### 2018-02-28 eval.Rmd
simple eval done

---
title: "Pteropod biomarker analysis"
author: ""
output: 
  html_document:
    code_folding: hide
self_contained: yes
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, fig.path = 'figs/', dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)

source("R/funcs.R")

data(envdat)
data(biodat)
```

Pairs plot of biomarkers and environmental data
```{r, fig.height = 8, fig.width = 8}
dat <- biodat %>% 
  left_join(envdat, by = 'CTD') %>% 
  mutate_if(is.character, as.numeric) %>% 
  gather('var', 'val', -CTD) %>% 
  group_by(CTD, var) %>% 
  summarise(val = mean(val, na.rm = TRUE)) %>% 
  ungroup %>% 
  spread(var, val)

pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
```

RDA triplot of biomarkers and environmental data
```{r, fig.height = 4, fig.width = 5}
dat_frm <- dat %>% 
  mutate(
    CAT = ifelse(is.nan(CAT), mean(CAT, na.rm = TRUE), CAT)
  ) %>% 
  data.frame %>% 
  column_to_rownames('CTD') %>%
  decostand(method = 'range')

env <- select(dat_frm, Ara, O2, pCO2, Sal, TA, Temp)
bio <- select(dat_frm, CAT, GR, GSHonGSSG, GST, LPX, ORAC, SOD)

mod <- rda(bio, env)
ggord(mod, ptslab = T, parse = T)
```

```{r}

# pearson, spearman correlation between biomarkers and env variables
cr <- dat_frm %>%
  gather('env_lab', 'env_val', Ara, O2, pCO2, Sal, TA, Temp) %>% 
  gather('bio_lab', 'bio_val', CAT, GR, GSHonGSSG, GST, LPX, ORAC, SOD) %>% 
  group_by(env_lab, bio_lab) %>% 
  summarise(
    sp_cr = round(cor.test(env_val, bio_val, method = 'spearman')$estimate, 2),
    sp_pv = p_ast(cor.test(env_val, bio_val, method = 'spearman')$p.value),
    pr_cr = round(cor.test(env_val, bio_val, method = 'pearson')$estimate, 2),
    pr_pv = p_ast(cor.test(env_val, bio_val, method = 'pearson')$p.value)
  ) %>% 
  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
  unite('pr_cr', pr_cr, pr_pv, sep = ' ')

sp_cr <- cr %>% 
  select(-pr_cr) %>% 
  spread(env_lab, sp_cr)

pr_cr <- cr %>% 
  select(-sp_cr) %>% 
  spread(env_lab, pr_cr)
```

# {.tabset}

## Spearman correlations
```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

## Pearson correlations
```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

# {.tabset}

Pairwise models with interactions for each biomarker
```{r}
# variables to model, chr strings only
env_cmb <- names(env) %>% 
  combn(2) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>%
  crossing(names(bio), .) %>% 
  t %>% 
  data.frame(stringsAsFactors = F) %>% 
  as.list

# models
mods <- env_cmb %>%
  map(function(x){

    # formula
    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
      as.formula

    # model, formatted
    mod <- lm(frm, data = dat_frm) %>% 
      summary %>% 
      .$coefficients %>% 
      .[-1, c(1, 4)] %>% 
      data.frame %>% 
      rownames_to_column('env_lab') %>% 
      mutate(
        bio_lab = x[1],
        Est = round(Estimate, 2), 
        Pvl = p_ast(Pr...t..)
      ) %>% 
      select(bio_lab, env_lab, Est, Pvl)

    return(mod)

  }) %>% 
  enframe('Model') %>% 
  unnest %>% 
  mutate(Model = gsub('^X', 'mod', Model)) %>% 
  group_by(bio_lab) %>% 
  nest
```

## CAT
```{r}
knitr::kable(filter(mods, bio_lab %in% 'CAT')$data[[1]], caption = 'CAT')
```

## GR
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GR')$data[[1]], caption = 'GR')
```

## GSHonGSSG
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GSHonGSSG')$data[[1]], caption = 'GSHonGSSG')
```

## GST
```{r}
knitr::kable(filter(mods, bio_lab %in% 'GST')$data[[1]], caption = 'GST')
```

## LPX
```{r}
knitr::kable(filter(mods, bio_lab %in% 'LPX')$data[[1]], caption = 'LPX')
```

## ORAC
```{r}
knitr::kable(filter(mods, bio_lab %in% 'ORAC')$data[[1]], caption = 'ORAC')
```

## SOD
```{r}
knitr::kable(filter(mods, bio_lab %in% 'SOD')$data[[1]], caption = 'SOD')
```

### 2018-02-28 R/dat_proc.R

@@ -1,17 +1,20 @@
 library(readxl)
 library(tidyverse)
 
-envdat <- read_excel('C:/Users/Marcus.SCCWRP2K/Desktop/MeanSurface100m2016WCOA.xlsx') %>% 
+envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
   select(`CTD Station`, `'pCO2'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
   rename(
     CTD = `CTD Station`, 
     Ara = `'Aragonite'`, 
     O2 = `'Oxygen'`,
-    pCO2 = `pCO2`,
+    pCO2 = `'pCO2'`,
     Temp = `'Temperature'`, 
     Sal = `'Salinity'`, 
     TA = `'Alkalinity'`
   )
-biodat <- read_excel('C:/Users/Marcus.SCCWRP2K/Desktop/PteropodIntegrated_oxibiomarkers.xlsx', 
+biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx', 
                      sheet = 'Pteropod30mbis') %>% 
-  select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOC)
+  select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD)
+
+save(envdat, file = 'data/envdat.RData', compress = 'xz')
+save(biodat, file = 'data/biodat.RData', compress = 'xz')

### 2018-03-01 eval.Rmd
new env vars, biomark ratio

@@
 data(biodat)
+
+envchr <- c('Ara', 'O2', 'pH', 'Fluor', 'CO3', 'pCO2', 'Sal', 'TA', 'Temp', 'Latitude')
+biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
 ```
@@
 
-env <- select(dat_frm, Ara, O2, pCO2, Sal, TA, Temp)
-bio <- select(dat_frm, CAT, GR, GSHonGSSG, GST, LPX, ORAC, SOD)
+env <- select(dat_frm, one_of(envchr))
+bio <- select(dat_frm, one_of(biochr))
 
@@
 cr <- dat_frm %>%
-  gather('env_lab', 'env_val', Ara, O2, pCO2, Sal, TA, Temp) %>% 
-  gather('bio_lab', 'bio_val', CAT, GR, GSHonGSSG, GST, LPX, ORAC, SOD) %>% 
+  gather('env_lab', 'env_val', one_of(envchr)) %>% 
+  gather('bio_lab', 'bio_val', one_of(biochr)) %>% 
   group_by(env_lab, bio_lab) %>% 
@@
 ```
+
+## ORACvLPX
+```{r}
+knitr::kable(filter(mods, bio_lab %in% 'ORACvLPX')$data[[1]], caption = 'ORACvLPX')
+```

### 2018-03-01 R/dat_proc.R

@@ -2,11 +2,14 @@ library(readxl)
 library(tidyverse)
 
 envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
-  select(`CTD Station`, `'pCO2'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
+  select(`CTD Station`, `'pCO2'`, `'pH'`, `'CO3'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
   rename(
     CTD = `CTD Station`, 
     Ara = `'Aragonite'`, 
     O2 = `'Oxygen'`,
+    pH = `'pH'`,
+    CO3 = `'CO3'`,
+    Fluor = `'Fluorescence'`,
     pCO2 = `'pCO2'`,
     Temp = `'Temperature'`, 
     Sal = `'Salinity'`, 
@@ -14,7 +17,10 @@ envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>%
   )
 biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx', 
                      sheet = 'Pteropod30mbis') %>% 
-  select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD)
+  select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD, Latitude) %>% 
+  mutate(
+    ORACvLPX = ORAC / LPX
+  )
 
 save(envdat, file = 'data/envdat.RData', compress = 'xz')
 save(biodat, file = 'data/biodat.RData', compress = 'xz')

### 2018-03-01 R/dat_proc.R
no LPX outlier

@@ -18,6 +18,7 @@ envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>%
 biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx', 
                      sheet = 'Pteropod30mbis') %>% 
   select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD, Latitude) %>% 
+  filter(LPX < 4) %>% 
   mutate(
     ORACvLPX = ORAC / LPX
   )

### 2018-03-01 eval.Rmd
removed latitude, sal, ta, fluor, co3
fixed pairs plot with envchr, biochr select

@@
 
-envchr <- c('Ara', 'O2', 'pH', 'Fluor', 'CO3', 'pCO2', 'Sal', 'TA', 'Temp', 'Latitude')
+envchr <- c('Ara', 'O2', 'pH', 'pCO2', 'Temp')
 biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
@@
   mutate_if(is.character, as.numeric) %>% 
+  select(one_of(c(envchr, biochr, 'CTD'))) %>% 
   gather('var', 'val', -CTD) %>% 

### 2018-03-01 R/dat_proc.R
lpx outlier as average

@@ -18,7 +18,10 @@ envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>%
 biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx', 
                      sheet = 'Pteropod30mbis') %>% 
   select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD, Latitude) %>% 
-  filter(LPX < 4) %>% 
+  mutate(
+    LPX = ifelse(LPX > 4, NA, LPX),
+    LPX = ifelse(is.na(LPX), mean(LPX, na.rm = T), LPX)
+    ) %>% 
   mutate(
     ORACvLPX = ORAC / LPX
   )

### 2018-03-01 eval.Rmd
lpx outlier as average
working with reactive effects plot

@@
 self_contained: yes
+runtime: shiny
 ---
@@
 library(vegan)
-library(ggord)
+# library(ggord)
+library(scales)
+library(effects)
+library(shiny)
 
@@
 
-envchr <- c('Ara', 'O2', 'pH', 'pCO2', 'Temp')
+envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
 biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
@@
 RDA triplot of biomarkers and environmental data
-```{r, fig.height = 4, fig.width = 5}
+```{r, fig.height = 4, fig.width = 5, eval = F}
 dat_frm <- dat %>% 
@@
   data.frame %>% 
-  column_to_rownames('CTD') %>%
+  column_to_rownames('CTD')
+
+env <- select(dat_frm, one_of(envchr)) %>% 
   decostand(method = 'range')
-
-env <- select(dat_frm, one_of(envchr))
-bio <- select(dat_frm, one_of(biochr))
+bio <- select(dat_frm, one_of(biochr)) %>% 
+  decostand(method = 'range')
 
@@
 
+```{r spheat, fig.height = 4, fig.width = 5}
+spplo <- cr %>% 
+  select(-pr_cr) %>% 
+  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
+  mutate(
+    cor = as.numeric(cor), 
+    bio_lab = factor(bio_lab, levels = rev(unique(bio_lab)))
+  )
+
+pbase <- theme(
+  panel.grid.major = element_blank(), 
+  panel.grid.minor = element_blank(), 
+  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+  axis.text.y = element_text(size = 8),
+  legend.position = 'top', 
+  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+  strip.background = element_blank(), 
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  )
+
+p <- ggplot(spplo, aes(y = bio_lab, x = env_lab, fill = cor)) + 
+  geom_tile(colour = 'black') + 
+  pbase +
+  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
+  scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
+  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
+p
+```
+
 ## Pearson correlations
@@
 knitr::kable(pr_cr, caption = 'Pearson correlations')
+```
+
+```{r prheat, fig.height = 4, fig.width = 5}
+prplo <- cr %>% 
+  select(-sp_cr) %>% 
+  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
+  mutate(
+    cor = as.numeric(cor), 
+    bio_lab = factor(bio_lab, levels = rev(unique(bio_lab)))
+  )
+
+pbase <- theme(
+  panel.grid.major = element_blank(), 
+  panel.grid.minor = element_blank(), 
+  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+  axis.text.y = element_text(size = 8),
+  legend.position = 'top', 
+  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+  strip.background = element_blank(), 
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  )
+
+p <- ggplot(prplo, aes(y = bio_lab, x = env_lab, fill = cor)) + 
+  geom_tile(colour = 'black') + 
+  pbase +
+  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
+  scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
+  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
+p
 ```
@@
 # models
-mods <- env_cmb %>%
+mod_all <- env_cmb %>%
   map(function(x){
@@
 
-    # model, formatted
-    mod <- lm(frm, data = dat_frm) %>% 
+    # model only
+    lm(frm, data = dat_frm)
+
+  }) %>% 
+  enframe('Model', 'Modobj') %>% 
+  mutate(
+    Model = gsub('^X', 'mod', Model), 
+    data = map(Modobj, function(x){
+
+      # response
+      yvar <- all.vars(formula(x))[1]
+
+      # coeff summary
+      out <- x %>% 
         summary %>% 
@@
         mutate(
-        bio_lab = x[1],
+          bio_lab = yvar,
           Est = round(Estimate, 2), 
@@
 
-    return(mod)
+    return(out)
 
-  }) %>% 
-  enframe('Model') %>% 
-  unnest %>% 
-  mutate(Model = gsub('^X', 'mod', Model)) %>% 
-  group_by(bio_lab) %>% 
-  nest
+    })
+  ) 
+mods <- mod_all %>% 
+  unnest(data)
+
+```
+
+```{r}
+selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))
+
+md <- reactive({
+
+  md <- filter(mod_all, Model %in% input$mod) %>% 
+  .$Modobj %>% 
+  .[[1]]
+
+  return(md)
+
+})
+
+levs1 <- reactive({
+
+  levs1 <- quantile(md()$model[, 2],  c(0.25, 0.75)) %>% 
+    list
+  names(levs1) <- names(md()$model)[2]
+  return(levs1)
+
+})
+
+levs2 <- reactive({
+
+  levs2 <- quantile(md()$model[, 3], c(0.25, 0.75)) %>% 
+    list
+  names(levs2) <- names(md()$model)[3]
+  return(levs2)
+
+})
+
+trm <- reactive({
+
+  md() %>% coefficients %>% names %>% .[4]
+
+})
+
+renderPlot({
+  plot(effect(trm(), md(), xlevels = levs1()), multiline = T)
+}, height = 300, width = 600)
+
+renderPlot({
+  plot(effect(trm(), md(), xlevels = levs2()), multiline = T)
+}, height = 300, width = 600)
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'CAT')$data[[1]], caption = 'CAT')
+knitr::kable(filter(mods, bio_lab %in% 'CAT'), caption = 'CAT')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'GR')$data[[1]], caption = 'GR')
+knitr::kable(filter(mods, bio_lab %in% 'GR'), caption = 'GR')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'GSHonGSSG')$data[[1]], caption = 'GSHonGSSG')
+knitr::kable(filter(mods, bio_lab %in% 'GSHonGSSG'), caption = 'GSHonGSSG')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'GST')$data[[1]], caption = 'GST')
+knitr::kable(filter(mods, bio_lab %in% 'GST'), caption = 'GST')
 ```
@@
 ## LPX
+
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'LPX')$data[[1]], caption = 'LPX')
+knitr::kable(filter(mods, bio_lab %in% 'LPX'), caption = 'LPX')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'ORAC')$data[[1]], caption = 'ORAC')
+knitr::kable(filter(mods, bio_lab %in% 'ORAC'), caption = 'ORAC')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'SOD')$data[[1]], caption = 'SOD')
+knitr::kable(filter(mods, bio_lab %in% 'SOD'), caption = 'SOD')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'ORACvLPX')$data[[1]], caption = 'ORACvLPX')
+knitr::kable(filter(mods, bio_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
 ```

### 2018-03-01 R/dat_proc.R

@@ -18,10 +18,10 @@ envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>%
 biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx', 
                      sheet = 'Pteropod30mbis') %>% 
   select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD, Latitude) %>% 
-  mutate(
-    LPX = ifelse(LPX > 4, NA, LPX),
-    LPX = ifelse(is.na(LPX), mean(LPX, na.rm = T), LPX)
-    ) %>% 
+  # mutate(
+  #   LPX = ifelse(LPX > 4, NA, LPX),
+  #   LPX = ifelse(is.na(LPX), mean(LPX, na.rm = T), LPX)
+  #   ) %>% 
   mutate(
     ORACvLPX = ORAC / LPX
   )

### 2018-03-01 eval.Rmd
trying to fix on shiny server

@@
 library(vegan)
-# library(ggord)
+library(ggord)
 library(scales)
@@
 RDA triplot of biomarkers and environmental data
-```{r, fig.height = 4, fig.width = 5, eval = F}
+```{r, fig.height = 4, fig.width = 5, eval = T}
 dat_frm <- dat %>% 

### 2018-03-02 index.Rmd
eval.Rmd changed to index.Rmd
renderPrint and ggplot for rmarkdown shiny

--- 2018-03-01 eval.Rmd
@@
 
+      # rsq
+      rsq <- summary(x) %>% 
+        .$r.squared
+
       # coeff summary
@@
           bio_lab = yvar,
+          Rsq = round(rsq, 2),
           Est = round(Estimate, 2), 
@@
         ) %>% 
-        select(bio_lab, env_lab, Est, Pvl)
+        select(bio_lab, env_lab, Rsq, Est, Pvl)
 
@@
 
+renderPrint({summary(md())})
+
+renderPlot({
+  toplo <- md()$model %>% 
+    gather('var', 'val', -1)
+
+  ggplot(toplo, aes_string(x='val', y=names(toplo)[1])) +
+    geom_point() + 
+    facet_wrap(~var, ncol = 2, scales = 'free_x') + 
+    theme_bw(base_family = 'serif') + 
+    geom_smooth(method = 'lm') + 
+    theme(axis.title.x = element_blank())
+
+}, height = 300, width = 600)
+
 renderPlot({
@@
 ## LPX
-
 ```{r}

### 2018-03-02 index.Rmd
removed figs folder for rmarkdown shiny server

@@
 library(knitr)
-opts_chunk$set(warning = FALSE, message = FALSE, fig.path = 'figs/', dev.args = list(bg = 'transparent'), eval = T)
+opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)
 

### 2018-03-02 index.Rmd
heatplot for model predictions

@@
 }, height = 300, width = 600)
+
+renderPlot({
+
+  x <- md()$model %>% 
+    .[, -1] %>% 
+    data.frame %>% 
+    as.list %>% 
+    map(range) %>%
+    map(function(x) seq(x[1], x[2], length = 100))
+
+  nms <- names(x) 
+  x <- crossing(x[[1]], x[[2]])
+  names(x) <- nms
+
+  prd_vl <- predict(md(), newdata = x) %>% 
+    data.frame(., x)
+  names(prd_vl)[1] <- all.vars(formula(md()))[1]
+
+  pbase <- theme(
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(), 
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+    axis.text.y = element_text(size = 8),
+    legend.position = 'top', 
+    plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+    strip.background = element_blank(), 
+    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  )
+
+  p <- ggplot(prd_vl, aes_string(y = names(prd_vl)[2], x = names(prd_vl)[3], fill = names(prd_vl)[1])) + 
+    geom_tile(colour = 'black') + 
+    pbase +
+    scale_y_discrete(expand = c(0, 0)) + 
+    scale_x_discrete(expand = c(0, 0)) +
+    scale_fill_distiller(palette = 'Spectral') +
+    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 12, angle = 0)))
+  p
+}, height = 700, width = 700)
 ```

### 2018-03-02 index.Rmd
messing with plotly, added heatmap

@@
 library(shiny)
+library(plotly)
 
@@
 
+prd_vl <- reactive({
+
+  x <- md()$model %>% 
+    .[, -1] %>% 
+    data.frame %>% 
+    as.list %>% 
+    map(range) %>%
+    map(function(x) seq(x[1], x[2], length = 100))
+
+  nms <- names(x) 
+  x <- crossing(x[[1]], x[[2]])
+  names(x) <- nms
+
+  prd_vl <- predict(md(), newdata = x) %>% 
+    data.frame(., x)
+  names(prd_vl)[1] <- all.vars(formula(md()))[1]
+
+  return(prd_vl)
+
+})
+
 renderPrint({summary(md())})
@@
 
-  x <- md()$model %>% 
-    .[, -1] %>% 
-    data.frame %>% 
-    as.list %>% 
-    map(range) %>%
-    map(function(x) seq(x[1], x[2], length = 100))
-
-  nms <- names(x) 
-  x <- crossing(x[[1]], x[[2]])
-  names(x) <- nms
-
-  prd_vl <- predict(md(), newdata = x) %>% 
-    data.frame(., x)
-  names(prd_vl)[1] <- all.vars(formula(md()))[1]
-
   pbase <- theme(
@@
 
-  p <- ggplot(prd_vl, aes_string(y = names(prd_vl)[2], x = names(prd_vl)[3], fill = names(prd_vl)[1])) + 
+  p <- ggplot(prd_vl(), aes_string(y = names(prd_vl())[2], x = names(prd_vl())[3], fill = names(prd_vl())[1])) + 
     geom_tile(colour = 'black') + 
+    theme_bw(base_family = 'serif') +
     pbase +
@@
     scale_fill_distiller(palette = 'Spectral') +
-    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 12, angle = 0)))
+    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 15, label.theme = element_text(size = 12, angle = 0)))
   p
 }, height = 700, width = 700)
+
+renderPlotly({
+
+  dat <- prd_vl()[, 1] %>% 
+    as.matrix
+
+  scene <- list(
+    aspectmode = 'manual', 
+    aspectratio = list(x = 0.5, y = 1, z = 0.3)
+  )
+
+  p <- plot_ly(z = ~dat) %>% add_surface() %>% 
+    layout(scene = scene)
+  p
+
+})
 ```

### 2018-03-02 index.Rmd
Fixed plotly, corrected axis labels

@@
 
-  dat <- prd_vl()[, 1] %>% 
+  dat <- prd_vl() %>% 
+    spread_(key_col = names(.)[2], value_col = names(.)[1]) %>% 
+    .[,-1] %>% 
     as.matrix
@@
     aspectmode = 'manual', 
-    aspectratio = list(x = 0.5, y = 1, z = 0.3)
+    aspectratio = list(x = 1, y = 1, z = 0.5), 
+    xaxis = list(title = names(prd_vl()[3])),
+    yaxis = list(title = names(prd_vl()[2])),
+    zaxis = list(title = names(prd_vl()[1]))
     )
 
-  p <- plot_ly(z = ~dat) %>% add_surface() %>% 
+  p <- plot_ly(z = ~dat) %>% add_surface(colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
     layout(scene = scene)

### 2018-03-05 index.Rmd

@@
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
+envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
 biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
@@
     panel.grid.minor = element_blank(), 
-    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
-    axis.text.y = element_text(size = 8),
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
+    axis.text.y = element_text(size = 10),
     legend.position = 'top', 
@@
     pbase +
-    scale_y_discrete(expand = c(0, 0)) + 
-    scale_x_discrete(expand = c(0, 0)) +
+    scale_y_continuous(expand = c(0, 0)) +
+    scale_x_continuous(expand = c(0, 0)) +
     scale_fill_distiller(palette = 'Spectral') +

### 2018-03-05 R/dat_proc.R
processing phys data

@@ -1,6 +1,7 @@
 library(readxl)
 library(tidyverse)
 
+# environmental data
 envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
   select(`CTD Station`, `'pCO2'`, `'pH'`, `'CO3'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
   rename(
@@ -14,17 +15,67 @@ envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>%
     Temp = `'Temperature'`, 
     Sal = `'Salinity'`, 
     TA = `'Alkalinity'`
-  )
+  ) %>% 
+  mutate_if(is.character, as.numeric)
+
+# biomarker data
 biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx', 
                      sheet = 'Pteropod30mbis') %>% 
   select(CTD, GSHonGSSG, GST, GR, LPX, CAT, ORAC, SOD, Latitude) %>% 
-  # mutate(
-  #   LPX = ifelse(LPX > 4, NA, LPX),
-  #   LPX = ifelse(is.na(LPX), mean(LPX, na.rm = T), LPX)
-  #   ) %>% 
   mutate(
     ORACvLPX = ORAC / LPX
-  )
+  ) %>% 
+  gather('var', 'val', -CTD) %>% 
+  group_by(CTD, var) %>% 
+  summarize(val = mean(val, na.rm = T)) %>% 
+  group_by(var) %>% 
+  mutate(val = ifelse(is.na(val), mean(val, na.rm = T), val)) %>% 
+  spread(var, val) %>% 
+  ungroup
 
+# abundance
+abudat <- read_excel('raw/Copy of abundances to be usef_Marcus2.xlsx') %>% 
+  rename(
+    CTD = `CTD station`,
+    abund = `real abundance`
+  ) %>% 
+  group_by(CTD) %>% 
+  summarize(
+    abu = mean(abund, na.rm = T)
+  ) %>% 
+  ungroup
+
+# shell dissolution
+disdat <- read_excel('raw/parameters for WCOA 2016 combined.xlsx', sheet = 'dissolution') %>% 
+  fill(station) %>% 
+  rename( 
+    CTD = station, 
+    dis = `dissolution extent`
+    ) %>% 
+  group_by(CTD) %>% 
+  summarise(
[…]

### 2018-03-05 biomark.Rmd
initial physio comp done

--- 2018-03-05 index.Rmd
@@
 data(envdat)
-data(biodat)
+data(ptedat)
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
-biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
+envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
+ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
 ```
@@
 ```{r, fig.height = 8, fig.width = 8}
-dat <- biodat %>% 
-  left_join(envdat, by = 'CTD') %>% 
-  mutate_if(is.character, as.numeric) %>% 
-  select(one_of(c(envchr, biochr, 'CTD'))) %>% 
-  gather('var', 'val', -CTD) %>% 
-  group_by(CTD, var) %>% 
-  summarise(val = mean(val, na.rm = TRUE)) %>%
-  ungroup %>% 
-  spread(var, val)
+dat <- ptedat %>% 
+  select(one_of('CTD', ptechr)) %>% 
+  na.omit %>% 
+  inner_join(envdat, by = 'CTD') %>% 
+  select(one_of(c('CTD', envchr, ptechr)))
 
@@
 dat_frm <- dat %>% 
-  mutate(
-    CAT = ifelse(is.nan(CAT), mean(CAT, na.rm = TRUE), CAT)
-  ) %>% 
   data.frame %>% 
@@
   decostand(method = 'range')
-bio <- select(dat_frm, one_of(biochr)) %>% 
+pte <- select(dat_frm, one_of(ptechr)) %>% 
   decostand(method = 'range')
 
-mod <- rda(bio, env)
+mod <- rda(pte, env)
 ggord(mod, ptslab = T, parse = T)
@@
   gather('env_lab', 'env_val', one_of(envchr)) %>% 
-  gather('bio_lab', 'bio_val', one_of(biochr)) %>% 
-  group_by(env_lab, bio_lab) %>% 
+  gather('pte_lab', 'pte_val', one_of(ptechr)) %>% 
+  group_by(env_lab, pte_lab) %>% 
   summarise(
-    sp_cr = round(cor.test(env_val, bio_val, method = 'spearman')$estimate, 2),
-    sp_pv = p_ast(cor.test(env_val, bio_val, method = 'spearman')$p.value),
-    pr_cr = round(cor.test(env_val, bio_val, method = 'pearson')$estimate, 2),
-    pr_pv = p_ast(cor.test(env_val, bio_val, method = 'pearson')$p.value)
+    sp_cr = round(cor.test(env_val, pte_val, method = 'spearman')$estimate, 2),
+    sp_pv = p_ast(cor.test(env_val, pte_val, method = 'spearman')$p.value),
+    pr_cr = round(cor.test(env_val, pte_val, method = 'pearson')$estimate, 2),
+    pr_pv = p_ast(cor.test(env_val, pte_val, method = 'pearson')$p.value)
   ) %>% 
@@
     cor = as.numeric(cor), 
-    bio_lab = factor(bio_lab, levels = rev(unique(bio_lab)))
+    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
   )
@@
 
-p <- ggplot(spplo, aes(y = bio_lab, x = env_lab, fill = cor)) + 
+p <- ggplot(spplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
   geom_tile(colour = 'black') + 
@@
 
-p <- ggplot(prplo, aes(y = bio_lab, x = env_lab, fill = cor)) + 
+p <- ggplot(prplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
   geom_tile(colour = 'black') + 
@@
   data.frame(stringsAsFactors = F) %>%
-  crossing(names(bio), .) %>% 
+  crossing(names(pte), .) %>% 
   t %>% 
@@
         mutate(
-          bio_lab = yvar,
+          pte_lab = yvar,
           Rsq = round(rsq, 2),
@@
         ) %>% 
-        select(bio_lab, env_lab, Rsq, Est, Pvl)
+        select(pte_lab, env_lab, Rsq, Est, Pvl)
 
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'CAT'), caption = 'CAT')
+knitr::kable(filter(mods, pte_lab %in% 'CAT'), caption = 'CAT')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'GR'), caption = 'GR')
+knitr::kable(filter(mods, pte_lab %in% 'GR'), caption = 'GR')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'GSHonGSSG'), caption = 'GSHonGSSG')
+knitr::kable(filter(mods, pte_lab %in% 'GSHonGSSG'), caption = 'GSHonGSSG')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'GST'), caption = 'GST')
+knitr::kable(filter(mods, pte_lab %in% 'GST'), caption = 'GST')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'LPX'), caption = 'LPX')
+knitr::kable(filter(mods, pte_lab %in% 'LPX'), caption = 'LPX')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'ORAC'), caption = 'ORAC')
+knitr::kable(filter(mods, pte_lab %in% 'ORAC'), caption = 'ORAC')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'SOD'), caption = 'SOD')
+knitr::kable(filter(mods, pte_lab %in% 'SOD'), caption = 'SOD')
 ```
@@
 ```{r}
-knitr::kable(filter(mods, bio_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
+knitr::kable(filter(mods, pte_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
 ```

### 2018-03-05 physio.Rmd
initial physio comp done

--- 2018-03-05 biomark.Rmd
@@
 ---
-title: "Pteropod biomarker analysis"
+title: "Pteropod physiological analysis"
 author: ""
@@
 envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
-ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
+ptechr <- c('abu', 'dis', 'len')
 ```
 
-Pairs plot of biomarkers and environmental data
+Pairs plot of physiological parameters and environmental data
 ```{r, fig.height = 8, fig.width = 8}
@@
   select(one_of('CTD', ptechr)) %>% 
-  na.omit %>% 
-  inner_join(envdat, by = 'CTD') %>% 
+  left_join(envdat, by = 'CTD') %>% 
   select(one_of(c('CTD', envchr, ptechr)))
@@
 
-RDA triplot of biomarkers and environmental data
+Pairs plot of physiological parameters and biomarkers
+```{r, fig.height = 8, fig.width = 8}
+biopr <- ptedat %>% 
+  select(-CTD, -Latitude)
+
+pairs(biopr, gap = 0, cex = 1, col = 'grey')
+```
+
+RDA triplot of physiological parameters and environmental data
 ```{r, fig.height = 4, fig.width = 5, eval = T}
-dat_frm <- dat %>% 
+pca_frm <- dat %>% 
+  na.omit %>% 
   data.frame %>% 
@@
 
-env <- select(dat_frm, one_of(envchr)) %>% 
+env <- select(pca_frm, one_of(envchr)) %>% 
   decostand(method = 'range')
-pte <- select(dat_frm, one_of(ptechr)) %>% 
+pte <- select(pca_frm, one_of(ptechr)) %>% 
   decostand(method = 'range')
@@
 ```{r}
-
-# pearson, spearman correlation between biomarkers and env variables
-cr <- dat_frm %>%
+# pearson, spearman correlation between physiological parameters and env variables
+cr <- dat %>%
   gather('env_lab', 'env_val', one_of(envchr)) %>% 
@@
 
-# {.tabset}
+# Correlation with environmental variables {.tabset}
 
-## Spearman correlations
+## Spearman correlations, environmental variables
 ```{r}
@@
 
-```{r spheat, fig.height = 4, fig.width = 5}
+```{r, fig.height = 4, fig.width = 5}
 spplo <- cr %>% 
@@
   pbase +
-  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
+  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
   scale_x_discrete('Environment', expand = c(0, 0)) +
@@
 
-## Pearson correlations
+## Pearson correlations, environmental variables
+
 ```{r}
@@
 
-```{r prheat, fig.height = 4, fig.width = 5}
+```{r, fig.height = 4, fig.width = 5}
 prplo <- cr %>% 
@@
 
-pbase <- theme(
-  panel.grid.major = element_blank(), 
-  panel.grid.minor = element_blank(), 
-  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
-  axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
-  strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
-  )
-
 p <- ggplot(prplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
@@
   pbase +
-  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
+  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
   scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
+  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
+p
+```
+
+# Correlation with biomarkers {.tabset}
+
+```{r}
+# pearson, spearman correlation between physiological parameters and biomarkers
+cr <- ptedat %>% 
+  gather('phy_lab', 'phy_val', one_of(ptechr)) %>% 
+  gather('bio_lab', 'bio_val', -phy_lab, -phy_val, -CTD) %>% 
+  group_by(phy_lab, bio_lab) %>% 
+  summarise(
+    sp_cr = round(cor.test(phy_val, bio_val, method = 'spearman')$estimate, 2),
+    sp_pv = p_ast(cor.test(phy_val, bio_val, method = 'spearman')$p.value),
+    pr_cr = round(cor.test(phy_val, bio_val, method = 'pearson')$estimate, 2),
+    pr_pv = p_ast(cor.test(phy_val, bio_val, method = 'pearson')$p.value)
+  ) %>% 
+  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
+  unite('pr_cr', pr_cr, pr_pv, sep = ' ') %>% 
+  ungroup
+
+sp_cr <- cr %>% 
+  select(-pr_cr) %>% 
+  spread(bio_lab, sp_cr)
+
+pr_cr <- cr %>% 
+  select(-sp_cr) %>% 
+  spread(bio_lab, pr_cr)
+```
+
+## Spearman correlations, biomarkers
+
+```{r}
+knitr::kable(sp_cr, caption = 'Spearman rank correlations')
+```
+
+```{r, fig.height = 4, fig.width = 5}
+spplo <- cr %>% 
+  select(-pr_cr) %>% 
+  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
+  mutate(
+    cor = as.numeric(cor), 
+    phy_lab = factor(phy_lab, levels = rev(unique(phy_lab)))
+  )
+
+p <- ggplot(spplo, aes(y = phy_lab, x = bio_lab, fill = cor)) + 
+  geom_tile(colour = 'black') + 
+  pbase +
+  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
+  scale_x_discrete('Biomarker', expand = c(0, 0)) +
+  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
+  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
+p
+```
+
+## Pearson correlations, biomarkers
+
+```{r}
+knitr::kable(pr_cr, caption = 'Pearson correlations')
+```
+
+```{r, fig.height = 4, fig.width = 5}
+prplo <- cr %>% 
+  select(-sp_cr) %>% 
+  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
+  mutate(
+    cor = as.numeric(cor), 
+    phy_lab = factor(phy_lab, levels = rev(unique(phy_lab)))
+  )
+
+p <- ggplot(prplo, aes(y = phy_lab, x = bio_lab, fill = cor)) + 
+  geom_tile(colour = 'black') + 
+  pbase +
+  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
+  scale_x_discrete('Biomarker', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 
-Pairwise models with interactions for each biomarker
+Pairwise models with interactions for each physiological parameter
 ```{r}
@@
     # model only
-    lm(frm, data = dat_frm)
+    lm(frm, data = dat)
 
@@
 
-## CAT
+## abundance
 ```{r}
-knitr::kable(filter(mods, pte_lab %in% 'CAT'), caption = 'CAT')
+knitr::kable(filter(mods, pte_lab %in% 'abu'), caption = 'abu')
 ```
 
-## GR
+## dissolution
 ```{r}
-knitr::kable(filter(mods, pte_lab %in% 'GR'), caption = 'GR')
+knitr::kable(filter(mods, pte_lab %in% 'dis'), caption = 'dis')
 ```
 
-## GSHonGSSG
+## length
 ```{r}
-knitr::kable(filter(mods, pte_lab %in% 'GSHonGSSG'), caption = 'GSHonGSSG')
+knitr::kable(filter(mods, pte_lab %in% 'len'), caption = 'len')
 ```
-
-## GST
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'GST'), caption = 'GST')
-```
-
-## LPX
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'LPX'), caption = 'LPX')
-```
-
-## ORAC
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'ORAC'), caption = 'ORAC')
-```
-
-## SOD
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'SOD'), caption = 'SOD')
-```
-
-## ORACvLPX
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
-```

### 2018-03-06 biomark.Rmd
added fluor to biomark and physio

@@
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
+envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
 ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
@@
 
-  levs1 <- quantile(md()$model[, 2],  c(0.25, 0.75)) %>% 
+  levs1 <- quantile(md()$model[, 2],  c(0, 1)) %>% 
     list
@@
 
-  levs2 <- quantile(md()$model[, 3], c(0.25, 0.75)) %>% 
+  levs2 <- quantile(md()$model[, 3], c(0, 1)) %>% 
     list

### 2018-03-06 physio.Rmd
added fluor to biomark and physio

@@
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp')
+envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
 ptechr <- c('abu', 'dis', 'len')

### 2018-03-06 R/dat_proc.R

@@ -39,9 +39,12 @@ abudat <- read_excel('raw/Copy of abundances to be usef_Marcus2.xlsx') %>%
     CTD = `CTD station`,
     abund = `real abundance`
   ) %>% 
+  mutate(
+    abu = log10(1 + abund)
+  ) %>% 
   group_by(CTD) %>% 
   summarize(
-    abu = mean(abund, na.rm = T)
+    abu = mean(abu, na.rm = T)
   ) %>% 
   ungroup
 
@@ -52,6 +55,7 @@ disdat <- read_excel('raw/parameters for WCOA 2016 combined.xlsx', sheet = 'diss
     CTD = station, 
     dis = `dissolution extent`
     ) %>% 
+  mutate(dis = asin(dis / 100)) %>% 
   group_by(CTD) %>% 
   summarise(
     dis = mean(dis, na.rm = T)

### 2018-03-09 physio.Rmd
rda loadings to biplot, adding ty2, ty3, scarring to physio

@@
 envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
-ptechr <- c('abu', 'dis', 'len')
+ptechr <- c('abu', 'dis', 'len', 'ty2', 'ty3', 'scr')
 ```
@@
 ggord(mod, ptslab = T, parse = T)
+```
+
+RDA axis loadings
+```{r}
+summary(mod)$biplot %>% round(2) %>% kable
 ```
@@
 
+## type 2 dissolution
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'ty2'), caption = 'ty2')
+```
+
+## type 3 dissolution
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'ty3'), caption = 'ty3')
+```
+
+## scarring
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'scr'), caption = 'scr')
+```
+
 ## length

### 2018-03-09 R/dat_proc.R

@@ -53,12 +53,18 @@ disdat <- read_excel('raw/parameters for WCOA 2016 combined.xlsx', sheet = 'diss
   fill(station) %>% 
   rename( 
     CTD = station, 
-    dis = `dissolution extent`
+    dis = `dissolution extent`, 
+    ty2 = `type II`,
+    ty3 = `type III`,
+    scr = scarring
     ) %>% 
   mutate(dis = asin(dis / 100)) %>% 
   group_by(CTD) %>% 
   summarise(
-    dis = mean(dis, na.rm = T)
+    dis = mean(dis, na.rm = T), 
+    ty2 = mean(ty2, na.rm = T),
+    ty3 = mean(ty3, na.rm = T),
+    scr = mean(scr, na.rm = T)
   ) %>% 
   ungroup
 

### 2018-03-09 biomark.Rmd
rda loadings to biplot, adding ty2, ty3, scarring to physio
better correlation matrices, vif for pairwise mods

@@
     code_folding: hide
+    css: kable.css
 self_contained: yes
@@
 
-Pairs plot of biomarkers and environmental data
+Pairs plot
 ```{r, fig.height = 8, fig.width = 8}
@@
 
+# {.tabset}
+
 ```{r}
+# all correlations
+crs <- crossing(var1 = names(dat_frm), var2 = names(dat_frm)) %>% 
+  filter(var1 != var2) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
+    crs = map(data, function(x){
 
-# pearson, spearman correlation between biomarkers and env variables
-cr <- dat_frm %>%
-  gather('env_lab', 'env_val', one_of(envchr)) %>% 
-  gather('pte_lab', 'pte_val', one_of(ptechr)) %>% 
-  group_by(env_lab, pte_lab) %>% 
-  summarise(
-    sp_cr = round(cor.test(env_val, pte_val, method = 'spearman')$estimate, 2),
-    sp_pv = p_ast(cor.test(env_val, pte_val, method = 'spearman')$p.value),
-    pr_cr = round(cor.test(env_val, pte_val, method = 'pearson')$estimate, 2),
-    pr_pv = p_ast(cor.test(env_val, pte_val, method = 'pearson')$p.value)
+      # variables
+      vr1 <- dat_frm[[x$var1]]
+      vr2 <- dat_frm[[x$var2]]
+
+      # pearson
+      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
+      pr_cr <- round(pr_ts$estimate, 2)
+      pr_pv <- p_ast(pr_ts$p.value)
+      pr <- paste(pr_cr, pr_pv)
+
+      # spearman
+      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
+      sp_cr <- round(sp_ts$estimate, 2)
+      sp_pv <- p_ast(sp_ts$p.value)
+      sp <- paste(sp_cr, sp_pv)
+
+      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
+      return(out)
+
+    })
   ) %>% 
-  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
-  unite('pr_cr', pr_cr, pr_pv, sep = ' ')
+  unnest %>% 
+  select(-rowname)
 
-sp_cr <- cr %>% 
-  select(-pr_cr) %>% 
-  spread(env_lab, sp_cr)
+sp_cr <- crs %>% 
+  select(var1, var2, sp) %>% 
+  spread(var2, sp) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1')
 
-pr_cr <- cr %>% 
-  select(-sp_cr) %>% 
-  spread(env_lab, pr_cr)
+pr_cr <- crs %>% 
+  select(var1, var2, pr) %>% 
+  spread(var2, pr) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1') 
 ```
 
-# {.tabset}
+## Spearman correlations
+<div class="verysmall">
 
-## Spearman correlations
 ```{r}
@@
 
-```{r spheat, fig.height = 4, fig.width = 5}
-spplo <- cr %>% 
-  select(-pr_cr) %>% 
-  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
+</div>
+
+```{r spheat, fig.height = 5, fig.width = 10}
+spplo <- crs %>% 
+  select(-pr) %>% 
+  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
   mutate(
     cor = as.numeric(cor), 
-    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
+    var1 = factor(var1, levels = rev(unique(.$var1)))
   )
@@
   strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
   )
 
-p <- ggplot(spplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
+p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
   geom_tile(colour = 'black') + 
   pbase +
-  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
-  scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 ## Pearson correlations
+<div class="verysmall">
+
 ```{r}
@@
 
-```{r prheat, fig.height = 4, fig.width = 5}
-prplo <- cr %>% 
-  select(-sp_cr) %>% 
-  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
+</div>
+
+```{r prheat, fig.height = 5, fig.width = 10}
+prplo <- crs %>% 
+  select(-sp) %>% 
+  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
   mutate(
     cor = as.numeric(cor), 
-    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
+    var1 = factor(var1, levels = rev(unique(.$var1)))
   )
@@
   strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
   )
 
-p <- ggplot(prplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
+p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
   geom_tile(colour = 'black') + 
   pbase +
-  scale_y_discrete('Biomarker', expand = c(0, 0)) + 
-  scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 ```
+
+# Pairwise VIF among environmental variables
+
+```{r}
+vifv <- crossing(var1 = envchr, var2 = envchr) %>% 
+  filter(var1 != var2) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
+    vifv = map(data, function(x){
+
+    vr1 <- x[['var1']]
+    vr2 <- x[['var2']]
+    frm <- formula(paste(vr1, '~', vr2))
+    mod <- lm(frm, dat_frm)
+    rsq <- mod %>% 
+      summary %>% 
+      .$r.squared
+    vif <- 1 / (1 - rsq)
+
+    return(vif)
+
+    })
+  ) %>% 
+  unnest %>% 
+  select(-rowname) %>% 
+  mutate(vifv = round(vifv, 2)) %>% 
+  spread(var2, vifv) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1') 
+```
+
+<div class="verysmall">
+
+```{r}
+knitr::kable(vifv, caption = 'VIF')
+```
+
+</div>
 

### 2018-03-09 physio.Rmd
better correlation matrices, vif for pairwise mods

@@
 
-Pairs plot of physiological parameters and environmental data
+Pairs plot
 ```{r, fig.height = 8, fig.width = 8}
@@
   select(one_of('CTD', ptechr)) %>% 
-  left_join(envdat, by = 'CTD') %>% 
+  na.omit %>% 
+  inner_join(envdat, by = 'CTD') %>% 
   select(one_of(c('CTD', envchr, ptechr)))
@@
 
-Pairs plot of physiological parameters and biomarkers
-```{r, fig.height = 8, fig.width = 8}
-biopr <- ptedat %>% 
-  select(-CTD, -Latitude)
-
-pairs(biopr, gap = 0, cex = 1, col = 'grey')
-```
-
-RDA triplot of physiological parameters and environmental data
+RDA triplot of biomarkers and environmental data
 ```{r, fig.height = 4, fig.width = 5, eval = T}
-pca_frm <- dat %>% 
-  na.omit %>% 
+dat_frm <- dat %>% 
   data.frame %>% 
@@
 
-env <- select(pca_frm, one_of(envchr)) %>% 
+env <- select(dat_frm, one_of(envchr)) %>% 
   decostand(method = 'range')
-pte <- select(pca_frm, one_of(ptechr)) %>% 
+pte <- select(dat_frm, one_of(ptechr)) %>% 
   decostand(method = 'range')
@@
 
-RDA axis loadings
+# Physiological correlations with biomarkers {.tabset}
+
 ```{r}
-summary(mod)$biplot %>% round(2) %>% kable
+tochk <- ptedat %>% 
+  select(-CTD, -Latitude)
+
+# all correlations
+crs <- crossing(var1 = names(tochk), var2 = names(tochk)) %>% 
+  filter(var1 != var2) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
+    crs = map(data, function(x){
+
+      # variables
+      vr1 <- tochk[[x$var1]]
+      vr2 <- tochk[[x$var2]]
+
+      # pearson
+      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
+      pr_cr <- round(pr_ts$estimate, 2)
+      pr_pv <- p_ast(pr_ts$p.value)
+      pr <- paste(pr_cr, pr_pv)
+
+      # spearman
+      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
+      sp_cr <- round(sp_ts$estimate, 2)
+      sp_pv <- p_ast(sp_ts$p.value)
+      sp <- paste(sp_cr, sp_pv)
+
+      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
+      return(out)
+
+    })
+  ) %>% 
+  unnest %>% 
+  select(-rowname)
+
+sp_cr <- crs %>% 
+  select(var1, var2, sp) %>% 
+  spread(var2, sp) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1')
+
+pr_cr <- crs %>% 
+  select(var1, var2, pr) %>% 
+  spread(var2, pr) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1') 
 ```
 
-```{r}
-# pearson, spearman correlation between physiological parameters and env variables
-cr <- dat %>%
-  gather('env_lab', 'env_val', one_of(envchr)) %>% 
-  gather('pte_lab', 'pte_val', one_of(ptechr)) %>% 
-  group_by(env_lab, pte_lab) %>% 
-  summarise(
-    sp_cr = round(cor.test(env_val, pte_val, method = 'spearman')$estimate, 2),
-    sp_pv = p_ast(cor.test(env_val, pte_val, method = 'spearman')$p.value),
-    pr_cr = round(cor.test(env_val, pte_val, method = 'pearson')$estimate, 2),
-    pr_pv = p_ast(cor.test(env_val, pte_val, method = 'pearson')$p.value)
-  ) %>% 
-  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
-  unite('pr_cr', pr_cr, pr_pv, sep = ' ')
+## Spearman correlations
 
-sp_cr <- cr %>% 
-  select(-pr_cr) %>% 
-  spread(env_lab, sp_cr)
+<div class="verysmall">
 
-pr_cr <- cr %>% 
-  select(-sp_cr) %>% 
-  spread(env_lab, pr_cr)
-```
-
-# Correlation with environmental variables {.tabset}
-
-## Spearman correlations, environmental variables
 ```{r}
@@
 
-```{r, fig.height = 4, fig.width = 5}
-spplo <- cr %>% 
-  select(-pr_cr) %>% 
-  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
+</div>
+
+```{r , fig.height = 5, fig.width = 10}
+spplo <- crs %>% 
+  select(-pr) %>% 
+  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
   mutate(
     cor = as.numeric(cor), 
-    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
+    var1 = factor(var1, levels = rev(unique(.$var1)))
   )
@@
   strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
   )
 
-p <- ggplot(spplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
+p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
   geom_tile(colour = 'black') + 
   pbase +
-  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
-  scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 
-## Pearson correlations, environmental variables
+## Pearson correlations
+<div class="verysmall">
 
@@
 
-```{r, fig.height = 4, fig.width = 5}
-prplo <- cr %>% 
-  select(-sp_cr) %>% 
-  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
+</div>
+
+```{r , fig.height = 5, fig.width = 10}
+prplo <- crs %>% 
+  select(-sp) %>% 
+  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
   mutate(
     cor = as.numeric(cor), 
-    pte_lab = factor(pte_lab, levels = rev(unique(pte_lab)))
+    var1 = factor(var1, levels = rev(unique(.$var1)))
   )
 
-p <- ggplot(prplo, aes(y = pte_lab, x = env_lab, fill = cor)) + 
+pbase <- theme(
+  panel.grid.major = element_blank(), 
+  panel.grid.minor = element_blank(), 
+  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+  axis.text.y = element_text(size = 8),
+  legend.position = 'top', 
+  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+  strip.background = element_blank(), 
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
+  )
+
+p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
   geom_tile(colour = 'black') + 
   pbase +
-  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
-  scale_x_discrete('Environment', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 
-# Correlation with biomarkers {.tabset}
+# Physiological correlations with environment {.tabset}
 
 ```{r}
-# pearson, spearman correlation between physiological parameters and biomarkers
-cr <- ptedat %>% 
-  gather('phy_lab', 'phy_val', one_of(ptechr)) %>% 
-  gather('bio_lab', 'bio_val', -phy_lab, -phy_val, -CTD) %>% 
-  group_by(phy_lab, bio_lab) %>% 
-  summarise(
-    sp_cr = round(cor.test(phy_val, bio_val, method = 'spearman')$estimate, 2),
-    sp_pv = p_ast(cor.test(phy_val, bio_val, method = 'spearman')$p.value),
-    pr_cr = round(cor.test(phy_val, bio_val, method = 'pearson')$estimate, 2),
-    pr_pv = p_ast(cor.test(phy_val, bio_val, method = 'pearson')$p.value)
+# all correlations
+crs <- crossing(var1 = names(dat_frm), var2 = names(dat_frm)) %>% 
+  filter(var1 != var2) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
+    crs = map(data, function(x){
+
+      # variables
+      vr1 <- dat_frm[[x$var1]]
+      vr2 <- dat_frm[[x$var2]]
+
+      # pearson
+      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
+      pr_cr <- round(pr_ts$estimate, 2)
+      pr_pv <- p_ast(pr_ts$p.value)
+      pr <- paste(pr_cr, pr_pv)
+
+      # spearman
+      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
+      sp_cr <- round(sp_ts$estimate, 2)
+      sp_pv <- p_ast(sp_ts$p.value)
+      sp <- paste(sp_cr, sp_pv)
+
+      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
+      return(out)
+
+    })
   ) %>% 
-  unite('sp_cr', sp_cr, sp_pv, sep = ' ') %>% 
-  unite('pr_cr', pr_cr, pr_pv, sep = ' ') %>% 
-  ungroup
+  unnest %>% 
+  select(-rowname)
 
-sp_cr <- cr %>% 
-  select(-pr_cr) %>% 
-  spread(bio_lab, sp_cr)
+sp_cr <- crs %>% 
+  select(var1, var2, sp) %>% 
+  spread(var2, sp) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1')
 
-pr_cr <- cr %>% 
-  select(-sp_cr) %>% 
-  spread(bio_lab, pr_cr)
+pr_cr <- crs %>% 
+  select(var1, var2, pr) %>% 
+  spread(var2, pr) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1') 
 ```
 
-## Spearman correlations, biomarkers
+## Spearman correlations
+<div class="verysmall">
 
@@
 
-```{r, fig.height = 4, fig.width = 5}
-spplo <- cr %>% 
-  select(-pr_cr) %>% 
-  separate(sp_cr, c('cor', 'sig'), sep = ' ') %>% 
+</div>
+
+```{r spheat, fig.height = 5, fig.width = 10}
+spplo <- crs %>% 
+  select(-pr) %>% 
+  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
   mutate(
     cor = as.numeric(cor), 
-    phy_lab = factor(phy_lab, levels = rev(unique(phy_lab)))
+    var1 = factor(var1, levels = rev(unique(.$var1)))
   )
 
-p <- ggplot(spplo, aes(y = phy_lab, x = bio_lab, fill = cor)) + 
+pbase <- theme(
+  panel.grid.major = element_blank(), 
+  panel.grid.minor = element_blank(), 
+  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+  axis.text.y = element_text(size = 8),
+  legend.position = 'top', 
+  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+  strip.background = element_blank(), 
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
+  )
+
+p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
   geom_tile(colour = 'black') + 
   pbase +
-  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
-  scale_x_discrete('Biomarker', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 
-## Pearson correlations, biomarkers
+## Pearson correlations
+<div class="verysmall">
 
@@
 
-```{r, fig.height = 4, fig.width = 5}
-prplo <- cr %>% 
-  select(-sp_cr) %>% 
-  separate(pr_cr, c('cor', 'sig'), sep = ' ') %>% 
+</div>
+
+```{r prheat, fig.height = 5, fig.width = 10}
+prplo <- crs %>% 
+  select(-sp) %>% 
+  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
   mutate(
     cor = as.numeric(cor), 
-    phy_lab = factor(phy_lab, levels = rev(unique(phy_lab)))
+    var1 = factor(var1, levels = rev(unique(.$var1)))
   )
 
-p <- ggplot(prplo, aes(y = phy_lab, x = bio_lab, fill = cor)) + 
+pbase <- theme(
+  panel.grid.major = element_blank(), 
+  panel.grid.minor = element_blank(), 
+  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+  axis.text.y = element_text(size = 8),
+  legend.position = 'top', 
+  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+  strip.background = element_blank(), 
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
+  )
+
+p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
   geom_tile(colour = 'black') + 
   pbase +
-  scale_y_discrete('Physio. param', expand = c(0, 0)) + 
-  scale_x_discrete('Biomarker', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
   scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
@@
 
+# Pairwise VIF among environmental variables
+
+```{r}
+vifv <- crossing(var1 = envchr, var2 = envchr) %>% 
+  filter(var1 != var2) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
+    vifv = map(data, function(x){
+
+    vr1 <- x[['var1']]
+    vr2 <- x[['var2']]
+    frm <- formula(paste(vr1, '~', vr2))
+    mod <- lm(frm, dat_frm)
+    rsq <- mod %>% 
+      summary %>% 
+      .$r.squared
+    vif <- 1 / (1 - rsq)
+
+    return(vif)
+
+    })
+  ) %>% 
+  unnest %>% 
+  select(-rowname) %>% 
+  mutate(vifv = round(vifv, 2)) %>% 
+  spread(var2, vifv) %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  column_to_rownames('var1') 
+```
+
+<div class="verysmall">
+
+```{r}
+knitr::kable(vifv, caption = 'VIF')
+```
+
+</div>
+
 # {.tabset}
 
-Pairwise models with interactions for each physiological parameter
+Pairwise models with interactions for each biomarker
 ```{r}
@@
     # model only
-    lm(frm, data = dat)
+    lm(frm, data = dat_frm)
 

### 2018-03-09 R/dat_proc.R

@@ -58,7 +58,12 @@ disdat <- read_excel('raw/parameters for WCOA 2016 combined.xlsx', sheet = 'diss
     ty3 = `type III`,
     scr = scarring
     ) %>% 
-  mutate(dis = asin(dis / 100)) %>% 
+  mutate(
+    dis = asin(dis / 100),
+    ty2 = asin(ty2),
+    ty3 = asin(ty3),
+    scr = asin(scr)
+    ) %>% 
   group_by(CTD) %>% 
   summarise(
     dis = mean(dis, na.rm = T), 

### 2018-03-09 R/funcs.R

@@ -10,3 +10,69 @@ p_ast <- function(x){
   return(out)
   
 }
+
+# vif function
+vif_func<-function(in_frame,thresh=10,trace=T,...){
+  
+  library(fmsb)
+  
+  if(any(!'data.frame' %in% class(in_frame))) in_frame<-data.frame(in_frame)
+  
+  #get initial vif value for all comparisons of variables
+  vif_init<-NULL
+  var_names <- names(in_frame)
+  for(val in var_names){
+    regressors <- var_names[-which(var_names == val)]
+    form <- paste(regressors, collapse = '+')
+    form_in <- formula(paste(val, '~', form))
+    vif_init<-rbind(vif_init, c(val, VIF(lm(form_in, data = in_frame, ...))))
+  }
+  vif_max<-max(as.numeric(vif_init[,2]), na.rm = TRUE)
+  
+  if(vif_max < thresh){
+    if(trace==T){ #print output of each iteration
+      prmatrix(vif_init,collab=c('var','vif'),rowlab=rep('',nrow(vif_init)),quote=F)
+      cat('\n')
+      cat(paste('All variables have VIF < ', thresh,', max VIF ',round(vif_max,2), sep=''),'\n\n')
+    }
+    return(var_names)
+  }
+  else{
+    
+    in_dat<-in_frame
+    
+    #backwards selection of explanatory variables, stops when all VIF values are below 'thresh'
+    while(vif_max >= thresh){
+      
+      vif_vals<-NULL
+      var_names <- names(in_dat)
+      
+      for(val in var_names){
+        regressors <- var_names[-which(var_names == val)]
+        form <- paste(regressors, collapse = '+')
+        form_in <- formula(paste(val, '~', form))
+        vif_add<-VIF(lm(form_in, data = in_dat, ...))
+        vif_vals<-rbind(vif_vals,c(val,vif_add))
+      }
+      max_row<-which(vif_vals[,2] == max(as.numeric(vif_vals[,2]), na.rm = TRUE))[1]
+      
+      vif_max<-as.numeric(vif_vals[max_row,2])
+      
+      if(vif_max<thresh) break
+      
+      if(trace==T){ #print output of each iteration
+        prmatrix(vif_vals,collab=c('var','vif'),rowlab=rep('',nrow(vif_vals)),quote=F)
+        cat('\n')
+        cat('removed: ',vif_vals[max_row,1],vif_max,'\n\n')
[…]

### 2018-03-10 expose.Rmd
added exposure analysis

---
title: "Pteropod exposure analysis"
author: ""
output: 
  html_document:
    code_folding: hide
    css: kable.css
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)

source("R/funcs.R")

data(envdat)
data(expdat)

envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
```

A simple plot:
```{r, fig.height = 4, fig.width = 7}
toplo <- expdat %>% 
  unite('temp, pCO2', temp_trt, pco2_trt, sep = '/') %>% 
  mutate(
    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
  )

ggplot(toplo, aes(x = `temp, pCO2`, y = alive)) +
  scale_y_continuous('% alive') +
  geom_boxplot(outlier.size = -1) +
  geom_jitter(alpha = 0.3, width = 0.15) + 
  theme_bw(base_family = 'serif')
```

Another simple plot:
```{r, fig.height = 4, fig.width = 7}
toplo <- expdat %>% 
  group_by(temp_trt, pco2_trt) %>% 
  summarize(
    ave = mean(alive, na.rm = T),
    lov = t.test(alive)$conf.int[1],
    hiv = t.test(alive)$conf.int[2]
  ) %>% 
  ungroup %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  )

allv <- expdat %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  ) %>%
  select(alive, temp_trt, pco2_trt, CTD) %>% 
  group_by(temp_trt, pco2_trt, CTD) %>%
  mutate(
    rep = 1, 
    rep = cumsum(rep)
    ) %>% 
  unite('uni', CTD, temp_trt, rep, remove = F)

ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
  geom_line(data = allv, aes(y = alive, group = uni), linetype = 'dotted') +
  geom_point(size = 5, aes(y = ave)) + 
  geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
  geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
  scale_y_continuous('% alive') + 
  theme_bw(base_family = 'serif')
```

Run a simple test:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt * pco2_trt, data = expdat)
summary(mod)
```

No interaction:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt + pco2_trt, data = expdat)
summary(mod)
```

Compare with environmental parameters:
```{r}
expdat <- expdat %>% 
  left_join(envdat, by = 'CTD')
```

```{r}
selectInput("lns", "Choose a variable for line thickness:", envchr)

allv <- allv %>% 
  left_join(envdat, by = 'CTD')

plo <- reactive({

  names(allv)[names(allv) %in% input$lns] <- 'thick'

  p <- ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
    geom_line(data = allv, aes(y = alive, group = uni, size = thick), alpha = 0.6) +
    geom_point(size = 5, aes(y = ave)) + 
    geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
    geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
    scale_y_continuous('% alive') + 
    theme_bw(base_family = 'serif') + 
    scale_size(input$lns, range = c(0.2, 1.5))

  return(p)

})

renderPlot({plo()}, height = 300, width = 600)

```

A model with environmental variable to model by treatment:
```{r}

selectInput("var", "Choose an environmental variable:", envchr)

# full
md <- reactive({

  frm <- formula(paste('asin(alive / 100) ~ (temp_trt + pco2_trt) * ', input$var))
  md <- lm(frm, data = allv)

  return(md)

})

# with select
md_sel <- reactive({

  frm <- formula(paste('asin(alive / 100) ~ (temp_trt + pco2_trt) * ', input$var))
  md_sel <- lm(frm, data = allv) %>% step(trace = 0)

  return(md_sel)

})

```

Full model:
```{r}
renderPrint({summary(md())})
```

Final selected model:
```{r}
renderPrint({summary(md_sel())})
```

### 2018-03-10 R/dat_proc.R

@@ -94,3 +94,19 @@ ptedat <- biodat %>%
 # save envdat, ptedat
 save(envdat, file = 'data/envdat.RData', compress = 'xz')
 save(ptedat, file = 'data/ptedat.RData', compress = 'xz')
+
+# exposure data
+expdat <- read_excel('raw/experimental treatment for Marcus.xlsx') %>% 
+  rename(
+    temp_trt = T, 
+    pco2_trt = PCO2, 
+    dead = `%Dead`, 
+    alive = `%Alive`
+  ) %>% 
+  mutate(
+    temp_trt = ifelse(temp_trt < 10, 'lo', 'hi'),
+    pco2_trt = ifelse(pco2_trt < 800, 'lo', 'hi')
+  ) %>% 
+  select(CTD, temp_trt, pco2_trt, alive)
+
+save(expdat, file = 'data/expdat.RData', compress = 'xz')

### 2018-03-10 biomark.Rmd
remove_rownames fix for shiny server

@@
   data.frame %>% 
+  remove_rownames %>% 
   column_to_rownames('CTD')
@@
   data.frame(stringsAsFactors = F) %>% 
+  remove_rownames %>% 
   column_to_rownames('var1')
@@
   data.frame(stringsAsFactors = F) %>% 
+  remove_rownames %>% 
   column_to_rownames('var1') 

### 2018-03-10 physio.Rmd
remove_rownames fix for shiny server

@@
   data.frame(stringsAsFactors = F) %>%
+  remove_rownames %>% 
   column_to_rownames('var1') 

### 2018-03-12 expose.Rmd
remove_rownames fix for shiny server
working on better surviror comparisons

@@
 ```{r}
-expdat <- expdat %>% 
-  left_join(envdat, by = 'CTD')
+
+toplo2 <- expdat %>% 
+  left_join(envdat, by = 'CTD') %>% 
+  unite(`temp, pCO2`, temp_trt, pco2_trt, sep = '/') %>% 
+  mutate(
+    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
+  )
+
+toplo3 <- toplo2 %>% 
+  gather('var', 'val', -CTD, -`temp, pCO2`) %>% 
+  group_by(CTD, var) %>% 
+  summarize(
+    val = mean(val, na.rm = T)
+    ) %>% 
+  spread(var, val)
+
+selectInput("vr", "Choose a variable to plot again survivorship:", envchr)
+
+plovr <- reactive({
+
+  names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'
+
+  p <- ggplot(toplo2, aes(x = xval, y = alive)) + 
+    geom_point(size = 5, aes(colour = `temp, pCO2`)) + 
+    scale_y_continuous('% alive') + 
+    theme_bw(base_family = 'serif') + 
+    scale_x_continuous(input$vr) + 
+    stat_smooth(method = 'lm') +
+    ggtitle('all observations')
+
+  return(p)
+
+})
+
+ploav <- reactive({
+
+  names(toplo3)[names(toplo3) %in% input$vr] <- 'xval'
+
+  p <- ggplot(toplo3, aes(x = xval, y = alive)) + 
+    geom_point(size = 5) + 
+    scale_y_continuous('% alive') + 
+    theme_bw(base_family = 'serif') + 
+    scale_x_continuous(input$vr) + 
+    stat_smooth(method = 'lm') + 
+    ggtitle('Averaged by site')
+
+  return(p)
+
+})
+
+# ranout <- reactive(lme(alive)
+
+renderPlot({plovr()}, height = 300, width = 600)
+renderPlot({ploav()}, height = 300, width = 600)
 ```
@@
 
-A model with environmental variable to model by treatment:
 ```{r}
 
-selectInput("var", "Choose an environmental variable:", envchr)
+selectInput("var", "Choose an environmental variable to model:", envchr)
 

### 2018-03-12 physio.Rmd
fixed issue with na.omit in physio

@@
 
-env <- select(dat_frm, one_of(envchr)) %>% 
+env <- dat_frm %>% 
+  na.omit %>% 
+  select(., one_of(envchr)) %>% 
   decostand(method = 'range')
-pte <- select(dat_frm, one_of(ptechr)) %>% 
+pte <- dat_frm %>% 
+  na.omit %>% 
+  select(one_of(ptechr)) %>% 
   decostand(method = 'range')

### 2018-03-12 expose.Rmd
push survivor cor to server

@@
 
-# ranout <- reactive(lme(alive)
-
 renderPlot({plovr()}, height = 300, width = 600)
@@
 ```{r}
-selectInput("lns", "Choose a variable for line thickness:", envchr)
+# all correlations
+crs <- crossing(var1 = 'alive', var2 = envchr) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
 
-allv <- allv %>% 
-  left_join(envdat, by = 'CTD')
+    crs = map(data, function(x){
 
-plo <- reactive({
+      # variables
+      vr1 <- toplo3[[x$var1]]
+      vr2 <- toplo3[[x$var2]]
 
-  names(allv)[names(allv) %in% input$lns] <- 'thick'
+      # pearson
+      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
+      pr_cr <- round(pr_ts$estimate, 2)
+      pr_pv <- p_ast(pr_ts$p.value)
+      pr <- paste(pr_cr, pr_pv)
 
-  p <- ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
-    geom_line(data = allv, aes(y = alive, group = uni, size = thick), alpha = 0.6) +
-    geom_point(size = 5, aes(y = ave)) + 
-    geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
-    geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
-    scale_y_continuous('% alive') + 
-    theme_bw(base_family = 'serif') + 
-    scale_size(input$lns, range = c(0.2, 1.5))
+      # spearman
+      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
+      sp_cr <- round(sp_ts$estimate, 2)
+      sp_pv <- p_ast(sp_ts$p.value)
+      sp <- paste(sp_cr, sp_pv)
 
-  return(p)
+      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
+      return(out)
 
     })
-
-renderPlot({plo()}, height = 300, width = 600)
-
+  ) %>% 
+  unnest %>% 
+  select(-rowname) %>% 
+  gather('cortyp', 'corval', pr, sp) %>% 
+  spread(var2, corval)
 ```
@@
 ```{r}
-
-selectInput("var", "Choose an environmental variable to model:", envchr)
-
-# full
-md <- reactive({
-
-  frm <- formula(paste('asin(alive / 100) ~ (temp_trt + pco2_trt) * ', input$var))
-  md <- lm(frm, data = allv)
-
-  return(md)
-
-})
-
-# with select
-md_sel <- reactive({
-
-  frm <- formula(paste('asin(alive / 100) ~ (temp_trt + pco2_trt) * ', input$var))
-  md_sel <- lm(frm, data = allv) %>% step(trace = 0)
-
-  return(md_sel)
-
-})
-
+knitr::kable(crs, caption = 'Pearson and Spearman correlations, based on averages by site')
 ```
-
-Full model:
-```{r}
-renderPrint({summary(md())})
-```
-
-Final selected model:
-```{r}
-renderPrint({summary(md_sel())})
-```

### 2018-03-15 expose.Rmd
comps by treatment for exposure

@@
   mutate(
-    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
+    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi')),
+    alive_tr = asin(alive / 100)
   )
 
-toplo3 <- toplo2 %>% 
-  gather('var', 'val', -CTD, -`temp, pCO2`) %>% 
-  group_by(CTD, var) %>% 
-  summarize(
-    val = mean(val, na.rm = T)
-    ) %>% 
-  spread(var, val)
+selectInput("vr", "Choose a variable to plot again survivorship:", envchr)
+selectInput("trt", "Choose a treatment (temp, pCO2):", as.character(levels(toplo2$`temp, pCO2`)))
 
-selectInput("vr", "Choose a variable to plot again survivorship:", envchr)
+plovrid <- reactive({
 
-plovr <- reactive({
-
+  # assign variable selection to name
   names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'
 
-  p <- ggplot(toplo2, aes(x = xval, y = alive)) + 
-    geom_point(size = 5, aes(colour = `temp, pCO2`)) + 
+  # filter by treatment
+  filt_toplo2 <- toplo2 %>% 
+    filter(`temp, pCO2` %in% input$trt)
+
+  p <- ggplot(filt_toplo2, aes(x = xval, y = alive)) + 
+    geom_point(size = 5) + 
     scale_y_continuous('% alive') + 
+    coord_cartesian(ylim = c(0, 105)) +
     theme_bw(base_family = 'serif') + 
@@
     stat_smooth(method = 'lm') +
-    ggtitle('all observations')
+    ggtitle('identity')
 
@@
 
-ploav <- reactive({
+plovr <- reactive({
 
-  names(toplo3)[names(toplo3) %in% input$vr] <- 'xval'
+  # assign variable selection to name
+  names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'
 
-  p <- ggplot(toplo3, aes(x = xval, y = alive)) + 
+  # filter by treatment
+  filt_toplo2 <- toplo2 %>% 
+    filter(`temp, pCO2` %in% input$trt)
+
+  p <- ggplot(filt_toplo2, aes(x = xval, y = alive_tr)) + 
     geom_point(size = 5) + 
-    scale_y_continuous('% alive') + 
+    scale_y_continuous('% alive (arcsine)') + 
+    coord_cartesian(ylim = c(0, 1.4)) +
     theme_bw(base_family = 'serif') + 
@@
     stat_smooth(method = 'lm') +
-    ggtitle('Averaged by site')
+    ggtitle('transformed')
 
@@
 
+renderPlot({plovrid()}, height = 300, width = 600)
 renderPlot({plovr()}, height = 300, width = 600)
-renderPlot({ploav()}, height = 300, width = 600)
 ```
 
-```{r}
+```{r, eval = T}
 # all correlations
-crs <- crossing(var1 = 'alive', var2 = envchr) %>% 
-  rownames_to_column() %>% 
-  group_by(rowname) %>% 
+crs <- toplo2 %>% 
+  select(-CTD) %>% 
+  filter(!`temp, pCO2` %in% 'hi/hi') %>% 
+  gather('var', 'val', -`temp, pCO2`, -alive_tr, -alive) %>% 
+  gather('alvvr', 'alvvl', alive_tr, alive) %>% 
+  group_by(`temp, pCO2`, alvvr, var) %>% 
   nest %>% 
@@
       # variables
-      vr1 <- toplo3[[x$var1]]
-      vr2 <- toplo3[[x$var2]]
+      vr1 <- x$alvvl
+      vr2 <- x$val
 
@@
   ) %>% 
+  select(-data) %>% 
   unnest %>% 
-  select(-rowname) %>% 
   gather('cortyp', 'corval', pr, sp) %>% 
-  spread(var2, corval)
+  spread(var, corval) %>% 
+  rename(response = alvvr) %>% 
+  mutate(response = factor(response, levels = c('alive', 'alive_tr'), labels = c('identity', 'arcsine'))) %>% 
+  arrange(cortyp, response, `temp, pCO2`)
 ```
 
-```{r}
-knitr::kable(crs, caption = 'Pearson and Spearman correlations, based on averages by site')
+```{r, eval = T}
+knitr::kable(crs, caption = 'Pearson and Spearman correlations of survival (identity and arcsine) by variable')
 ```

### 2018-03-15 biomark.Rmd
methods added for stats analysis

@@
 ```
+
+## Methods
+
+The response of pteropods to multiple stressors at the cellular, physiological, and population levels were evaluated using correlation analysis, multiple linear regression, and multivariate analysis.
+The general approach was to identify relationships among variables as potentially additive, synergistic, and antagonistic by evaluating pairwise effects of stressors on each response measure.
+Pearson correlations of environmental parameters (OA parameters, temperature, and chlorophyll) with cellular, physiological, and population response were first evaluated to identify potential associations with individual variables.
+Correlations between cellular and physiological responses were also evaluated to identify links between the two levels of biological organization.
+Non-continuous or skewed variables were transformed to better satisfy assumptions of the parametric tests (e.g., abundances were logarithmically transformed, proportions were arcsine transformed).
+All analyses were performed with the R statistical programming language (R Core Team 2017).
+
+Linear models were developed for every pairwise combination of environmental variables to evaluate additive and interactive effects on stressor response.
+Each model included separate terms for individual variables and a third term for the interaction of the pair.
+Evidence for additive effects were based solely on the magnitude of the estimated parameter for each variable, whereas synergistic or antagonistic effects were evaluated from the estimated parameter for the interaction.
+A positive interaction was evidence of a synergistic effect and a negative interaction was evidence of an antagonistic effect.
+Results of the linear models were further evaluated using effects plots to characterize the relationship of a pteropod response measure to continuous values for one stressor given two different values for the second stressor (constant at the minimum and maximum observed values).
+Thresholds for synergistic or antagonistic effects between two variables were identified from the effects plots as the intersection between the two fitted lines.
+
+Variance Inflation Factors (VIF) were quantified for all pairwise combinations of environmental variables used in the regression analysis to evaluate collinearity, such that:
+
+VIFj = 1 /( 1 − R2j) 		(1)
+
+where VIF is the reciprocal of the unexplained variance (1 − R2) of the linear regression of variable j against all other explanatory variables (Zuur et al., 2007).
+Although there are no specific VIF thresholds beyond which a variable is considered overly collinear, values less than 50 are generally considered adequate (Zuur et al.
+2007).
+Comparisons of variables indicated that all VIF were below this threshold, except combinations of CO3/Aragonite (70.6), pCO2/CO3 (51.4), and ph/pCO2 (79.2).
+
+Finally, multivariate comparisons between the cellular response measures and environmental variables were assessed using redundancy analysis (RDA) to jointly characterize relationships between all sampling stations.
+This analysis is conceptually similar to principal components analysis with an additional constraint on the environmental matrix, where the relationships are further partitioned based on covariance among response measures at each site in addition to the covariance between environmental variables.
+The final triplot (two biplots of environmental and response matrices) can be used to evaluate which environmental variables are correlated, as well as their relationships to the cellular response measures at each site.
+The environmental and response matrices were standardized to range from 0-1 prior to RDA.
+The vegan package for R was used for standardization and RDA (Oksanen et al.
+2017).
+
+Oksanen, J., Guillaume Blanchet, F.,  Friendly, M., Kindt, R.,  Legendre, P., McGlinn, D., Minchin, P.R., O'Hara, R.B., Simpson, G.L., Solymos, P., Henry, M., Stevens, H., Szoecs, E., Wagner, H.
+2017.
+vegan: Community Ecology Package.
+R package version 2.4-5.
+https://CRAN.R-project.org/package=vegan
+
+R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
+
+Zuur, A.F., Ieno, E.N., Smith, G.M. 2007. Analysing Ecological DAta. Springer, New York, NY. 
+MOntgomery, D.C., Peck, E.A. 1992. Introduction to Linear Regression Analysis. Wiley, New York, NY.
 

### 2018-03-16 expose.Rmd
pairwise models by trt for exposure

@@
 
-toplo2 <- expdat %>% 
+tomod <- expdat %>% 
   left_join(envdat, by = 'CTD') %>% 
@@
 selectInput("vr", "Choose a variable to plot again survivorship:", envchr)
-selectInput("trt", "Choose a treatment (temp, pCO2):", as.character(levels(toplo2$`temp, pCO2`)))
+selectInput("trt", "Choose a treatment (temp, pCO2):", as.character(levels(tomod$`temp, pCO2`)))
 
@@
   # assign variable selection to name
-  names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'
+  names(tomod)[names(tomod) %in% input$vr] <- 'xval'
 
   # filter by treatment
-  filt_toplo2 <- toplo2 %>% 
+  filt_tomod <- tomod %>% 
     filter(`temp, pCO2` %in% input$trt)
 
-  p <- ggplot(filt_toplo2, aes(x = xval, y = alive)) + 
+  p <- ggplot(filt_tomod, aes(x = xval, y = alive)) + 
     geom_point(size = 5) + 
@@
   # assign variable selection to name
-  names(toplo2)[names(toplo2) %in% input$vr] <- 'xval'
+  names(tomod)[names(tomod) %in% input$vr] <- 'xval'
 
   # filter by treatment
-  filt_toplo2 <- toplo2 %>% 
+  filt_tomod <- tomod %>% 
     filter(`temp, pCO2` %in% input$trt)
 
-  p <- ggplot(filt_toplo2, aes(x = xval, y = alive_tr)) + 
+  p <- ggplot(filt_tomod, aes(x = xval, y = alive_tr)) + 
     geom_point(size = 5) + 
@@
 # all correlations
-crs <- toplo2 %>% 
+crs <- tomod %>% 
   select(-CTD) %>% 
@@
 ```
+
+# Pairwise linear models by treatment
+
+```{r}
+
+env_cmb <- envchr  %>% 
+  combn(2) %>% 
+  t %>% 
+  data.frame(stringsAsFactors = F) %>%
+  crossing(trt = levels(tomod$`temp, pCO2`), res = c('alive', 'alive_tr'), .) %>% 
+  filter(!trt %in% c("hi/hi"))
+
+# all correlations
+mod_all <- env_cmb %>% 
+  t %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  as.list %>% 
+  map(function(x){
+
+    # formula
+    frm <- paste(x[[2]], '~', x[[3]], '*', x[[4]]) %>% 
+      as.formula
+
+    # model only
+    trt <<- which(tomod$`temp, pCO2` %in% x[[1]])
+    mod <- lm(frm, tomod, subset = trt)
+
+    levs1 <- quantile(mod$model[, 2],  c(0, 1)) %>% 
+      list
+    names(levs1) <- names(mod$model)[2]
+
+    levs2 <- quantile(mod$model[, 3], c(0, 1)) %>% 
+      list
+    names(levs2) <- names(mod$model)[3]
+
+    trm <- mod %>% coefficients %>% names %>% .[4]    
+    eff1 <- plot(effect(trm, mod, xlevels = levs1), multiline = T)
+    eff2 <- plot(effect(trm, mod, xlevels = levs2), multiline = T)
+
+    out <- list(mod = mod, eff1 = eff1, eff2 = eff2)
+
+    return(out)
+
+  }) %>% 
+  enframe('Model', 'Modobj') %>% 
+  mutate(Model = gsub('^X', 'mod', Model)) %>% 
+  bind_cols(env_cmb) %>% 
+  mutate(
+    Modsum = map(Modobj, function(x){
+
+      x <- x$mod
+
+      # rsq
+      rsq <- summary(x) %>% 
+        .$r.squared
+
+      # coeff summary
+      out <- x %>% 
+        summary %>% 
+        .$coefficients %>% 
+        .[-1, c(1, 4)] %>% 
+        data.frame %>% 
+        rownames_to_column('Env') %>% 
+        mutate(
+          Res = formula(x)[2] %>% as.character,
+          Rsq = round(rsq, 2),
+          Est = round(Estimate, 2), 
+          Pvl = p_ast(Pr...t..)
+        ) %>% 
+        select(Res, Env, Rsq, Est, Pvl)
+
+    return(out)
+
+    })
+  )
+
+mods <- mod_all %>%
+  select(-Modobj) %>%
+  unnest %>% 
+  select(-res, -X1, -X2)
+
+```
+
+```{r}
+selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mod_all))))
+
+md <- reactive({
+
+  md <- filter(mod_all, Model %in% input$mod) %>% 
+  .$Modobj %>% 
+  .[[1]]
+
+  return(md)
+
+})
+
+renderPrint({summary(md()$mod)})
+
+renderPlot({
+  toplo <- md()$mod %>% 
+    .$model %>% 
+    gather('var', 'val', -1)
+
+  ggplot(toplo, aes_string(x='val', y=names(toplo)[1])) +
+    geom_point() + 
+    facet_wrap(~var, ncol = 2, scales = 'free_x') + 
+    theme_bw(base_family = 'serif') + 
+    geom_smooth(method = 'lm') + 
+    theme(axis.title.x = element_blank())
+
+}, height = 300, width = 600)
+
+prd_vl <- reactive({
+
+  md <- md()$mod
+  x <- md$model %>% 
+    .[, -1] %>% 
+    data.frame %>% 
+    as.list %>% 
+    map(range) %>%
+    map(function(x) seq(x[1], x[2], length = 100))
+
+  nms <- names(x) 
+  x <- crossing(x[[1]], x[[2]])
+  names(x) <- nms
+
+  prd_vl <- predict(md, newdata = x) %>% 
+    data.frame(., x)
+  names(prd_vl)[1] <- all.vars(formula(md))[1]
+
+  return(prd_vl)
+
+})
+
+renderPlot({
+  md()$eff1
+}, height = 300, width = 600)
+
+renderPlot({
+  md()$eff2
+}, height = 300, width = 600)
+
+renderPlot({
+
+  pbase <- theme(
+    panel.grid.major = element_blank(), 
+    panel.grid.minor = element_blank(), 
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
+    axis.text.y = element_text(size = 10),
+    legend.position = 'top', 
+    plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+    strip.background = element_blank(), 
+    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
+  )
+
+  p <- ggplot(prd_vl(), aes_string(y = names(prd_vl())[2], x = names(prd_vl())[3], fill = names(prd_vl())[1])) + 
+    geom_tile(colour = 'black') + 
+    theme_bw(base_family = 'serif') +
+    pbase +
+    scale_y_continuous(expand = c(0, 0)) +
+    scale_x_continuous(expand = c(0, 0)) +
+    scale_fill_distiller(palette = 'Spectral') +
+    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 15, label.theme = element_text(size = 12, angle = 0)))
+  p
+}, height = 700, width = 700)
+
+renderPlotly({
+
+  dat <- prd_vl() %>% 
+    spread_(key_col = names(.)[2], value_col = names(.)[1]) %>% 
+    .[,-1] %>% 
+    as.matrix
+
+  scene <- list(
+    aspectmode = 'manual', 
+    aspectratio = list(x = 1, y = 1, z = 0.5), 
+    xaxis = list(title = names(prd_vl()[3])),
+    yaxis = list(title = names(prd_vl()[2])),
+    zaxis = list(title = names(prd_vl()[1]))
+    )
+
+  p <- plot_ly(z = ~dat) %>% add_surface(colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
+    layout(scene = scene)
+  p
+
+})
+```
+
+# {.tabset}
+
+## Temp hi, pCO2 lo
+
+Identity response:
+
+```{r}
+knitr::kable(filter(mods, mods$trt %in% 'hi/lo' & mods$Res %in% 'alive'))
+```
+
+Arcsine response:
+
+```{r}
+knitr::kable(filter(mods, mods$trt %in% 'hi/lo' & mods$Res %in% 'alive_tr'))
+```
+
+## Temp lo, pCO2 hi
+
+Identity response:
+
+```{r}
+knitr::kable(filter(mods, mods$trt %in% 'lo/hi' & mods$Res %in% 'alive'))
+```
+
+Arcsine response:
+
+```{r}
+knitr::kable(filter(mods, mods$trt %in% 'lo/hi' & mods$Res %in% 'alive_tr'))
+```
+
+## Temp lo, pCO2 lo
+
+Identity response:
+
+```{r}
+knitr::kable(filter(mods, mods$trt %in% 'lo/lo' & mods$Res %in% 'alive'))
+```
+
+Arcsine response:
+
+```{r}
+knitr::kable(filter(mods, mods$trt %in% 'lo/lo' & mods$Res %in% 'alive_tr'))
+```

### 2018-03-16 expose.Rmd
better model sel input for expose

@@
 ```{r}
-selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mod_all))))
+sellab <- env_cmb %>% 
+  bind_cols(md = paste0('mod', seq(1:nrow(mod_all))), .) %>% 
+  unite('vr', trt, res, X1, X2, sep = ', ') %>% 
+  unite('vr2', md, vr, sep = ': ') %>% 
+  unlist %>% 
+  as.character
+
+selectInput("mod", "Choose model:", sellab)
 
@@
 
-  md <- filter(mod_all, Model %in% input$mod) %>% 
+  tosel <- strsplit(input$mod, ':')[[1]][1]
+
+  md <- filter(mod_all, Model %in% tosel) %>% 
   .$Modobj %>% 

### 2018-03-17 R/dat_proc.R
exposure analysis averaged by ctd, trt

@@ -107,6 +107,9 @@ expdat <- read_excel('raw/experimental treatment for Marcus.xlsx') %>%
     temp_trt = ifelse(temp_trt < 10, 'lo', 'hi'),
     pco2_trt = ifelse(pco2_trt < 800, 'lo', 'hi')
   ) %>% 
-  select(CTD, temp_trt, pco2_trt, alive)
+  select(CTD, temp_trt, pco2_trt, alive) %>% 
+  group_by(CTD, temp_trt, pco2_trt) %>% 
+  summarize(alive = mean(alive)) %>% 
+  ungroup
 
 save(expdat, file = 'data/expdat.RData', compress = 'xz')

### 2018-03-20 figures.Rmd
started figures

---
output:
  html_document:
    keep_md: yes
    self_contained: no
    code_folding: hide
---

# Pteropod stressor interaction: figures

```{r echo = T, message = F, warning = F}
library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(knitr)
library(gridExtra)

source("R/funcs.R")

opts_chunk$set(fig.align = 'center', message = F, echo = T, cache = F, dev = 'png', dev.args = list(family = 'serif'), dpi = 400, fig.pos = '!h', warning = F, background = 'white', out.width = '100%',
               fig.process = function(x) {
                 x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                 if (file.rename(x, x2)) x2 else x
               })

data(envdat)
data(ptedat)

envchr <- c('Lat', 'pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
phychr <- c('abu', 'dis', 'len', 'ty2', 'ty3', 'scr')
```

```{r rda_bio}
# biomarker rda model
dat_bio <- ptedat %>% 
  select(one_of('CTD', biochr)) %>% 
  na.omit %>% 
  inner_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, biochr))) %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

envphy <- select(dat_bio, one_of(envchr)) %>% 
  select(-Lat) %>% 
  decostand(method = 'range')
ptephy <- select(dat_bio, one_of(biochr)) %>% 
  decostand(method = 'range')

biomod <- rda(ptephy, envphy)

# physiology rda model
dat_phy <- ptedat %>% 
  select(one_of('CTD', phychr)) %>% 
  na.omit %>% 
  inner_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, phychr))) %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

envphy <- select(dat_phy, one_of(envchr)) %>% 
  select(-Lat) %>% 
  decostand(method = 'range')
ptephy <- select(dat_phy, one_of(phychr)) %>% 
  decostand(method = 'range')

phymod <- rda(ptephy, envphy)

```

RDA plots:
```{r rda_plt, fig.height = 4.5, fig.width = 9}
# globals
coord_fix <- F
repel <- F
addsize <- 3
alpha <- 0.8
arrow <- 0.3
cols <- 'lightblue'

p1a <- ggord(biomod, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_bio$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
  ggtitle('(a) Cellular stressors')
p2a <- ggord(phymod, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_phy$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
  ggtitle('(b) Physiological stressors')

grid.arrange(p1a, p2a, ncol = 2)
```

### 2018-03-20 R/dat_proc.R

@@ -3,9 +3,10 @@ library(tidyverse)
 
 # environmental data
 envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
-  select(`CTD Station`, `'pCO2'`, `'pH'`, `'CO3'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
+  select(`CTD Station`, `Latitude`, `'pCO2'`, `'pH'`, `'CO3'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
   rename(
-    CTD = `CTD Station`, 
+    CTD = `CTD Station`,
+    Lat = `Latitude`,
     Ara = `'Aragonite'`, 
     O2 = `'Oxygen'`,
     pH = `'pH'`,

### 2018-03-21 R/dat_proc.R
full abu data added

@@ -35,7 +35,7 @@ biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx',
   ungroup
 
 # abundance
-abudat <- read_excel('raw/Copy of abundances to be usef_Marcus2.xlsx') %>% 
+abudat <- read_excel('raw/abundances to be usef_Marcus.xlsx') %>% 
   rename(
     CTD = `CTD station`,
     abund = `real abundance`

### 2018-03-21 figures.Rmd
correlation heatmap fig added

@@
 library(gridExtra)
+library(grid)
 
@@
 ```
+
+Heatmaps of correlations:
+
+```{r}
+# physiology rda model
+dat_cor <- ptedat %>% 
+  select(one_of('CTD', phychr, biochr)) %>% 
+  # na.omit %>% 
+  left_join(envdat, by = 'CTD') %>% 
+  data.frame %>% 
+  remove_rownames %>% 
+  column_to_rownames('CTD')
+
+# all correlations
+crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>% 
+  filter(var1 != var2) %>% 
+  rownames_to_column() %>% 
+  group_by(rowname) %>% 
+  nest %>% 
+  mutate(
+    crs = map(data, function(x){
+
+      # variables
+      vr1 <- dat_cor[[x$var1]]
+      vr2 <- dat_cor[[x$var2]]
+
+      # pearson
+      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
+      pr_cr <- round(pr_ts$estimate, 2)
+      pr_pv <- p_ast(pr_ts$p.value)
+      pr <- paste(pr_cr, pr_pv)
+
+      out <- data.frame(pr = pr, stringsAsFactors = F)
+      return(out)
+
+    })
+  ) %>% 
+  unnest %>% 
+  select(-rowname)
+```
+
+```{r corplo, fig.height = 6, fig.width = 7}
+levs <- c(sort(envchr), sort(biochr), sort(phychr))
+prplo <- crs %>% 
+  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
+  filter(var1 %in% levs & var2 %in% levs) %>% 
+  mutate(
+    cor = as.numeric(cor), 
+    var1 = factor(var1, levels = rev(levs)), 
+    var2 = factor(var2, levels = rev(levs)), 
+    sig = gsub('ns', '', sig)
+  )
+
+pbase <- theme(
+  panel.grid.major = element_blank(), 
+  panel.grid.minor = element_blank(), 
+  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
+  axis.text.y = element_text(size = 8),
+  legend.position = 'top', 
+  plot.margin = unit(c(0,4,0,0), "lines"),
+  strip.background = element_blank(), 
+  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
+  panel.background = element_rect(fill = 'black')
+  )
+
+outlab <- data.frame(
+  y = c(3.5, 10.5, 18.5), 
+  lab = c('Physiology', 'Cellular', 'Environment')
+)
+
+p <- ggplot(prplo) + 
+  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') + 
+  geom_text(aes(y = var1, x = var2, label = sig)) +
+  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 23, xmax = 23) +
+  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 23, xmax = 23) +  
+  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 23, xmax = 23) +
+  pbase +
+  scale_y_discrete('', expand = c(0, 0)) + 
+  scale_x_discrete('', expand = c(0, 0)) +
+  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
+  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0))) +
+  geom_hline(yintercept = 6.5, size = 1.5) +
+  geom_hline(yintercept = 14.5, size = 1.5) +
+  geom_vline(xintercept = 6.5, size = 1.5) +
+  geom_vline(xintercept = 14.5, size = 1.5)
+
+# Code to override clipping
+gt <- ggplot_gtable(ggplot_build(p))
+gt$layout$clip[gt$layout$name == "panel"] <- "off"
+grid.draw(gt)
+```

### 2018-03-23 figures.Rmd
effects plots added

@@
 
-RDA plots:
 ```{r rda_plt, fig.height = 4.5, fig.width = 9}
@@
 p1a <- ggord(biomod, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_bio$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
-  ggtitle('(a) Cellular stressors')
+  ggtitle('(a) Cellular endpoints')
 p2a <- ggord(phymod, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_phy$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
-  ggtitle('(b) Physiological stressors')
+  ggtitle('(b) Physiological endpoints')
 
@@
 
-Heatmaps of correlations:
+Fig.
+1 Results of redundancy analyses for environmental variables with (a) cellular and (b) physiological endpoints of pteropod response to OA stressors.
+Points are site locations in multivariate space with the size proportional to latitude.
+Separate RDAs were created for cellular and physiological endpoints because not all data were available across all stations.
 
@@
   y = c(3.5, 10.5, 18.5), 
-  lab = c('Physiology', 'Cellular', 'Environment')
+  lab = c('Population/\nPhysiology', 'Cellular', 'Environment')
 )
@@
 ```
+Fig.
+2 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
+Darker red values are strong positive correlations and darker purple values are strong negative correlations.
+Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*).
+
+```{r effplo, fig.height = 4, fig.width = 9}
+# abu, no syngeristics
+data(modex1)
+
+# lpx, synergistic
+data(modex2)
+
+x <- modex2$model %>% 
+  .[, -1] %>% 
+  data.frame %>% 
+  as.list %>% 
+  map(range) %>%
+  map(function(x) seq(x[1], x[2], length = 100))
+
+x$pH <- modex2$model$pH %>% quantile
+
+nms <- names(x) 
+x <- crossing(x[[1]], x[[2]])
+names(x) <- nms
+
+prd_vl <- predict(modex2, newdata = x) %>% 
+  data.frame(., x)
+names(prd_vl)[1] <- all.vars(formula(modex2))[1]
+
+toplo <- prd_vl
+
+yval <- toplo %>% group_by(pH) %>% summarize(LPX = min(LPX))
+xlab <- data.frame(
+  lab = rev(c('Max', '75th', 'Med', '25th', 'Min')), 
+  x = 1.23, y = yval$LPX,
+  stringsAsFactors = F)
+xlab[1, 3] <- 4.2
+
+p1 <- ggplot() +
+  geom_line(data = toplo, aes(x = Ara, y = LPX, group = pH, colour = pH), size = 1) + 
+  geom_text(data = xlab, aes(x= x, y = y, label = lab), hjust = 1) +
+  theme_bw() +
+  coord_cartesian(xlim = c(1.2, 1.75)) +
+  xlab(expression(Omega[AR])) +
+  scale_colour_gradientn(colours = RColorBrewer::brewer.pal(9, 'RdBu')) +
+  ggtitle("(a) Mitigating effect of pH on low\n aragonite saturation, synergistic")
+
+# abu additive
+data(modex1)
+
+x <- modex1$model %>% 
+  .[, -1] %>% 
+  data.frame %>% 
+  as.list %>% 
+  map(range) %>%
+  map(function(x) seq(x[1], x[2], length = 100))
+
+x$O2 <- modex1$model$O2 %>% quantile
+
+nms <- names(x) 
+x <- crossing(x[[1]], x[[2]])
+names(x) <- nms
+
+prd_vl <- predict(modex1, newdata = x, type = 'response') %>% 
+  data.frame(., x)
+names(prd_vl)[1] <- all.vars(formula(modex1))[1]
+
+toplo <- prd_vl
+
+yval <- toplo %>% group_by(O2) %>% summarize(abu = min(abu))
+xlab <- data.frame(
+  lab = rev(c('Max', '75th', 'Med', '25th', 'Min')), 
+  x = 390, y = yval$abu,
+  stringsAsFactors = F)
+
+p2 <- ggplot() +
+  geom_line(data = toplo, aes(x = pCO2, y = abu, group = O2, colour = O2), size = 1) + 
+  geom_text(data = xlab, aes(x= x, y = y, label = lab), hjust = 1) +
+  theme_bw() +
+  coord_cartesian(xlim = c(360, 860)) +
+  scale_colour_gradientn(colours = RColorBrewer::brewer.pal(9, 'RdBu')) + 
+  scale_y_continuous('Pteropod abundance') +
+  ggtitle("(b) Mitigating effect of oxygen on\nOA stress, additive")
+
+grid.arrange(p1, p2, ncol = 2)
+
+```
+Fig.
+3 Examples of model interactions of co-occuring environmental variables on pteropod response measures.
+Plot (a) shows a synergistic effect between pH and aragonite saturation on LPX expression.
+Plot (b) shows an additive response of oxyxgen and OA stress (pCO2) on pteropod abundance.
+Both y-axes are transformed to conform to model output.
+Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.

### 2018-03-23 physio.Rmd
effects plots added

@@
 
+# # back-transform abu for models
+# dat_frm <- dat_frm %>% 
+#   mutate(
+#     abu = (10^abu) - 1,
+#     abu = as.integer(round(abu, 0))
+#   )
+#   
 # models
@@
 
+    # if(x[1] == 'abu'){
+    #   
+    #   # zinf model
+    #   out <- zeroinfl(frm, data = dat_frm)
+    # 
+    # } else {
+
       # model only
-    lm(frm, data = dat_frm)
+      out <- lm(frm, data = dat_frm)
+
+    # }
+
+    return(out)
 

### 2018-03-23 biomark.Rmd
stressor mods created in dat_proc

@@
 data(ptedat)
+data(biomod)
 
@@
 Pairwise models with interactions for each biomarker
-```{r}
-# variables to model, chr strings only
-env_cmb <- names(env) %>% 
-  combn(2) %>% 
-  t %>% 
-  data.frame(stringsAsFactors = F) %>%
-  crossing(names(pte), .) %>% 
-  t %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  as.list
-
-# models
-mod_all <- env_cmb %>%
-  map(function(x){
-
-    # formula
-    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
-      as.formula
-
-    # model only
-    lm(frm, data = dat_frm)
-
-  }) %>% 
-  enframe('Model', 'Modobj') %>% 
-  mutate(
-    Model = gsub('^X', 'mod', Model), 
-    data = map(Modobj, function(x){
-
-      # response
-      yvar <- all.vars(formula(x))[1]
-
-      # rsq
-      rsq <- summary(x) %>% 
-        .$r.squared
-
-      # coeff summary
-      out <- x %>% 
-        summary %>% 
-        .$coefficients %>% 
-        .[-1, c(1, 4)] %>% 
-        data.frame %>% 
-        rownames_to_column('env_lab') %>% 
-        mutate(
-          pte_lab = yvar,
-          Rsq = round(rsq, 2),
-          Est = round(Estimate, 2), 
-          Pvl = p_ast(Pr...t..)
-        ) %>% 
-        select(pte_lab, env_lab, Rsq, Est, Pvl)
-
-    return(out)
-
-    })
-  ) 
-mods <- mod_all %>% 
-  unnest(data)
-
-```
 
 ```{r}
+# model data
+mods <- biomod %>% 
+  unnest(data)
+mod_all <- biomod
+
 selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))

### 2018-03-23 physio.Rmd
stressor mods created in dat_proc

--- 2018-03-12 physio.Rmd
@@
 data(ptedat)
+data(phymod)
 
@@
 Pairwise models with interactions for each biomarker
-```{r}
-# variables to model, chr strings only
-env_cmb <- names(env) %>% 
-  combn(2) %>% 
-  t %>% 
-  data.frame(stringsAsFactors = F) %>%
-  crossing(names(pte), .) %>% 
-  t %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  as.list
-
-# models
-mod_all <- env_cmb %>%
-  map(function(x){
-
-    # formula
-    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
-      as.formula
-
-    # model only
-    lm(frm, data = dat_frm)
-
-  }) %>% 
-  enframe('Model', 'Modobj') %>% 
-  mutate(
-    Model = gsub('^X', 'mod', Model), 
-    data = map(Modobj, function(x){
-
-      # response
-      yvar <- all.vars(formula(x))[1]
-
-      # rsq
-      rsq <- summary(x) %>% 
-        .$r.squared
-
-      # coeff summary
-      out <- x %>% 
-        summary %>% 
-        .$coefficients %>% 
-        .[-1, c(1, 4)] %>% 
-        data.frame %>% 
-        rownames_to_column('env_lab') %>% 
-        mutate(
-          pte_lab = yvar,
-          Rsq = round(rsq, 2),
-          Est = round(Estimate, 2), 
-          Pvl = p_ast(Pr...t..)
-        ) %>% 
-        select(pte_lab, env_lab, Rsq, Est, Pvl)
-
-    return(out)
-
-    })
-  ) 
-mods <- mod_all %>% 
-  unnest(data)
-
-```
 
 ```{r}
+# model data
+mods <- phymod %>% 
+  unnest(data)
+mod_all <- phymod
+
 selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))

### 2018-03-23 R/dat_proc.R

@@ -1,6 +1,8 @@
 library(readxl)
 library(tidyverse)
 
+source('R/funcs.R')
+
 # environmental data
 envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
   select(`CTD Station`, `Latitude`, `'pCO2'`, `'pH'`, `'CO3'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
@@ -114,3 +116,177 @@ expdat <- read_excel('raw/experimental treatment for Marcus.xlsx') %>%
   ungroup
 
 save(expdat, file = 'data/expdat.RData', compress = 'xz')
+
+######
+# all stressor covar models
+
+##
+# cellular
+
+# data
+data(envdat)
+data(ptedat)
+
+envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
+ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
+
+dat_frm <- ptedat %>% 
+  select(one_of('CTD', ptechr)) %>% 
+  na.omit %>% 
+  inner_join(envdat, by = 'CTD') %>% 
+  select(one_of(c('CTD', envchr, ptechr))) %>% 
+  data.frame %>% 
+  remove_rownames %>% 
+  column_to_rownames('CTD')
+
+# models
+env_cmb <- envchr %>% 
+  combn(2) %>% 
+  t %>% 
+  data.frame(stringsAsFactors = F) %>%
+  crossing(ptechr, .) %>% 
+  t %>% 
+  data.frame(stringsAsFactors = F) %>% 
+  as.list
+
+# models
+mod_all <- env_cmb %>%
+  map(function(x, dat = dat_frm){
+    
+    # formula
+    frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
+      as.formula
+    
+    # model only
+    lm(frm, data = dat)
+    
+  }) %>% 
+  enframe('Model', 'Modobj') %>% 
+  mutate(
+    Model = gsub('^X', 'mod', Model), 
+    data = map(Modobj, function(x){
+      
+      # response
+      yvar <- all.vars(formula(x))[1]
+      
+      # rsq
+      rsq <- summary(x) %>% 
+        .$r.squared
+      
[…]

### 2018-03-23 figures.Rmd
stressor mods created in dat_proc
all effects plots done

--- 2018-03-21 figures.Rmd
@@
 data(ptedat)
+data(phymod)
+data(biomod)
 
@@
 
-biomod <- rda(ptephy, envphy)
+biorda<- rda(ptephy, envphy)
 
@@
 
-phymod <- rda(ptephy, envphy)
+phyrda <- rda(ptephy, envphy)
 
@@
 
-RDA plots:
-```{r rda_plt, fig.height = 4.5, fig.width = 9}
+```{r rdaplo, fig.height = 4.5, fig.width = 9}
 # globals
@@
 
-p1a <- ggord(biomod, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_bio$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
-  ggtitle('(a) Cellular stressors')
-p2a <- ggord(phymod, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_phy$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
-  ggtitle('(b) Physiological stressors')
+p1a <- ggord(biorda, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_bio$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
+  ggtitle('(a) Cellular endpoints')
+p2a <- ggord(phyrda, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_phy$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
+  ggtitle('(b) Physiological endpoints')
 
@@
 ```
+Fig.
+2 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
+Darker red values are strong positive correlations and darker purple values are strong negative correlations.
+Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*).
+
+```{r effbio, fig.height = 8, fig.width = 9}
+# cellular 92 (lpx, ara, ph - syn), 101 (lpx, ara, t - neg add), 164 (sod, ara, t, neg but no thrsh) 
+biosel <- biomod %>% 
+  filter(Model %in% paste0('mod', c(92, 101, 164))) %>% 
+  dplyr::select(-data)
+
+cols <- RColorBrewer::brewer.pal(9, 'RdBu')
+
+##
+# plots
+
+# mod 92 cell (lpx, ara, ph - syn)
+pl1 <- filter(biosel, grepl('92$', Model)) %>% 
+  .$Modobj %>% 
+  .[[1]] %>% 
+  get_pldat(., 'pH')
+
+p1 <- ggplot() +
+  geom_line(data = pl1[[1]], aes(x = Ara, y = LPX, group = pH, colour = pH), size = 1) + 
+  geom_text(data = pl1[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
+  theme_bw() +
+  scale_x_continuous(expression(Omega[Ar])) +
+  scale_colour_gradientn(colours = cols) + 
+  scale_y_continuous('LPX') +
+  ggtitle("(a) Synergistic effect of pH on aragonite\nsaturation")
+
+# mod 101 cell (lpx, ara, t - neg add), 164 (sod, ara, t, neg but no thrsh) 
+pl2 <- filter(biosel, grepl('101$', Model)) %>% 
+  .$Modobj %>% 
+  .[[1]] %>% 
+  get_pldat(., 'Temp')
+
+p2 <- ggplot() +
+  geom_line(data = pl2[[1]], aes(x = Ara, y = LPX, group = Temp, colour = Temp), size = 1) + 
+  geom_text(data = pl2[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
+  theme_bw() +
+  scale_x_continuous(expression(Omega[Ar])) +
+  scale_colour_gradientn(colours = cols) + 
+  scale_y_continuous('LPX') +
+  ggtitle("(b) Negative additive effect below threshold of\n temperature on aragonite saturation")
+
+# mod 164 cell (sod, ara, t, neg but no thrsh) 
+pl3 <- filter(biosel, grepl('164$', Model)) %>% 
+  .$Modobj %>% 
+  .[[1]] %>% 
+  get_pldat(., 'Temp')
+
+p3 <- ggplot() +
+  geom_line(data = pl3[[1]], aes(x = Ara, y = SOD, group = Temp, colour = Temp), size = 1) + 
+  geom_text(data = pl3[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
+  theme_bw() +
+  scale_x_continuous(expression(Omega[Ar])) +
+  scale_colour_gradientn(colours = cols) + 
+  scale_y_continuous('SOD') +
+  ggtitle("(c) Negative additive effect of temperature on \naragonite saturation")
+
+grid.arrange(p1, p2, p3, ncol = 2)
+
+```
+Fig.
+3 Examples of model interactions of co-occuring environmental variables on cellular response measures.
+Each subplot shows a different relationship as either additive or synergistic effects between the variables.
+All y-axes are transformed to conform to model output.
+Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.
+
+```{r effphy, fig.height = 4, fig.width = 9}
+
+# physio 4 (abu, o2, pco2 - pos additive), 38 (dis, ara, t - syn)
+physel <- phymod %>% 
+  filter(Model %in% paste0('mod', c(4, 38))) %>% 
+  dplyr::select(-data)
+
+cols <- RColorBrewer::brewer.pal(9, 'RdBu')
+
+# mod 4 physio (abu, o2, pco2 - pos additive)
+pl4 <- filter(physel, grepl('4$', Model)) %>% 
+  .$Modobj %>% 
+  .[[1]] %>% 
+  get_pldat(., 'O2', fct = 0.85)
+
+p4 <- ggplot() +
+  geom_line(data = pl4[[1]], aes(x = pCO2, y = abu, group = O2, colour = O2), size = 1) + 
+  geom_text(data = pl4[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
+  theme_bw() +
+  scale_colour_gradientn(colours = cols) + 
+  scale_y_continuous('Abundance') +
+  ggtitle("(a) Positive additive effect of O2 on\npCO2 stress")
+
+# mod 38 physio (dis, ara, t - syn)
+pl5 <- filter(physel, grepl('38$', Model)) %>% 
+  .$Modobj %>% 
+  .[[1]] %>% 
+  get_pldat(., 'Temp', 'right', fct = 1.05)
+
+p5 <- ggplot() +
+  geom_line(data = pl5[[1]], aes(x = Ara, y = dis, group = Temp, colour = Temp), size = 1) + 
+  geom_text(data = pl5[[2]], aes(x= x, y = y, label = lab), hjust = 1) +
+  theme_bw() +
+  scale_colour_gradientn(colours = cols) + 
+  scale_y_continuous('Shell dissolution') +
+  scale_x_continuous(expression(Omega[Ar])) +
+  ggtitle("(b) Synergistic effect of temperature on \naragonite saturation")
+
+grid.arrange(p4, p5, ncol = 2)
+```
+Fig.
+4 Examples of model interactions of co-occuring environmental variables on abundance and shell dissolution.
+Each subplot shows a different relationship as either additive or synergistic effects between the variables.
+All y-axes are transformed to conform to model output.
+Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.

### 2018-03-23 R/funcs.R

@@ -76,3 +76,64 @@ vif_func<-function(in_frame,thresh=10,trace=T,...){
   }
   
 }
+
+# Data to plot for effects
+#
+# modin lm model
+# cvar chr string of variable to hold constant
+# pos is where the labels are, left or right of effects line
+# fct is scaling factor for labels from end of lines
+get_pldat <- function(modin, cvar, pos = c('left', 'right'), fct = NULL){
+  
+  pos <- match.arg(pos)
+  
+  # crossing of model data by range
+  x <- modin$model %>% 
+    .[, -1] %>% 
+    data.frame %>% 
+    as.list %>% 
+    map(range) %>%
+    map(function(x) seq(x[1], x[2], length = 100))
+  
+  # quantiles for cvar
+  x[[cvar]] <- modin$model[[cvar]]%>% quantile
+  
+  # make data frame
+  nms <- names(x) 
+  x <- crossing(x[[1]], x[[2]])
+  names(x) <- nms
+  
+  # get predictions, combine with exp vars
+  prd_vl <- predict(modin, newdata = x) %>% 
+    data.frame(., x)
+  names(prd_vl)[1] <- all.vars(formula(modin))[1]
+  
+  # min x axis values for quantile labels
+  yvar <- names(prd_vl)[1]
+  xvar <- all.vars(formula(modin))
+  xvar <- xvar[!xvar %in% c(yvar, cvar)]
+
+  locs <- prd_vl %>% 
+    group_by(.dots = list(cvar))   
+  if(pos == 'right'){
+    if(is.null(fct)) fct <- 1.05
+    locs <- filter(locs, row_number() == n())
+  } else {
+    if(is.null(fct)) fct <- 0.95
+    locs <- filter(locs, row_number() == 1)
+  }
+  
+  yval <- locs[[yvar]]
+  xval <- locs[[xvar]] %>% unique %>% `*`(fct)
+  xlab <- data.frame(
+    lab = c('Max', '75th', 'Med', '25th', 'Min'), 
+    x = xval, y = yval,
+    stringsAsFactors = F)
+  dr <- locs[[cvar]] %>% range %>% diff %>% sign
+  if(dr == 1) xlab$lab <- rev(xlab$lab)
+  
+  # output
+  out <- list(prd_vl = prd_vl, xlab = xlab)
+  return(out)
+  
+}

### 2018-03-24 figures.Rmd
biomark table done

@@
   scale_y_continuous('Abundance') +
-  ggtitle("(a) Positive additive effect of O2 on\npCO2 stress")
+  ggtitle("(a) Synergistic effect of O2 on\npCO2 stress")
 
@@
 Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.
+
+```{r}
+mods <- readxl::read_excel('raw/info for table.xlsx') %>% 
+  .[-1, -1] %>% 
+  gather('var', 'Model', everything()) %>% 
+  mutate(
+    org = ifelse(var %in% c('abundance', 'dissolution', 'length'), 'phy', 'bio'),
+    Model = paste0('mod', Model)
+    ) %>% 
+  split(.$org)
+
+biotab <- mods$bio %>% 
+  left_join(biomod, by = 'Model') %>% 
+  select(Model, data) %>%
+  mutate(data = map(data, function(x){
+    x %>% mutate(vr = seq(1:nrow(.)))
+  })) %>% 
+  unnest %>% 
+  mutate(
+    Model = gsub('mod', '', Model),
+    Model = as.numeric(Model)
+    ) %>% 
+  group_by(Model) %>% 
+  mutate(
+    Pvl = gsub('ns', '', Pvl), 
+    Rsq = ifelse(duplicated(Rsq), '', Rsq)
+    ) %>% 
+  ungroup %>% 
+  unite('Est', Est, Pvl, sep = '') %>% 
+  gather('estnm', 'estvl', Rsq, Est) %>% 
+  unite('est', pte_lab, estnm, sep = ', ', remove = F) %>% 
+  select(-estnm) %>% 
+  split(.$pte_lab) %>%
+  lapply(., function(x){
+
+    out <- spread(x, est, estvl) %>% 
+      select(-pte_lab) %>%
+      arrange(Model, vr) %>% 
+      select(-vr, -Model)
+
+    if(!any(grepl('^LPX', names(out))))
+      out <- out %>% 
+        select(-env_lab)
+
+    return(out)
+
+  }) %>% 
+  do.call('cbind', .) %>% 
+  mutate(
+    Model = rep(letters[1:(nrow(.)/3)], each = 3),
+    Model = paste0('(', Model, ')'),
+    Model = ifelse(duplicated(Model), '', Model)
+  ) %>% 
+  select(Model, everything())
+
+names(biotab) <- gsub('^.*\\.', '', names(biotab))
+vrs <- gsub('(^.*),\\s.*$', '\\1', names(biotab))
+vrs[duplicated(vrs)] <- ''
+vrs <- paste0(vrs, '<br>')
+vrs[vrs %in% 'env_lab<br>'] <- 'Parameter<br>'
+vrs2 <- gsub('^.*,\\s|Model|env_lab', '', names(biotab))
+vrs <- paste(vrs, vrs2, sep = '')
+
+names(biotab) <- vrs
+
+knitr::kable(biotab)
+```

### 2018-03-24 figures.Rmd
all table sdone

@@
 
-knitr::kable(biotab)
+knitr::kable(biotab, caption = "Table 1: Model results for pteropod cellular response to pairs of co-occurring environmental variables.
+The estimated joint effects of variables in each model are shown.
+The overall R-squred value for each model is also shown.
+Significance of each effect is noted as * p < 0.05, ** p < 0.005.")
 ```
+
+```{r}
+phytab <- mods$phy %>% 
+  left_join(phymod, by = 'Model') %>% 
+  select(Model, data) %>%
+  mutate(data = map(data, function(x){
+    x %>% mutate(vr = seq(1:nrow(.)))
+  })) %>% 
+  unnest %>% 
+  mutate(
+    Model = gsub('mod', '', Model),
+    Model = as.numeric(Model)
+  ) %>% 
+  group_by(Model) %>% 
+  mutate(
+    Pvl = gsub('ns', '', Pvl), 
+    Rsq = ifelse(duplicated(Rsq), '', Rsq)
+  ) %>% 
+  ungroup %>% 
+  unite('Est', Est, Pvl, sep = '') %>% 
+  gather('estnm', 'estvl', Rsq, Est) %>% 
+  unite('est', pte_lab, estnm, sep = ', ', remove = F) %>% 
+  select(-estnm) %>% 
+  split(.$pte_lab) %>%
+  lapply(., function(x){
+
+    out <- spread(x, est, estvl) %>% 
+      select(-pte_lab) %>%
+      arrange(Model, vr) %>% 
+      select(-vr, -Model)
+
+    if(!any(grepl('^abu', names(out))))
+      out <- out %>% 
+        select(-env_lab)
+
+    return(out)
+
+  }) %>% 
+  do.call('cbind', .) %>% 
+  mutate(
+    Model = rep(letters[1:(nrow(.)/3)], each = 3),
+    Model = paste0('(', Model, ')'),
+    Model = ifelse(duplicated(Model), '', Model)
+  ) %>% 
+  select(Model, everything())
+
+names(phytab) <- gsub('^.*\\.', '', names(phytab))
+vrs <- gsub('(^.*),\\s.*$', '\\1', names(phytab))
+vrs[duplicated(vrs)] <- ''
+vrs <- paste0(vrs, '<br>')
+vrs[vrs %in% 'env_lab<br>'] <- 'Parameter<br>'
+vrs2 <- gsub('^.*,\\s|Model|env_lab', '', names(phytab))
+vrs <- paste(vrs, vrs2, sep = '')
+vrs <- gsub('abu', 'Abundance', vrs)
+vrs <- gsub('len', 'Length', vrs)
+vrs <- gsub('dis', 'Shell dissolution', vrs)
+
+names(phytab) <- vrs
+
+knitr::kable(phytab, caption = "Table 2: Model results for pteropod physiological and population response to pairs of co-occurring environmental variables.
+The estimated joint effects of variables in each model are shown.
+The overall R-squred value for each model is also shown.
+Significance of each effect is noted as * p < 0.05, ** p < 0.005.")
+```

### 2018-03-27 figures.Rmd
better labels on correlation heatmap

@@
 levs <- c(sort(envchr), sort(biochr), sort(phychr))
+labs <- c('Omega[ar]', 'CO[3]^-2', 'chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>% 
@@
     cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(levs)), 
-    var2 = factor(var2, levels = rev(levs)), 
+    var1 = factor(var1, levels = rev(levs), labels = rev(labs)), 
+    var2 = factor(var2, levels = rev(levs), labels = rev(labs)), 
     sig = gsub('ns', '', sig)
@@
   axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = unit(c(0,4,0,0), "lines"),
+  legend.position = c(0.5, 1.12),
+  legend.direction = 'horizontal',
+  plot.margin = unit(c(4,4,0,0), "lines"),
   strip.background = element_blank(), 
@@
                     ymin = outlab$y[3], ymax = outlab$y[3], xmin = 23, xmax = 23) +
+  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 23.5, ymax = 23.5) +
+  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 23.5, ymax = 23.5) +  
+  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 23.5, ymax = 23.5) +
   pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
+  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) + 
+  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
   scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +

### 2018-03-29 figures.Rmd
switch axis fig 4

@@
   .[[1]] %>% 
-  get_pldat(., 'O2', fct = 0.85)
+  get_pldat(., 'pCO2', fct = 0.9)
 
 p4 <- ggplot() +
-  geom_line(data = pl4[[1]], aes(x = pCO2, y = abu, group = O2, colour = O2), size = 1) + 
+  geom_line(data = pl4[[1]], aes(x = O2, y = abu, group = pCO2, colour = pCO2), size = 1) + 
   geom_text(data = pl4[[2]], aes(x= x, y = y, label = lab), hjust = 0) +

### 2018-03-30 R/dat_proc.R
do not use abundance from south range

@@ -37,7 +37,7 @@ biodat <- read_excel('raw/PteropodIntegrated_oxibiomarkers.xlsx',
   ungroup
 
 # abundance
-abudat <- read_excel('raw/abundances to be usef_Marcus.xlsx') %>% 
+abudat <- read_excel('raw/Copy of abundances to be usef_Marcus2.xlsx') %>% 
   rename(
     CTD = `CTD station`,
     abund = `real abundance`

### 2018-04-02 biomark.Rmd
ggally added with smooths on biomark

@@
 library(plotly)
+library(GGally)
 
@@
 pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
+```
+
+Pairs plot showing only LPX, SOD, ORAC/LPX with omega, O2, T:
+```{r, fig.height = 8, fig.width = 8}
+dat2 <- ptedat %>% 
+  inner_join(envdat, by = 'CTD') %>% 
+  select(one_of(c('LPX', 'SOD', 'ORACvLPX', 'Ara', 'O2', 'Temp')))
+
+# Function to return points and geom_smooth
+my_fn <- function(data, mapping, method="loess", ...){
+      p <- ggplot(data = data, mapping = mapping) + 
+      geom_point() + 
+      geom_smooth(method=method, ...)
+      p
+    }
+
+# Default loess curve    
+ggpairs(dat2, lower = list(continuous = my_fn))
 ```

### 2018-04-02 biomark.Rmd
effects plots with confidence intervals, added observed data to surface plots

@@
 
+  # observed data
+  toplo <- md()$model
+  names(toplo) <- c('z', 'x', 'y')
+
+  # surface data
   dat <- prd_vl() %>% 
     spread_(key_col = names(.)[2], value_col = names(.)[1]) %>% 
-    .[,-1] %>% 
     as.matrix
+  x <- colnames(dat)[-1] %>% as.numeric
+  y <- dat[, 1]
+  dat <- dat[, -1]
 
@@
     aspectmode = 'manual', 
-    aspectratio = list(x = 1, y = 1, z = 0.5), 
-    xaxis = list(title = names(prd_vl()[3])),
-    yaxis = list(title = names(prd_vl()[2])),
+    aspectratio = list(x = 1, y = 1, z = 0.6), 
+    xaxis = list(title = names(prd_vl()[2])),
+    yaxis = list(title = names(prd_vl()[3])),
     zaxis = list(title = names(prd_vl()[1]))
@@
 
-  p <- plot_ly(z = ~dat) %>% add_surface(colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
-    layout(scene = scene)
+  p <- plot_ly() %>% 
+    add_surface(z = ~dat, x = ~x, y = ~y, colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
+  add_markers(data = toplo, x = ~x, y = ~y, z = ~z) %>% 
+    layout(scene = scene, width = 800, height = 400)
   p
-
 })

### 2018-04-02 figures.Rmd
effects plots with confidence intervals, added observed data to surface plots

@@
 p1 <- ggplot() +
+  geom_ribbon(data = pl1[[1]], aes(x = Ara, ymin = LPX - se.fit, ymax = LPX + se.fit, group = pH), alpha = 0.5, fill = 'grey') +
   geom_line(data = pl1[[1]], aes(x = Ara, y = LPX, group = pH, colour = pH), size = 1) + 
@@
 p2 <- ggplot() +
+    geom_ribbon(data = pl2[[1]], aes(x = Ara, ymin = LPX - se.fit, ymax = LPX + se.fit, group = Temp), alpha = 0.5, fill = 'grey') +
   geom_line(data = pl2[[1]], aes(x = Ara, y = LPX, group = Temp, colour = Temp), size = 1) + 
@@
 p3 <- ggplot() +
+    geom_ribbon(data = pl3[[1]], aes(x = Ara, ymin = SOD - se.fit, ymax = SOD + se.fit, group = Temp), alpha = 0.5, fill = 'grey') +
   geom_line(data = pl3[[1]], aes(x = Ara, y = SOD, group = Temp, colour = Temp), size = 1) + 
@@
 p4 <- ggplot() +
+  geom_ribbon(data = pl4[[1]], aes(x = O2, ymin = abu - se.fit, ymax = abu + se.fit, group = pCO2), alpha = 0.5, fill = 'grey') +
   geom_line(data = pl4[[1]], aes(x = O2, y = abu, group = pCO2, colour = pCO2), size = 1) + 
@@
 p5 <- ggplot() +
+  geom_ribbon(data = pl5[[1]], aes(x = Ara, ymin = dis - se.fit, ymax = dis + se.fit, group = Temp), alpha = 0.5, fill = 'grey') +
   geom_line(data = pl5[[1]], aes(x = Ara, y = dis, group = Temp, colour = Temp), size = 1) + 

### 2018-04-02 physio.Rmd
effects plots with confidence intervals, added observed data to surface plots

@@
 
-```{r}
+```{r, fig.height = 9}
 # model data

### 2018-04-02 R/funcs.R

@@ -96,7 +96,7 @@ get_pldat <- function(modin, cvar, pos = c('left', 'right'), fct = NULL){
     map(function(x) seq(x[1], x[2], length = 100))
   
   # quantiles for cvar
-  x[[cvar]] <- modin$model[[cvar]]%>% quantile
+  x[[cvar]] <- modin$model[[cvar]]%>% quantile(., c(0, 1))
   
   # make data frame
   nms <- names(x) 
@@ -104,8 +104,9 @@ get_pldat <- function(modin, cvar, pos = c('left', 'right'), fct = NULL){
   names(x) <- nms
   
   # get predictions, combine with exp vars
-  prd_vl <- predict(modin, newdata = x) %>% 
-    data.frame(., x)
+  prd_vl <- predict(modin, newdata = x, se = T) %>% 
+    data.frame(., x) %>% 
+    dplyr::select(-df, -residual.scale)
   names(prd_vl)[1] <- all.vars(formula(modin))[1]
   
   # min x axis values for quantile labels
@@ -126,7 +127,7 @@ get_pldat <- function(modin, cvar, pos = c('left', 'right'), fct = NULL){
   yval <- locs[[yvar]]
   xval <- locs[[xvar]] %>% unique %>% `*`(fct)
   xlab <- data.frame(
-    lab = c('Max', '75th', 'Med', '25th', 'Min'), 
+    lab = c('Max', 'Min'), 
     x = xval, y = yval,
     stringsAsFactors = F)
   dr <- locs[[cvar]] %>% range %>% diff %>% sign

### 2018-04-12 biomark.Rmd
no mods with vif > 11, removed CO3 from analysis

@@
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
+envchr <- c('pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
 ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
@@
 
-Pairwise models with interactions for each biomarker
+Pairwise models with interactions for each biomarker, only pairwise models with VIF < 11 are shown
 
@@
 
-selectInput("mod", "Choose model:", paste0('mod', seq(1:nrow(mods))))
+selectInput("mod", "Choose model:", mod_all[['Model']])
 

### 2018-04-12 physio.Rmd
no mods with vif > 11, removed CO3 from analysis

@@
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
+envchr <- c('pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
 ptechr <- c('abu', 'dis', 'len', 'ty2', 'ty3', 'scr')

### 2018-04-12 R/dat_proc.R

@@ -5,14 +5,13 @@ source('R/funcs.R')
 
 # environmental data
 envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
-  select(`CTD Station`, `Latitude`, `'pCO2'`, `'pH'`, `'CO3'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
+  select(`CTD Station`, `Latitude`, `'pCO2'`, `'pH'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
   rename(
     CTD = `CTD Station`,
     Lat = `Latitude`,
     Ara = `'Aragonite'`, 
     O2 = `'Oxygen'`,
     pH = `'pH'`,
-    CO3 = `'CO3'`,
     Fluor = `'Fluorescence'`,
     pCO2 = `'pCO2'`,
     Temp = `'Temperature'`, 
@@ -127,7 +126,7 @@ save(expdat, file = 'data/expdat.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
+envchr <- c('pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
 ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
 
 dat_frm <- ptedat %>% 
@@ -152,6 +151,17 @@ env_cmb <- envchr %>%
 # models
 mod_all <- env_cmb %>%
   map(function(x, dat = dat_frm){
+
+    # check vif
+    rsq <- paste(x[2], '~', x[3]) %>% 
+      as.formula %>% 
+      lm(data = dat) %>% 
+      summary %>% 
+      .$r.squared 
+    vif <- 1 / (1 - rsq)
+      
+    # exit if vif large
+    if(vif > 11) return(NA)
     
     # formula
     frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
@@ -162,6 +172,7 @@ mod_all <- env_cmb %>%
     
   }) %>% 
   enframe('Model', 'Modobj') %>% 
+  filter(!map(.$Modobj, is.logical) %>% unlist) %>% 
[…]

### 2018-04-12 biomark.Rmd
only mods with low VIF, only select response/exp vars, model sel results, removed methods text

@@
 library(GGally)
+library(grid)
 
@@
 
-## Methods
-
-The response of pteropods to multiple stressors at the cellular, physiological, and population levels were evaluated using correlation analysis, multiple linear regression, and multivariate analysis.
-The general approach was to identify relationships among variables as potentially additive, synergistic, and antagonistic by evaluating pairwise effects of stressors on each response measure.
-Pearson correlations of environmental parameters (OA parameters, temperature, and chlorophyll) with cellular, physiological, and population response were first evaluated to identify potential associations with individual variables.
-Correlations between cellular and physiological responses were also evaluated to identify links between the two levels of biological organization.
-Non-continuous or skewed variables were transformed to better satisfy assumptions of the parametric tests (e.g., abundances were logarithmically transformed, proportions were arcsine transformed).
-All analyses were performed with the R statistical programming language (R Core Team 2017).
-
-Linear models were developed for every pairwise combination of environmental variables to evaluate additive and interactive effects on stressor response.
-Each model included separate terms for individual variables and a third term for the interaction of the pair.
-Evidence for additive effects were based solely on the magnitude of the estimated parameter for each variable, whereas synergistic or antagonistic effects were evaluated from the estimated parameter for the interaction.
-A positive interaction was evidence of a synergistic effect and a negative interaction was evidence of an antagonistic effect.
-Results of the linear models were further evaluated using effects plots to characterize the relationship of a pteropod response measure to continuous values for one stressor given two different values for the second stressor (constant at the minimum and maximum observed values).
-Thresholds for synergistic or antagonistic effects between two variables were identified from the effects plots as the intersection between the two fitted lines.
-
-Variance Inflation Factors (VIF) were quantified for all pairwise combinations of environmental variables used in the regression analysis to evaluate collinearity, such that:
-
-VIFj = 1 /( 1 − R2j) 		(1)
-
-where VIF is the reciprocal of the unexplained variance (1 − R2) of the linear regression of variable j against all other explanatory variables (Zuur et al., 2007).
-Although there are no specific VIF thresholds beyond which a variable is considered overly collinear, values less than 50 are generally considered adequate (Zuur et al.
-2007).
-Comparisons of variables indicated that all VIF were below this threshold, except combinations of CO3/Aragonite (70.6), pCO2/CO3 (51.4), and ph/pCO2 (79.2).
-
-Finally, multivariate comparisons between the cellular response measures and environmental variables were assessed using redundancy analysis (RDA) to jointly characterize relationships between all sampling stations.
-This analysis is conceptually similar to principal components analysis with an additional constraint on the environmental matrix, where the relationships are further partitioned based on covariance among response measures at each site in addition to the covariance between environmental variables.
-The final triplot (two biplots of environmental and response matrices) can be used to evaluate which environmental variables are correlated, as well as their relationships to the cellular response measures at each site.
-The environmental and response matrices were standardized to range from 0-1 prior to RDA.
-The vegan package for R was used for standardization and RDA (Oksanen et al.
-2017).
-
-Oksanen, J., Guillaume Blanchet, F.,  Friendly, M., Kindt, R.,  Legendre, P., McGlinn, D., Minchin, P.R., O'Hara, R.B., Simpson, G.L., Solymos, P., Henry, M., Stevens, H., Szoecs, E., Wagner, H.
-2017.
-vegan: Community Ecology Package.
-R package version 2.4-5.
-https://CRAN.R-project.org/package=vegan
-
-R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
-
-Zuur, A.F., Ieno, E.N., Smith, G.M. 2007. Analysing Ecological DAta. Springer, New York, NY. 
-MOntgomery, D.C., Peck, E.A. 1992. Introduction to Linear Regression Analysis. Wiley, New York, NY.
-
 Pairs plot
@@
 pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
-```
-
-Pairs plot showing only LPX, SOD, ORAC/LPX with omega, O2, T:
-```{r, fig.height = 8, fig.width = 8}
-dat2 <- ptedat %>% 
-  inner_join(envdat, by = 'CTD') %>% 
-  select(one_of(c('LPX', 'SOD', 'ORACvLPX', 'Ara', 'O2', 'Temp')))
-
-# Function to return points and geom_smooth
-my_fn <- function(data, mapping, method="loess", ...){
-      p <- ggplot(data = data, mapping = mapping) + 
-      geom_point() + 
-      geom_smooth(method=method, ...)
-      p
-    }
-
-# Default loess curve    
-ggpairs(dat2, lower = list(continuous = my_fn))
 ```
@@
 
-# {.tabset}
+# Correlations
 
 ```{r}
+# physiology rda model
+dat_cor <- ptedat %>% 
+  select(one_of('CTD', ptechr)) %>% 
+  # na.omit %>% 
+  left_join(envdat, by = 'CTD') %>% 
+  data.frame %>% 
+  remove_rownames %>% 
+  column_to_rownames('CTD')
+
 # all correlations
-crs <- crossing(var1 = names(dat_frm), var2 = names(dat_frm)) %>% 
+crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>% 
   filter(var1 != var2) %>% 
@@
       # variables
-      vr1 <- dat_frm[[x$var1]]
-      vr2 <- dat_frm[[x$var2]]
+      vr1 <- dat_cor[[x$var1]]
+      vr2 <- dat_cor[[x$var2]]
 
@@
 
-      # spearman
-      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
-      sp_cr <- round(sp_ts$estimate, 2)
-      sp_pv <- p_ast(sp_ts$p.value)
-      sp <- paste(sp_cr, sp_pv)
-
-      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
+      out <- data.frame(pr = pr, stringsAsFactors = F)
       return(out)
@@
   select(-rowname)
-
-sp_cr <- crs %>% 
-  select(var1, var2, sp) %>% 
-  spread(var2, sp) %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  remove_rownames %>% 
-  column_to_rownames('var1')
-
-pr_cr <- crs %>% 
-  select(var1, var2, pr) %>% 
-  spread(var2, pr) %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  remove_rownames %>% 
-  column_to_rownames('var1') 
 ```
 
-## Spearman correlations
-<div class="verysmall">
-
-```{r}
-knitr::kable(sp_cr, caption = 'Spearman rank correlations')
-```
-
-</div>
-
-```{r spheat, fig.height = 5, fig.width = 10}
-spplo <- crs %>% 
-  select(-pr) %>% 
-  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
+```{r, fig.height = 6, fig.width = 6}
+levs <- c(sort(envchr), sort(ptechr))
+labs <- c('Omega[ar]', 'chla', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
+prplo <- crs %>% 
+  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
+  filter(var1 %in% levs & var2 %in% levs) %>%  
   mutate(
     cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(unique(.$var1)))
+    var1 = factor(var1, levels = rev(levs), labels = rev(labs)), 
+    var2 = factor(var2, levels = rev(levs), labels = rev(labs)), 
+    sig = gsub('ns', '', sig)
   )
@@
   axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
+  legend.position = c(0.5, 1.2),
+  legend.direction = 'horizontal',
+  plot.margin = unit(c(8,4,0,0), "lines"),
   strip.background = element_blank(), 
@@
 
-p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
-  geom_tile(colour = 'black') + 
-  pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
-  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
-  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
-p
-```
-
-## Pearson correlations
-<div class="verysmall">
-
-```{r}
-knitr::kable(pr_cr, caption = 'Pearson correlations')
-```
-
-</div>
-
-```{r prheat, fig.height = 5, fig.width = 10}
-prplo <- crs %>% 
-  select(-sp) %>% 
-  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
-  mutate(
-    cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(unique(.$var1)))
+outlab <- data.frame(
+  y = c(4.5, 11.5), 
+  lab = c('Cellular', 'Environment')
 )
 
-pbase <- theme(
-  panel.grid.major = element_blank(), 
-  panel.grid.minor = element_blank(), 
-  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
-  axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
-  strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
-  panel.background = element_rect(fill = 'black')
-  )
+p <- ggplot(prplo) + 
+  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') + 
+  geom_text(aes(y = var1, x = var2, label = sig)) +
+  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 15, xmax = 15) +
+  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 15, xmax = 15) +
+  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 15.5, ymax = 15.5) +
+  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 15.5, ymax = 15.5) +
+  pbase +
+  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) + 
+  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
+  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
+  geom_hline(yintercept = 8.5, size = 1.5) +
+  geom_vline(xintercept = 8.5, size = 1.5)
 
-p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
-  geom_tile(colour = 'black') + 
-  pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
-  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
-  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
-p
+# Code to override clipping
+gt <- ggplot_gtable(ggplot_build(p))
+gt$layout$clip[gt$layout$name == "panel"] <- "off"
+grid.draw(gt)
 ```
@@
 
-# {.tabset}
-
-Pairwise models with interactions for each biomarker, only pairwise models with VIF < 11 are shown
+# Model results {.tabset}
 
@@
 mod_all <- biomod
+```
+
+## LPX
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'LPX'), caption = 'LPX')
+```
+
+## ORAC
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'ORAC'), caption = 'ORAC')
+```
+
+## SOD
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'SOD'), caption = 'SOD')
+```
+
+# {.tabset}
+
+Pairwise models with interactions for relevant response measures, only pairwise models with VIF < 11 are shown
+
+```{r}
 
@@
 
-levs1 <- reactive({
-
-  levs1 <- quantile(md()$model[, 2],  c(0, 1)) %>% 
-    list
-  names(levs1) <- names(md()$model)[2]
-  return(levs1)
-
-})
-
-levs2 <- reactive({
-
-  levs2 <- quantile(md()$model[, 3], c(0, 1)) %>% 
-    list
-  names(levs2) <- names(md()$model)[3]
-  return(levs2)
-
-})
-
 trm <- reactive({
 
-  md() %>% coefficients %>% names %>% .[4]
+  coefnms <- md() %>% coefficients %>% names
+  if(length(coefnms) < 3) return(NULL)
+  else out <- coefnms[3]
+
+  if(input$cvr) out <- coefnms[2]
+
+  return(out)
 
@@
 
-renderPlot({
-  plot(effect(trm(), md(), xlevels = levs1()), multiline = T)
-}, height = 300, width = 600)
-
-renderPlot({
-  plot(effect(trm(), md(), xlevels = levs2()), multiline = T)
-}, height = 300, width = 600)
+selectInput("cvr", "Switch covariate:", c(F, T))
 
@@
 
-  pbase <- theme(
-    panel.grid.major = element_blank(), 
-    panel.grid.minor = element_blank(), 
-    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
-    axis.text.y = element_text(size = 10),
-    legend.position = 'top', 
-    plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
-    strip.background = element_blank(), 
-    strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5)
-  )
+  if(is.null(trm())) stop("Model is too simple, you broke me Nina!")
 
-  p <- ggplot(prd_vl(), aes_string(y = names(prd_vl())[2], x = names(prd_vl())[3], fill = names(prd_vl())[1])) + 
-    geom_tile(colour = 'black') + 
-    theme_bw(base_family = 'serif') +
-    pbase +
-    scale_y_continuous(expand = c(0, 0)) +
-    scale_x_continuous(expand = c(0, 0)) +
-    scale_fill_distiller(palette = 'Spectral') +
-    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 15, label.theme = element_text(size = 12, angle = 0)))
-  p
-}, height = 700, width = 700)
+  # colors
+  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
+
+  # get data to plot
+  pl <- md() %>% 
+    get_pldat(., trm())
+  pl1 <- pl[[1]]
+  pl2 <- pl[[2]]
+
+# names for aes_string
+nms <- names(pl1)
+
+p1 <- ggplot() +
+  geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
+  geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
+  geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
+  theme_bw() +
+  scale_x_continuous(nms[3]) +
+  scale_colour_gradientn(colours = cols) + 
+  scale_y_continuous(nms[1])
+
+return(p1)
+
+}, height = 300, width = 600)
 
@@
   toplo <- md()$model
+  if(ncol(toplo) < 3) stop("Model is too simple, you broke me Nina!")
+
   names(toplo) <- c('z', 'x', 'y')
@@
 ```
-
-## CAT
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'CAT'), caption = 'CAT')
-```
-
-## GR
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'GR'), caption = 'GR')
-```
-
-## GSHonGSSG
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'GSHonGSSG'), caption = 'GSHonGSSG')
-```
-
-## GST
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'GST'), caption = 'GST')
-```
-
-## LPX
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'LPX'), caption = 'LPX')
-```
-
-## ORAC
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'ORAC'), caption = 'ORAC')
-```
-
-## SOD
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'SOD'), caption = 'SOD')
-```
-
-## ORACvLPX
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
-```

### 2018-04-12 figures.Rmd
only mods with low VIF, only select response/exp vars, model sel results, removed methods text

@@
 
-envchr <- c('Lat', 'pCO2', 'pH', 'CO3', 'Ara', 'O2', 'Temp', 'Fluor')
+envchr <- c('Lat', 'pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
 biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
@@
 levs <- c(sort(envchr), sort(biochr), sort(phychr))
-labs <- c('Omega[ar]', 'CO[3]^-2', 'chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
+labs <- c('Omega[ar]', 'chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>% 
@@
   annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
-                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 23, xmax = 23) +
+                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 22, xmax = 22) +
   annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
-                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 23, xmax = 23) +  
+                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 22, xmax = 22) +  
   annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
-                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 23, xmax = 23) +
+                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 22, xmax = 22) +
   annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
-                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 23.5, ymax = 23.5) +
+                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 22.5, ymax = 22.5) +
   annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
-                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 23.5, ymax = 23.5) +  
+                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 22.5, ymax = 22.5) +  
   annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
-                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 23.5, ymax = 23.5) +
+                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 22.5, ymax = 22.5) +
   pbase +
@@
 
-```{r effbio, fig.height = 8, fig.width = 9}
+```{r effbio, fig.height = 8, fig.width = 9, eval = F}
 # cellular 92 (lpx, ara, ph - syn), 101 (lpx, ara, t - neg add), 164 (sod, ara, t, neg but no thrsh) 
@@
 
-```{r effphy, fig.height = 4, fig.width = 9}
+```{r effphy, fig.height = 4, fig.width = 9, eval = F}
 
@@
 
-```{r}
+```{r, eval = F}
 mods <- readxl::read_excel('raw/info for table.xlsx') %>% 
@@
 
-```{r}
+```{r, eval = F}
 phytab <- mods$phy %>% 

### 2018-04-12 physio.Rmd
only mods with low VIF, only select response/exp vars, model sel results, removed methods text

@@
 library(plotly)
+library(grid)
 
@@
 
-# Physiological correlations with biomarkers {.tabset}
+# Correlations
 
 ```{r}
-tochk <- ptedat %>% 
-  select(-CTD, -Latitude)
+# physiology rda model
+dat_cor <- ptedat %>% 
+  select(one_of('CTD', ptechr)) %>% 
+  # na.omit %>% 
+  left_join(envdat, by = 'CTD') %>% 
+  data.frame %>% 
+  remove_rownames %>% 
+  column_to_rownames('CTD')
 
 # all correlations
-crs <- crossing(var1 = names(tochk), var2 = names(tochk)) %>% 
+crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>% 
   filter(var1 != var2) %>% 
@@
       # variables
-      vr1 <- tochk[[x$var1]]
-      vr2 <- tochk[[x$var2]]
+      vr1 <- dat_cor[[x$var1]]
+      vr2 <- dat_cor[[x$var2]]
 
@@
   select(-rowname)
-
-sp_cr <- crs %>% 
-  select(var1, var2, sp) %>% 
-  spread(var2, sp) %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  remove_rownames %>% 
-  column_to_rownames('var1')
-
-pr_cr <- crs %>% 
-  select(var1, var2, pr) %>% 
-  spread(var2, pr) %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  remove_rownames %>% 
-  column_to_rownames('var1') 
 ```
 
-## Spearman correlations
-
-<div class="verysmall">
-
-```{r}
-knitr::kable(sp_cr, caption = 'Spearman rank correlations')
-```
-
-</div>
-
-```{r , fig.height = 5, fig.width = 10}
-spplo <- crs %>% 
-  select(-pr) %>% 
-  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
+```{r, fig.height = 6, fig.width = 6}
+levs <- c(sort(envchr), sort(ptechr))
+labs <- c('Omega[ar]', 'chla', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
+prplo <- crs %>% 
+  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
+  filter(var1 %in% levs & var2 %in% levs) %>%  
   mutate(
     cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(unique(.$var1)))
+    var1 = factor(var1, levels = rev(levs), labels = rev(labs)), 
+    var2 = factor(var2, levels = rev(levs), labels = rev(labs)), 
+    sig = gsub('ns', '', sig)
   )
@@
 
-p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
-  geom_tile(colour = 'black') + 
-  pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
-  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
-  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
-p
-```
-
-## Pearson correlations
-<div class="verysmall">
-
-```{r}
-knitr::kable(pr_cr, caption = 'Pearson correlations')
-```
-
-</div>
-
-```{r , fig.height = 5, fig.width = 10}
-prplo <- crs %>% 
-  select(-sp) %>% 
-  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
-  mutate(
-    cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(unique(.$var1)))
+outlab <- data.frame(
+  y = c(3.5, 9.5), 
+  lab = c('Population/\nPhysiology', 'Environment')
 )
 
-pbase <- theme(
-  panel.grid.major = element_blank(), 
-  panel.grid.minor = element_blank(), 
-  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
-  axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
-  strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
-  panel.background = element_rect(fill = 'black')
-  )
+p <- ggplot(prplo) + 
+  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') + 
+  geom_text(aes(y = var1, x = var2, label = sig)) +
+  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 13, xmax = 13) +
+  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
+                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 13, xmax = 13) +
+  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 13.5, ymax = 13.5) +
+  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
+                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 13.5, ymax = 13.5) +
+  pbase +
+  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) + 
+  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
+  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
+  geom_hline(yintercept = 6.5, size = 1.5) +
+  geom_vline(xintercept = 6.5, size = 1.5)
 
-p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
-  geom_tile(colour = 'black') + 
-  pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
-  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
-  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
-p
-```
-
-# Physiological correlations with environment {.tabset}
-
-```{r}
-# all correlations
-crs <- crossing(var1 = names(dat_frm), var2 = names(dat_frm)) %>% 
-  filter(var1 != var2) %>% 
-  rownames_to_column() %>% 
-  group_by(rowname) %>% 
-  nest %>% 
-  mutate(
-    crs = map(data, function(x){
-
-      # variables
-      vr1 <- dat_frm[[x$var1]]
-      vr2 <- dat_frm[[x$var2]]
-
-      # pearson
-      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
-      pr_cr <- round(pr_ts$estimate, 2)
-      pr_pv <- p_ast(pr_ts$p.value)
-      pr <- paste(pr_cr, pr_pv)
-
-      # spearman
-      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
-      sp_cr <- round(sp_ts$estimate, 2)
-      sp_pv <- p_ast(sp_ts$p.value)
-      sp <- paste(sp_cr, sp_pv)
-
-      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
-      return(out)
-
-    })
-  ) %>% 
-  unnest %>% 
-  select(-rowname)
-
-sp_cr <- crs %>% 
-  select(var1, var2, sp) %>% 
-  spread(var2, sp) %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  remove_rownames %>% 
-  column_to_rownames('var1')
-
-pr_cr <- crs %>% 
-  select(var1, var2, pr) %>% 
-  spread(var2, pr) %>% 
-  data.frame(stringsAsFactors = F) %>% 
-  remove_rownames %>% 
-  column_to_rownames('var1') 
-```
-
-## Spearman correlations
-<div class="verysmall">
-
-```{r}
-knitr::kable(sp_cr, caption = 'Spearman rank correlations')
-```
-
-</div>
-
-```{r spheat, fig.height = 5, fig.width = 10}
-spplo <- crs %>% 
-  select(-pr) %>% 
-  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
-  mutate(
-    cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(unique(.$var1)))
-  )
-
-pbase <- theme(
-  panel.grid.major = element_blank(), 
-  panel.grid.minor = element_blank(), 
-  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
-  axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
-  strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
-  panel.background = element_rect(fill = 'black')
-  )
-
-p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
-  geom_tile(colour = 'black') + 
-  pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
-  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
-  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
-p
-```
-
-## Pearson correlations
-<div class="verysmall">
-
-```{r}
-knitr::kable(pr_cr, caption = 'Pearson correlations')
-```
-
-</div>
-
-```{r prheat, fig.height = 5, fig.width = 10}
-prplo <- crs %>% 
-  select(-sp) %>% 
-  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
-  mutate(
-    cor = as.numeric(cor), 
-    var1 = factor(var1, levels = rev(unique(.$var1)))
-  )
-
-pbase <- theme(
-  panel.grid.major = element_blank(), 
-  panel.grid.minor = element_blank(), 
-  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
-  axis.text.y = element_text(size = 8),
-  legend.position = 'top', 
-  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
-  strip.background = element_blank(), 
-  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
-  panel.background = element_rect(fill = 'black')
-  )
-
-p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
-  geom_tile(colour = 'black') + 
-  pbase +
-  scale_y_discrete('', expand = c(0, 0)) + 
-  scale_x_discrete('', expand = c(0, 0)) +
-  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
-  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
-p
+# Code to override clipping
+gt <- ggplot_gtable(ggplot_build(p))
+gt$layout$clip[gt$layout$name == "panel"] <- "off"
+grid.draw(gt)
 ```
@@
 
-# {.tabset}
+# Model results {.tabset}
 
-Pairwise models with interactions for each biomarker, only pairwise models with VIF < 11 are shown
-
-```{r, fig.height = 9}
+```{r}
 # model data
@@
 mod_all <- phymod
+```
+
+## abundance
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'abu'), caption = 'abu')
+```
+
+## dissolution
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'dis'), caption = 'dis')
+```
+
+# {.tabset}
+
+Pairwise models with interactions for relevant response measures, only pairwise models with VIF < 11 are shown
+
+```{r, fig.height = 9}
 
@@
 ```
-
-## abundance
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'abu'), caption = 'abu')
-```
-
-## dissolution
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'dis'), caption = 'dis')
-```
-
-## type 2 dissolution
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'ty2'), caption = 'ty2')
-```
-
-## type 3 dissolution
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'ty3'), caption = 'ty3')
-```
-
-## scarring
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'scr'), caption = 'scr')
-```
-
-## length
-```{r}
-knitr::kable(filter(mods, pte_lab %in% 'len'), caption = 'len')
-```

### 2018-04-12 R/dat_proc.R

@@ -126,8 +126,8 @@ save(expdat, file = 'data/expdat.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
-ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
+envchr <- c('pCO2', 'Ara', 'Temp')
+ptechr <- c('LPX', 'ORAC', 'SOD')
 
 dat_frm <- ptedat %>% 
   select(one_of('CTD', ptechr)) %>% 
@@ -168,13 +168,29 @@ mod_all <- env_cmb %>%
       as.formula
     
     # model only
-    lm(frm, data = dat)
+    datin <- dat[, x] %>% na.omit
+    modout <- lm(frm, data = datin, na.action = na.pass) %>% 
+      dredge %>% 
+      get.models(subset = 1) %>% 
+      .[[1]]
+    
+    # check pval
+    pval <- modout %>%
+      summary %>%
+      .$fstatistic
+    
+    # pval <- p.adjust(pval, method = 'holm', n = 88)
+    if(is.null(pval)) return(NA)
+    else pval <- pf(pval[1], pval[2], pval[3], lower.tail = F)
+    if(pval >= 0.05) return(NA)
+
+    return(modout)
     
   }) %>% 
   enframe('Model', 'Modobj') %>% 
   filter(!map(.$Modobj, is.logical) %>% unlist) %>% 
   mutate(
-    Model = gsub('^X', 'mod', Model), 
+    Model = paste0('mod', 1:nrow(.)),
     data = map(Modobj, function(x){
       
       # response
@@ -183,12 +199,12 @@ mod_all <- env_cmb %>%
       # rsq
       rsq <- summary(x) %>% 
         .$r.squared
-      
+
       # coeff summary
       out <- x %>% 
         summary %>% 
         .$coefficients %>% 
-        .[-1, c(1, 4)] %>% 
+        .[-1, c(1, 4), drop = F] %>% 
         data.frame %>% 
         rownames_to_column('env_lab') %>% 
         mutate(
@@ -214,8 +230,8 @@ save(biomod, file = 'data/biomod.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
[…]

### 2018-04-12 R/funcs.R

@@ -102,11 +102,16 @@ get_pldat <- function(modin, cvar, pos = c('left', 'right'), fct = NULL){
   nms <- names(x) 
   x <- crossing(x[[1]], x[[2]])
   names(x) <- nms
-  
+  x <- x[, c(names(x)[!names(x) %in% cvar], cvar)]
+
   # get predictions, combine with exp vars
   prd_vl <- predict(modin, newdata = x, se = T) %>% 
     data.frame(., x) %>% 
-    dplyr::select(-df, -residual.scale)
+    dplyr::select(-df, -residual.scale) %>% 
+    mutate(
+      hi = fit + se.fit, 
+      lo = fit - se.fit
+      )
   names(prd_vl)[1] <- all.vars(formula(modin))[1]
   
   # min x axis values for quantile labels

### 2018-04-13 R/dat_proc.R
added oracvlpx, len, o2 v ara for abund

@@ -1,5 +1,6 @@
 library(readxl)
 library(tidyverse)
+library(MuMIn)
 
 source('R/funcs.R')
 
@@ -127,7 +128,7 @@ data(envdat)
 data(ptedat)
 
 envchr <- c('pCO2', 'Ara', 'Temp')
-ptechr <- c('LPX', 'ORAC', 'SOD')
+ptechr <- c('LPX', 'ORAC', 'ORACvLPX', 'SOD')
 
 dat_frm <- ptedat %>% 
   select(one_of('CTD', ptechr)) %>% 
@@ -230,8 +231,8 @@ save(biomod, file = 'data/biomod.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'Ara', 'Temp')
-ptechr <- c('abu', 'dis')
+envchr <- c('pCO2', 'Ara', 'Temp', 'O2')
+ptechr <- c('abu', 'dis', 'len')
 
 dat_frm <- ptedat %>% 
   select(one_of('CTD', ptechr)) %>% 
@@ -247,19 +248,12 @@ env_cmb <- envchr %>%
   combn(2) %>% 
   t %>% 
   data.frame(stringsAsFactors = F) %>%
+  filter(!(X1 %in% 'pCO2' & X2 %in% 'O2')) %>%  # remove pCO2 and O2 comps
   crossing(ptechr, .) %>% 
   t %>% 
   data.frame(stringsAsFactors = F) %>% 
   as.list
 
-# # back-transform abu for models
-# dat_frm <- dat_frm %>% 
-#   mutate(
-#     abu = (10^abu) - 1,
-#     abu = as.integer(round(abu, 0))
-#   )
-#   
-# models
 mod_all <- env_cmb %>%
   
   map(function(x, dat = dat_frm){
@@ -295,7 +289,7 @@ mod_all <- env_cmb %>%
     if(is.null(pval)) return(NA)
     else pval <- pf(pval[1], pval[2], pval[3], lower.tail = F)
     if(pval >= 0.05) return(NA)
-    
+
     return(modout)    
     
   }) %>% 

### 2018-04-13 R/dat_proc.R
add flourescence

@@ -127,7 +127,7 @@ save(expdat, file = 'data/expdat.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'Ara', 'Temp')
+envchr <- c('pCO2', 'Ara', 'Temp', 'Fluor')
 ptechr <- c('LPX', 'ORAC', 'ORACvLPX', 'SOD')
 
 dat_frm <- ptedat %>% 
@@ -231,8 +231,8 @@ save(biomod, file = 'data/biomod.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'Ara', 'Temp', 'O2')
-ptechr <- c('abu', 'dis', 'len')
+envchr <- c('pCO2', 'Ara', 'Temp', 'O2', 'Fluor')
+ptechr <- c('abu', 'dis')
 
 dat_frm <- ptedat %>% 
   select(one_of('CTD', ptechr)) %>% 

### 2018-04-13 doc/methods.Rmd
methods doc added and updated

---
title: "Methods"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
urlcolor: blue
link-citations: true
---

The general approach was to identify relationships among variables as potentially additive, synergistic, and antagonistic by evaluating pairwise effects of stressors on each response measure. Pearson correlations of environmental parameters (OA parameters, temperature, and chlorophyll) with cellular, physiological, and population response were first evaluated to identify potential associations with individual variables. Correlations between cellular and physiological responses were also evaluated to identify links between the two levels of biological organization. Non-continuous or skewed variables were transformed to better satisfy assumptions of parametric tests (e.g., abundances were logarithmically transformed, proportions were arcsine transformed). All analyses were performed with the R statistical programming language (R Core Team 2017).

Multivariate comparisons between the cellular response measures and environmental variables were assessed using redundancy analysis (RDA) to jointly characterize relationships between all sampling stations. This analysis is conceptually similar to principal components analysis with an additional constraint on the environmental matrix, where the relationships are further partitioned based on covariance among response measures at each site in addition to the covariance between environmental variables. The final triplot (two biplots of environmental and response matrices) can be used to evaluate which environmental variables are correlated, as well as their relationships to the cellular response measures at each site. The environmental and response matrices were standardized to range from 0-1 prior to RDA. The vegan package for R was used for standardization and RDA (Oksanen et al. 2017).

Linear models were then developed for pairwise combinations of environmental variables to evaluate additive and interactive effects on stressor response. To reduce the likelihood of false positive results from multiple comparisons, variables were chosen a priori that were considered most relevant for describing pteropod response to stressors.  Only LPX, ORAC, ORACvLPX, and SOD biomarkers were evaluated for cellular response and only abundance, dissolution, and length were evaluated for physiological and population responses.  Environmental variables were selected for analysis that were orthogonal in multivariate space to reduce collinearity and included aragonite saturation, fluorescence, pCO2, and temperature. Oxygen and aragonite saturation were evaluated for relationships with physiologial and population response measures.

Variance Inflation Factors (VIF) were quantified for all pairwise combinations of environmental variables to estimate potential collinearity in each model, such that:

  VIFj = 1 /( 1 − R2j) 		(1)

where VIF is the reciprocal of the unexplained variance (1 − R2) of the linear regression of variable j against all other explanatory variables. Zuur et al. (2007) suggest that VIF values less than fifty may be appropriate for analysis, but we chose a value of ten for excluding combinations of environmental variables. We chose this conservative value to further reduce the potential for false positive results by reducing the number of combinations that were evaluated, in addition to reducing the likelihood of spurious results from collinear explanatory variables.

Linear models for the selected pairwise combinations of environmental variables and response measures included separate terms for individual variables and a third term for the interaction of the pair.  A model selection procedure was then used to compare every smaller subset of the global model to identify the most parsimonious solution. The final model for each pairwise combination was chosen based on a minimization of corrected Akaike Information Criterion (Burnham and Anderson 2002, Barton 2018).  Further, models with probability values greater than alpha of 0.05 for the overall model fit were excluded. These p-values were not adjusted for multiple comparisons due to the relatively small sample sizes of each model (n = 11 for all, except n = 35 for abundance).

Evidence for additive effects were based solely on the magnitude of the estimated parameter for each variable, whereas synergistic or antagonistic effects were evaluated from the estimated parameter for the interaction if included. A positive interaction was evidence of a synergistic effect and a negative interaction was evidence of an antagonistic effect. Results of the linear models were further evaluated using effects plots to characterize the relationship of a pteropod response measure to continuous values for one stressor given two different values for the second stressor (constant at the minimum and maximum observed values).

Barton, K. 2018. MuMIn: Multi-Model Inference. R package version 1.40.4. https://CRAN.R-project.org/package=MuMIn

Burnham, K.P., Anderson, D.R. 2002. Model selection and multimodel inference: a practical information-theoretic approach. 2nd ed. New York, Springer-Verlag.

Oksanen, J., Guillaume Blanchet, F.,  Friendly, M., Kindt, R.,  Legendre, P., McGlinn, D., Minchin, P.R., O'Hara, R.B., Simpson, G.L., Solymos, P., Henry, M., Stevens, H., Szoecs, E., Wagner, H. 2017. vegan: Community Ecology Package. R package version 2.4-5. https://CRAN.R-project.org/package=vegan

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

Zuur, A.F., Ieno, E.N., Smith, G.M. 2007. Analysing Ecological DAta. Springer, New York, NY.

### 2018-04-13 R/dat_proc.R

@@ -201,6 +201,9 @@ mod_all <- env_cmb %>%
       rsq <- summary(x) %>% 
         .$r.squared
 
+      # n
+      n <- nrow(x$model)
+      
       # coeff summary
       out <- x %>% 
         summary %>% 
@@ -210,11 +213,12 @@ mod_all <- env_cmb %>%
         rownames_to_column('env_lab') %>% 
         mutate(
           pte_lab = yvar,
+          n = n, 
           Rsq = round(rsq, 2),
           Est = round(Estimate, 2), 
           Pvl = p_ast(Pr...t..)
         ) %>% 
-        select(pte_lab, env_lab, Rsq, Est, Pvl)
+        select(pte_lab, env_lab, n, Rsq, Est, Pvl)
       
       return(out)
       
@@ -306,6 +310,9 @@ mod_all <- env_cmb %>%
       rsq <- summary(x) %>% 
         .$r.squared
       
+      # n
+      n <- nrow(x$model)
+      
       # coeff summary
       out <- x %>% 
         summary %>% 
@@ -315,11 +322,12 @@ mod_all <- env_cmb %>%
         rownames_to_column('env_lab') %>% 
         mutate(
           pte_lab = yvar,
+          n = n,
           Rsq = round(rsq, 2),
           Est = round(Estimate, 2), 
           Pvl = p_ast(Pr...t..)
         ) %>% 
-        select(pte_lab, env_lab, Rsq, Est, Pvl)
+        select(pte_lab, env_lab, n, Rsq, Est, Pvl)
       
       return(out)
       

### 2018-04-18 figures.Rmd
mod tables added

@@
     code_folding: hide
+    css: styles.css
 ---
+
+<a href="https://github.com/fawda123/Pteropod_biomarker/" class="github-corner" aria-label="View source on Github"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#FD6C6C; color:#fff; position: absolute; top: 0; border: 0; left: 0; transform: scale(-1, 1);" aria-hidden="true"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a><style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style>
 
@@
 library(grid)
+library(stargazer)
 
@@
 ```
-Fig.
+<!-- Fig.
 3 Examples of model interactions of co-occuring environmental variables on cellular response measures.
@@
 Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.
+-->
 
@@
 ```
-Fig.
+<!-- Fig.
 4 Examples of model interactions of co-occuring environmental variables on abundance and shell dissolution.
@@
 Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.
+-->
 
-```{r, eval = F}
-mods <- readxl::read_excel('raw/info for table.xlsx') %>% 
-  .[-1, -1] %>% 
-  gather('var', 'Model', everything()) %>% 
-  mutate(
-    org = ifelse(var %in% c('abundance', 'dissolution', 'length'), 'phy', 'bio'),
-    Model = paste0('mod', Model)
-    ) %>% 
-  split(.$org)
+```{r, eval = T, results = 'asis'}
+biotab <- biomod %>% 
+  unnest(data) %>% 
+  group_by(Model) %>% 
+  mutate(n = n()) %>% 
+  filter(n > 1) %>% 
+  dplyr::select(-n)
+biotab <- filter(biomod, Model %in% unique(biotab$Model))
 
-biotab <- mods$bio %>% 
-  left_join(biomod, by = 'Model') %>% 
-  select(Model, data) %>%
-  mutate(data = map(data, function(x){
-    x %>% mutate(vr = seq(1:nrow(.)))
-  })) %>% 
-  unnest %>% 
-  mutate(
-    Model = gsub('mod', '', Model),
-    Model = as.numeric(Model)
-    ) %>% 
-  group_by(Model) %>% 
-  mutate(
-    Pvl = gsub('ns', '', Pvl), 
-    Rsq = ifelse(duplicated(Rsq), '', Rsq)
-    ) %>% 
-  ungroup %>% 
-  unite('Est', Est, Pvl, sep = '') %>% 
-  gather('estnm', 'estvl', Rsq, Est) %>% 
-  unite('est', pte_lab, estnm, sep = ', ', remove = F) %>% 
-  select(-estnm) %>% 
-  split(.$pte_lab) %>%
-  lapply(., function(x){
+cap <- 'Table 1: Model results for pteropod cellular response to pairs of co-occurring environmental variables.
+The estimated joint effects of variables and their significance in each model are shown.
+Parameter estimates with standard errors (in parentheses) are included with overall model statistics (R-squared, F statistic) at the bottom.
+Models with the same cellular response measure can be identified with the column names at the top.'
 
-    out <- spread(x, est, estvl) %>% 
-      select(-pte_lab) %>%
-      arrange(Model, vr) %>% 
-      select(-vr, -Model)
-
-    if(!any(grepl('^LPX', names(out))))
-      out <- out %>% 
-        select(-env_lab)
-
-    return(out)
-
-  }) %>% 
-  do.call('cbind', .) %>% 
-  mutate(
-    Model = rep(letters[1:(nrow(.)/3)], each = 3),
-    Model = paste0('(', Model, ')'),
-    Model = ifelse(duplicated(Model), '', Model)
-  ) %>% 
-  select(Model, everything())
-
-names(biotab) <- gsub('^.*\\.', '', names(biotab))
-vrs <- gsub('(^.*),\\s.*$', '\\1', names(biotab))
-vrs[duplicated(vrs)] <- ''
-vrs <- paste0(vrs, '<br>')
-vrs[vrs %in% 'env_lab<br>'] <- 'Parameter<br>'
-vrs2 <- gsub('^.*,\\s|Model|env_lab', '', names(biotab))
-vrs <- paste(vrs, vrs2, sep = '')
-
-names(biotab) <- vrs
-
-knitr::kable(biotab, caption = "Table 1: Model results for pteropod cellular response to pairs of co-occurring environmental variables.
-The estimated joint effects of variables in each model are shown.
-The overall R-squred value for each model is also shown.
-Significance of each effect is noted as * p < 0.05, ** p < 0.005.")
+stargazer(biotab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Cellular response measures')
 ```
 
-```{r, eval = F}
-phytab <- mods$phy %>% 
-  left_join(phymod, by = 'Model') %>% 
-  select(Model, data) %>%
-  mutate(data = map(data, function(x){
-    x %>% mutate(vr = seq(1:nrow(.)))
-  })) %>% 
-  unnest %>% 
-  mutate(
-    Model = gsub('mod', '', Model),
-    Model = as.numeric(Model)
-  ) %>% 
+```{r, results = 'asis'}
+phytab <- phymod %>% 
+  unnest(data) %>% 
   group_by(Model) %>% 
-  mutate(
-    Pvl = gsub('ns', '', Pvl), 
-    Rsq = ifelse(duplicated(Rsq), '', Rsq)
-  ) %>% 
-  ungroup %>% 
-  unite('Est', Est, Pvl, sep = '') %>% 
-  gather('estnm', 'estvl', Rsq, Est) %>% 
-  unite('est', pte_lab, estnm, sep = ', ', remove = F) %>% 
-  select(-estnm) %>% 
-  split(.$pte_lab) %>%
-  lapply(., function(x){
+  mutate(n = n()) %>% 
+  filter(n > 1) %>% 
+  dplyr::select(-n)
+phytab <- filter(phymod, Model %in% unique(phytab$Model))
 
-    out <- spread(x, est, estvl) %>% 
-      select(-pte_lab) %>%
-      arrange(Model, vr) %>% 
-      select(-vr, -Model)
+cap <- 'Table 2: Model results for physiological and abundance responses to pairs of co-occurring environmental variables.
+The estimated joint effects of variables and their significance in each model are shown.
+Parameter estimates with standard errors (in parentheses) are included with overall model statistics (R-squared, F statistic) at the bottom.
+Models with the same response measure can be identified with the column names at the top.'
 
-    if(!any(grepl('^abu', names(out))))
-      out <- out %>% 
-        select(-env_lab)
+stargazer(phytab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Physiological response measures', dep.var.labels = c('Abundance', 'Shell dissolution'))
 
-    return(out)
-
-  }) %>% 
-  do.call('cbind', .) %>% 
-  mutate(
-    Model = rep(letters[1:(nrow(.)/3)], each = 3),
-    Model = paste0('(', Model, ')'),
-    Model = ifelse(duplicated(Model), '', Model)
-  ) %>% 
-  select(Model, everything())
-
-names(phytab) <- gsub('^.*\\.', '', names(phytab))
-vrs <- gsub('(^.*),\\s.*$', '\\1', names(phytab))
-vrs[duplicated(vrs)] <- ''
-vrs <- paste0(vrs, '<br>')
-vrs[vrs %in% 'env_lab<br>'] <- 'Parameter<br>'
-vrs2 <- gsub('^.*,\\s|Model|env_lab', '', names(phytab))
-vrs <- paste(vrs, vrs2, sep = '')
-vrs <- gsub('abu', 'Abundance', vrs)
-vrs <- gsub('len', 'Length', vrs)
-vrs <- gsub('dis', 'Shell dissolution', vrs)
-
-names(phytab) <- vrs
-
-knitr::kable(phytab, caption = "Table 2: Model results for pteropod physiological and population response to pairs of co-occurring environmental variables.
-The estimated joint effects of variables in each model are shown.
-The overall R-squred value for each model is also shown.
-Significance of each effect is noted as * p < 0.05, ** p < 0.005.")
 ```

### 2018-04-18 R/dat_proc.R

@@ -6,10 +6,11 @@ source('R/funcs.R')
 
 # environmental data
 envdat <- read_excel('raw/MeanSurface100m2016WCOA.xlsx') %>% 
-  select(`CTD Station`, `Latitude`, `'pCO2'`, `'pH'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
+  select(`CTD Station`, `Latitude`, `Longitude`, `'pCO2'`, `'pH'`, `'Fluorescence'`, `'Aragonite'`, `'Oxygen'`, `'Temperature'`, `'Salinity'`, `'Alkalinity'`) %>% 
   rename(
     CTD = `CTD Station`,
     Lat = `Latitude`,
+    Lon = `Longitude`,
     Ara = `'Aragonite'`, 
     O2 = `'Oxygen'`,
     pH = `'pH'`,
@@ -201,8 +202,10 @@ mod_all <- env_cmb %>%
       rsq <- summary(x) %>% 
         .$r.squared
 
-      # n
-      n <- nrow(x$model)
+      # pval
+      pval <- summary(x) %>%
+        .$fstatistic
+      pval <- pf(pval[1], pval[2], pval[3], lower.tail = F)
       
       # coeff summary
       out <- x %>% 
@@ -213,12 +216,12 @@ mod_all <- env_cmb %>%
         rownames_to_column('env_lab') %>% 
         mutate(
           pte_lab = yvar,
-          n = n, 
+          pval = p_ast(pval), 
           Rsq = round(rsq, 2),
           Est = round(Estimate, 2), 
           Pvl = p_ast(Pr...t..)
         ) %>% 
-        select(pte_lab, env_lab, n, Rsq, Est, Pvl)
+        select(pte_lab, env_lab, pval, Rsq, Est, Pvl)
       
       return(out)
       
@@ -310,8 +313,10 @@ mod_all <- env_cmb %>%
       rsq <- summary(x) %>% 
         .$r.squared
       
-      # n
-      n <- nrow(x$model)
+      # pval
+      pval <- summary(x) %>%
+        .$fstatistic
+      pval <- pf(pval[1], pval[2], pval[3], lower.tail = F)
       
       # coeff summary
       out <- x %>% 
@@ -322,12 +327,12 @@ mod_all <- env_cmb %>%
         rownames_to_column('env_lab') %>% 
         mutate(
           pte_lab = yvar,
[…]

### 2018-04-25 physio.Rmd
three explanatory variables for dissolution

@@
 library(grid)
+library(MuMIn)
 
@@
 ```
+
+Three explanatory variables:
+
+```{r}
+envchr <- c('Ara', 'Temp', 'O2')
+ptechr <- c('dis')
+
+dat_frm <- ptedat %>% 
+  select(one_of('CTD', ptechr)) %>% 
+  left_join(envdat, by = 'CTD') %>% 
+  data.frame %>% 
+  remove_rownames %>% 
+  column_to_rownames('CTD') %>% 
+  na.omit
+
+frm <- formula(c('dis ~ (Ara + Temp + O2)^2'))
+modoutglo <- lm(frm, data = dat_frm, na.action = na.pass)
+
+modoutsel <- lm(frm, data = dat_frm, na.action = na.pass) %>% 
+  dredge %>% 
+  get.models(subset = 1) %>% 
+  .[[1]]
+
+summary(modoutglo)
+
+summary(modoutsel)
+
+```

### 2018-05-20 doc/discussion.Rmd
discussion paragraph

---
title: "Discussion"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
urlcolor: blue
link-citations: true
---

Our analysis approach to evaluate the effects of co-occurring stressors on pteropod response measures was designed to balance the tradeoffs between characteristics of the dataset and the ability to identify an ecologically meaningful signal.  Several characteristics of the dataset presented challenges that required a conservative statistical design that was related primarily to the sample size and the occurrence of several correlated environmental variables.  Our restriction to evaluate only variables below our chosen threshold for collinearity and use of response measures that were chosen a priori reflects this limitation.  Further, our pairwise evaluation of co-occurring variables on the cellular and physiological response measures provided an indication of relative importance, but it is not a definitive explanation of mechanisms.  Causal description through purely correlative statistics is problematic in these cases.  The interpretation of our results must be considered in this context, yet patterns shown by the models are consistent with known mechanisms of pteropod response to OA stressors or changes in temperature. Our ability to evaluate joint effects of stressors is a novel contribution of this study and the interpretation within the limits of the analysis is supported by past work (citations).  This study is the first to evaluate pteropod response to both OA stressors and warming events as they occur in the environment.  As such, the results can inform further work with the goal of identifying how multiple stressors interact to influence pteropods at multiple levels of biological organization.

### 2018-05-20 doc/discussion.Rmd
recommendation sentence

@@
 Our ability to evaluate joint effects of stressors is a novel contribution of this study and the interpretation within the limits of the analysis is supported by past work (citations).
-This study is the first to evaluate pteropod response to both OA stressors and warming events as they occur in the environment.
+This study is the first to evaluate pteropod response to both OA stressors and warming events as they occur in the environment and our statistical design could be used in similar contexts to address challenges that are common in environmental datasets.
 As such, the results can inform further work with the goal of identifying how multiple stressors interact to influence pteropods at multiple levels of biological organization.

### 2018-07-06 figures.Rmd
working on figures for revisions

@@
   remove_rownames %>% 
-  column_to_rownames('CTD')
+  column_to_rownames('CTD') %>% 
+  rename(
+    abundance = abu, 
+    dissolution = dis, 
+    length = len, 
+    scarring = scr, 
+    typeII = ty2, 
+    typeIII = ty3
+  )
 
@@
   decostand(method = 'range')
-ptephy <- select(dat_phy, one_of(phychr)) %>% 
+ptephy <- select(dat_phy, -one_of(envchr)) %>% 
   decostand(method = 'range')
@@
 Fig.
-1 Results of redundancy analyses for environmental variables with (a) cellular and (b) physiological endpoints of pteropod response to OA stressors.
+2 Results of redundancy analyses for environmental variables with (a) cellular and (b) physiological endpoints of pteropod response to OA stressors.
 Points are site locations in multivariate space with the size proportional to latitude.
@@
 levs <- c(sort(envchr), sort(biochr), sort(phychr))
-labs <- c('Omega[ar]', 'chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
+labs <- c('Omega[ar]', 'chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>% 
@@
 Fig.
-2 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
+3 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
 Darker red values are strong positive correlations and darker purple values are strong negative correlations.
@@
 
-```{r effbio, fig.height = 8, fig.width = 9, eval = F}
-# cellular 92 (lpx, ara, ph - syn), 101 (lpx, ara, t - neg add), 164 (sod, ara, t, neg but no thrsh) 
-biosel <- biomod %>% 
-  filter(Model %in% paste0('mod', c(92, 101, 164))) %>% 
-  dplyr::select(-data)
+```{r effbio, fig.height = 8, fig.width = 6, eval = T}
+toplo <- biomod %>% 
+  filter(Model %in% c('mod3', 'mod4', 'mod17')) %>% 
+  mutate(
+    cvar = c('pCO2', 'Temp', 'Temp'),
+    xvar = c('Fluorescence', 'Omega[ar]', 'Omega[ar]'), 
+    yvar = c('LPX', 'LPX', 'SOD'), 
+    legr = c('pCO[2]', 'Temp', 'Temp'),
+    fct = c(0.5, 0.95, 0.95)
+    ) %>% 
+  mutate(
+    pldat = pmap(list(Modobj, cvar, fct), function(Modobj, cvar, fct){
 
+      get_pldat(Modobj, cvar, fct = fct)
+
+    }),
+    plos = pmap(list(pldat, xvar, legr), function(pldat, xvar, legr){
+
+      pl1 <- pldat[[1]]
+      pl2 <- pldat[[2]]
+
+      # colors
       cols <- RColorBrewer::brewer.pal(9, 'RdBu')
 
-##
-# plots
-
-# mod 92 cell (lpx, ara, ph - syn)
-pl1 <- filter(biosel, grepl('92$', Model)) %>% 
-  .$Modobj %>% 
-  .[[1]] %>% 
-  get_pldat(., 'pH')
+      # names for aes_string
+      nms <- names(pl1)
 
       p1 <- ggplot() +
-  geom_ribbon(data = pl1[[1]], aes(x = Ara, ymin = LPX - se.fit, ymax = LPX + se.fit, group = pH), alpha = 0.5, fill = 'grey') +
-  geom_line(data = pl1[[1]], aes(x = Ara, y = LPX, group = pH, colour = pH), size = 1) + 
-  geom_text(data = pl1[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
+        geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
+        geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
+        geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
         theme_bw() +
-  scale_x_continuous(expression(Omega[Ar])) +
-  scale_colour_gradientn(colours = cols) + 
-  scale_y_continuous('LPX') +
-  ggtitle("(a) Synergistic effect of pH on aragonite\nsaturation")
+        xlab(parse(text = xvar)) +
+        scale_colour_gradientn(parse(text = legr), colours = cols) +
+        scale_y_continuous(nms[1])
 
-# mod 101 cell (lpx, ara, t - neg add), 164 (sod, ara, t, neg but no thrsh) 
-pl2 <- filter(biosel, grepl('101$', Model)) %>% 
-  .$Modobj %>% 
-  .[[1]] %>% 
-  get_pldat(., 'Temp')
+      return(p1)
 
-p2 <- ggplot() +
-    geom_ribbon(data = pl2[[1]], aes(x = Ara, ymin = LPX - se.fit, ymax = LPX + se.fit, group = Temp), alpha = 0.5, fill = 'grey') +
-  geom_line(data = pl2[[1]], aes(x = Ara, y = LPX, group = Temp, colour = Temp), size = 1) + 
-  geom_text(data = pl2[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
-  theme_bw() +
-  scale_x_continuous(expression(Omega[Ar])) +
-  scale_colour_gradientn(colours = cols) + 
-  scale_y_continuous('LPX') +
-  ggtitle("(b) Negative additive effect below threshold of\n temperature on aragonite saturation")
+    })
+  )
 
-# mod 164 cell (sod, ara, t, neg but no thrsh) 
-pl3 <- filter(biosel, grepl('164$', Model)) %>% 
-  .$Modobj %>% 
-  .[[1]] %>% 
-  get_pldat(., 'Temp')
+p1 <- toplo$plos[[1]] + ggtitle('(b)')
+p2 <- toplo$plos[[2]] + ggtitle('(a)')
+p3 <- toplo$plos[[3]] + ggtitle('(c)')
 
-p3 <- ggplot() +
-    geom_ribbon(data = pl3[[1]], aes(x = Ara, ymin = SOD - se.fit, ymax = SOD + se.fit, group = Temp), alpha = 0.5, fill = 'grey') +
-  geom_line(data = pl3[[1]], aes(x = Ara, y = SOD, group = Temp, colour = Temp), size = 1) + 
-  geom_text(data = pl3[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
-  theme_bw() +
-  scale_x_continuous(expression(Omega[Ar])) +
-  scale_colour_gradientn(colours = cols) + 
-  scale_y_continuous('SOD') +
-  ggtitle("(c) Negative additive effect of temperature on \naragonite saturation")
+pA <- ggplot_gtable(ggplot_build(p1))
+pB <- ggplot_gtable(ggplot_build(p2))
+pC <- ggplot_gtable(ggplot_build(p3))
 
-grid.arrange(p1, p2, p3, ncol = 2)
+maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[2:3])
+
+pA$widths[2:3] <- maxWidth
+pB$widths[2:3] <- maxWidth
+pC$widths[2:3] <- maxWidth
+
+grid.arrange(
+ pB, pA, pC, ncol = 1 
+)
 
 ```
-<!-- Fig.
-3 Examples of model interactions of co-occuring environmental variables on cellular response measures.
-Each subplot shows a different relationship as either additive or synergistic effects between the variables.
-All y-axes are transformed to conform to model output.
-Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.
--->
+Fig.
+4 Examples of model interactions of co-occuring environmental variables on pteropod cellular responses.
+Plot (a) shows a negative additive effect between aragonite saturation state (Ωar) and temperature on cellular toxicity (lipid peroxidation; LPX); plot (b) shows negative additive effects of food availability and pCO2 on LPX; plot (c) shows positive additive effects of temperature and Ωar on antioxidative activity (sodium dismutase, SOD).
+Both y-axes are transformed to conform to model output.
+Covarying environmental variables were held constant at the minimum and maximum values in the observed data.
 
-```{r effphy, fig.height = 4, fig.width = 9, eval = F}
+```{r effphy, fig.height = 5.33, fig.width = 6, eval = T}
+toplo <- phymod %>% 
+  filter(Model %in% c('mod2', 'mod9')) %>% 
+  mutate(
+    cvar = c('Temp', 'Temp'),
+    xvar = c('Omega[ar]', 'Omega[ar]'), 
+    yvar = c('Abundance', 'Dissolution'), 
+    legr = c('Temp', 'Temp'),
+    pos = c('left', 'right'), 
+    fct = c(0.92, 1.04), 
+    hjs = c(0, 1)
+    ) %>% 
+  mutate(
+    pldat = pmap(list(Modobj, cvar, pos, fct), function(Modobj, cvar, pos, fct){
 
-# physio 4 (abu, o2, pco2 - pos additive), 38 (dis, ara, t - syn)
-physel <- phymod %>% 
-  filter(Model %in% paste0('mod', c(4, 38))) %>% 
-  dplyr::select(-data)
+      get_pldat(Modobj, cvar, pos = pos, fct = fct)
 
+    }),
+    plos = pmap(list(pldat, xvar, yvar, legr, hjs), function(pldat, xvar, yvar, legr, hjs){
+
+      pl1 <- pldat[[1]]
+      pl2 <- pldat[[2]]
+
+      # colors
       cols <- RColorBrewer::brewer.pal(9, 'RdBu')
 
-# mod 4 physio (abu, o2, pco2 - pos additive)
-pl4 <- filter(physel, grepl('4$', Model)) %>% 
-  .$Modobj %>% 
-  .[[1]] %>% 
-  get_pldat(., 'pCO2', fct = 0.9)
+      # names for aes_string
+      nms <- names(pl1)
 
-p4 <- ggplot() +
-  geom_ribbon(data = pl4[[1]], aes(x = O2, ymin = abu - se.fit, ymax = abu + se.fit, group = pCO2), alpha = 0.5, fill = 'grey') +
-  geom_line(data = pl4[[1]], aes(x = O2, y = abu, group = pCO2, colour = pCO2), size = 1) + 
-  geom_text(data = pl4[[2]], aes(x= x, y = y, label = lab), hjust = 0) +
+      p1 <- ggplot() +
+        geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
+        geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
+        geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = hjs) +
         theme_bw() +
-  scale_colour_gradientn(colours = cols) + 
-  scale_y_continuous('Abundance') +
-  ggtitle("(a) Synergistic effect of O2 on\npCO2 stress")
+        xlab(parse(text = xvar)) +
+        scale_colour_gradientn(parse(text = legr), colours = cols) +
+        scale_y_continuous(yvar)
 
-# mod 38 physio (dis, ara, t - syn)
-pl5 <- filter(physel, grepl('38$', Model)) %>% 
-  .$Modobj %>% 
-  .[[1]] %>% 
-  get_pldat(., 'Temp', 'right', fct = 1.05)
+      return(p1)
 
-p5 <- ggplot() +
-  geom_ribbon(data = pl5[[1]], aes(x = Ara, ymin = dis - se.fit, ymax = dis + se.fit, group = Temp), alpha = 0.5, fill = 'grey') +
-  geom_line(data = pl5[[1]], aes(x = Ara, y = dis, group = Temp, colour = Temp), size = 1) + 
-  geom_text(data = pl5[[2]], aes(x= x, y = y, label = lab), hjust = 1) +
-  theme_bw() +
-  scale_colour_gradientn(colours = cols) + 
-  scale_y_continuous('Shell dissolution') +
-  scale_x_continuous(expression(Omega[Ar])) +
-  ggtitle("(b) Synergistic effect of temperature on \naragonite saturation")
+    })
+  )
 
-grid.arrange(p4, p5, ncol = 2)
+p1 <- toplo$plos[[1]] + ggtitle('(b)')
+p2 <- toplo$plos[[2]] + ggtitle('(a)')
+
+pA <- ggplot_gtable(ggplot_build(p1))
+pB <- ggplot_gtable(ggplot_build(p2))
+
+maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3])
+
+pA$widths[2:3] <- maxWidth
+pB$widths[2:3] <- maxWidth
+
+grid.arrange(
+ pB, pA, ncol = 1 
+)
+
 ```
-<!-- Fig.
-4 Examples of model interactions of co-occuring environmental variables on abundance and shell dissolution.
-Each subplot shows a different relationship as either additive or synergistic effects between the variables.
+Fig.
+5 An example of model interactions of co-occuring environmental variables with biomineralization (shell dissolution) and population (abundance) responses.
+The interaction effects between aragonite saturation state (Ωar) and temperature on shell dissolution, as well as the interaction between Ωar and oxygen is synergistic (a, b, respectively).
 All y-axes are transformed to conform to model output.
-Covarying environmental variables were held constant at the minimum, 25th, median, 75th, and maximum values in the observed data.
--->
+Covarying environmental variables were held constant at the minimum and maximum values in the observed data.
 

### 2018-07-06 physio.Rmd
checking three factor model for length - ns
working on figures for revisions

@@
 levs <- c(sort(envchr), sort(ptechr))
-labs <- c('Omega[ar]', 'chla', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
+labs <- c('Omega[ar]', 'chla', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>% 
@@
 
-Three explanatory variables:
+### Three explanatory variables
+
+Shell dissolution:
 
@@
 ```
+
+Length/growth:
+
+```{r}
+envchr <- c('Ara', 'Temp', 'O2')
+ptechr <- c('len')
+
+dat_frm <- ptedat %>% 
+  select(one_of('CTD', ptechr)) %>% 
+  left_join(envdat, by = 'CTD') %>% 
+  data.frame %>% 
+  remove_rownames %>% 
+  column_to_rownames('CTD') %>% 
+  na.omit
+
+frm <- formula(c('len ~ (Ara + Temp + O2)^2'))
+modoutglo <- lm(frm, data = dat_frm, na.action = na.pass)
+
+modoutsel <- lm(frm, data = dat_frm, na.action = na.pass) %>% 
+  dredge %>% 
+  get.models(subset = 1) %>% 
+  .[[1]]
+
+summary(modoutglo)
+
+summary(modoutsel)
+
+```

### 2018-07-20 figures.Rmd
revisions for frontiers to nina

@@
 
+scalbio <- range(dat_bio$Lat)
+scalbio <- seq(scalbio[1], scalbio[2], length  = 5)
+scalphy <- range(dat_phy$Lat)
+scalphy <- seq(scalphy[1], scalphy[2], length  = 5)
+
 p1a <- ggord(biorda, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_bio$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
-  ggtitle('(a) Cellular endpoints')
+  ggtitle('(a) Cellular endpoints') + 
+  scale_size(breaks = scalbio, labels = round(scalbio, 1), range = c(1, 8))
 p2a <- ggord(phyrda, ptslab = T, parse = T, repel = repel, coord_fix = coord_fix, addsize = addsize, size = dat_phy$Lat, sizelab = 'Latitude', alpha = alpha, arrow = arrow, ellipse = F) + 
-  ggtitle('(b) Physiological endpoints')
+  ggtitle('(b) Physiological endpoints') + 
+  scale_size(breaks = scalphy, labels = round(scalphy, 1), range = c(1, 8))
 
@@
 levs <- c(sort(envchr), sort(biochr), sort(phychr))
-labs <- c('Omega[ar]', 'chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
+labs <- c('Omega[ar]', 'Fluor', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>% 
@@
 Darker red values are strong positive correlations and darker purple values are strong negative correlations.
-Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*).
+Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*\*).
 
@@
 The estimated joint effects of variables and their significance in each model are shown.
-Parameter estimates with standard errors (in parentheses) are included with overall model statistics (R-squared, F statistic) at the bottom.
-Models with the same cellular response measure can be identified with the column names at the top.'
+Parameter estimates are included with overall model statistics (R-squared, F statistic) at the bottom.
+Models with the same cellular response measure can be identified with the column names at the top.
+* p < 0.1, ** p < 0.05, *** p < 0.01'
 
-stargazer(biotab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Cellular response measures')
+stargazer(biotab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Cellular response measures', report = 'vc*')
 ```
@@
 The estimated joint effects of variables and their significance in each model are shown.
-Parameter estimates with standard errors (in parentheses) are included with overall model statistics (R-squared, F statistic) at the bottom.
-Models with the same response measure can be identified with the column names at the top.'
+Parameter estimates are included with overall model statistics (R-squared, F statistic) at the bottom.
+Models with the same response measure can be identified with the column names at the top.
+* p < 0.1, ** p < 0.05, *** p < 0.01'
 
-stargazer(phytab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Physiological response measures', dep.var.labels = c('Abundance', 'Shell dissolution'))
+stargazer(phytab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Physiological response measures', dep.var.labels = c('Abundance', 'Shell dissolution'), report = 'vc*')
 

### 2018-07-23 figures.Rmd
O2 back in cellular models, updated figures.Rmd

@@
 toplo <- biomod %>% 
-  filter(Model %in% c('mod3', 'mod4', 'mod17')) %>% 
+  filter(Model %in% c('mod3', 'mod4', 'mod24')) %>% 
   mutate(

### 2018-07-23 R/dat_proc.R

@@ -128,7 +128,7 @@ save(expdat, file = 'data/expdat.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'Ara', 'Temp', 'Fluor')
+envchr <- c('pCO2', 'Ara', 'Temp', 'O2', 'Fluor')
 ptechr <- c('LPX', 'ORAC', 'ORACvLPX', 'SOD')
 
 dat_frm <- ptedat %>% 

### 2018-07-24 R/dat_proc.R
ph back in mods save dat_proc

@@ -128,7 +128,7 @@ save(expdat, file = 'data/expdat.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'Ara', 'Temp', 'O2', 'Fluor')
+envchr <- c('pCO2', 'Ara', 'Temp', 'O2', 'pH', 'Fluor')
 ptechr <- c('LPX', 'ORAC', 'ORACvLPX', 'SOD')
 
 dat_frm <- ptedat %>% 
@@ -238,7 +238,7 @@ save(biomod, file = 'data/biomod.RData', compress = 'xz')
 data(envdat)
 data(ptedat)
 
-envchr <- c('pCO2', 'Ara', 'Temp', 'O2', 'Fluor')
+envchr <- c('pCO2', 'Ara', 'Temp', 'O2', 'pH', 'Fluor')
 ptechr <- c('abu', 'dis')
 
 dat_frm <- ptedat %>% 

### 2018-07-24 figures.Rmd
vif > 5 excluded

@@
 toplo <- biomod %>% 
-  filter(Model %in% c('mod3', 'mod4', 'mod24')) %>% 
+  filter(Model %in% c('mod2', 'mod3', 'mod23')) %>% 
   mutate(
@@
 toplo <- phymod %>% 
-  filter(Model %in% c('mod2', 'mod9')) %>% 
+  filter(Model %in% c('mod2', 'mod8')) %>% 
   mutate(

### 2018-07-24 R/dat_proc.R

@@ -163,7 +163,7 @@ mod_all <- env_cmb %>%
     vif <- 1 / (1 - rsq)
       
     # exit if vif large
-    if(vif > 11) return(NA)
+    if(vif > 5) return(NA)
     
     # formula
     frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 
@@ -274,7 +274,7 @@ mod_all <- env_cmb %>%
     vif <- 1 / (1 - rsq)
     
     # exit if vif large
-    if(vif > 11) return(NA)
+    if(vif > 5) return(NA)
     
     # formula
     frm <- paste(x[1], '~', x[2], '*', x[3]) %>% 

### 2018-07-25 figures.Rmd
tables redone with vif < 5, tables without constant

@@
 Models with the same cellular response measure can be identified with the column names at the top.
-* p < 0.1, ** p < 0.05, *** p < 0.01'
+* p < 0.05, ** p < 0.01'
 
-stargazer(biotab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Cellular response measures', report = 'vc*')
+stargazer(biotab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Cellular response measures', report = 'vc*', star.char = c("", "*", "**"), omit = 'Constant')
 ```
@@
 Models with the same response measure can be identified with the column names at the top.
-* p < 0.1, ** p < 0.05, *** p < 0.01'
+* p < 0.05, ** p < 0.01'
 
-stargazer(phytab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Physiological response measures', dep.var.labels = c('Abundance', 'Shell dissolution'), report = 'vc*')
+stargazer(phytab$Modobj, type = 'html', title = cap, digits = 2, keep.stat = c('adj.rsq'), intercept.top = T, intercept.bottom = F, dep.var.caption = 'Physiological response measures', dep.var.labels = c('Abundance', 'Shell dissolution'), report = 'vc*', star.char = c("", "*", "**"), omit = 'Constant')
 

### 2018-07-30 figures.Rmd
figure 7 updates

@@
 Fig.
-2 Results of redundancy analyses for environmental variables with (a) cellular and (b) physiological endpoints of pteropod response to OA stressors.
+4 Results of redundancy analyses for environmental variables with (a) cellular and (b) physiological endpoints of pteropod response to OA stressors.
 Points are site locations in multivariate space with the size proportional to latitude.
@@
 Fig.
-3 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
+5 Correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
 Darker red values are strong positive correlations and darker purple values are strong negative correlations.
@@
 Fig.
-4 Examples of model interactions of co-occuring environmental variables on pteropod cellular responses.
+6 Examples of model interactions of co-occuring environmental variables on pteropod cellular responses.
 Plot (a) shows a negative additive effect between aragonite saturation state (Ωar) and temperature on cellular toxicity (lipid peroxidation; LPX); plot (b) shows negative additive effects of food availability and pCO2 on LPX; plot (c) shows positive additive effects of temperature and Ωar on antioxidative activity (sodium dismutase, SOD).
@@
 
-```{r effphy, fig.height = 5.33, fig.width = 6, eval = T}
-toplo <- phymod %>% 
-  filter(Model %in% c('mod2', 'mod8')) %>% 
+```{r effphy, fig.height = 5.33, fig.width = 12, eval = T}
+toplo <- rbind(
+    phymod[phymod$Model == 'mod8', ],
+    phymod[phymod$Model == 'mod8', ],
+    phymod[phymod$Model == 'mod10', ],
+    phymod[phymod$Model == 'mod2', ]
+  ) %>% 
   mutate(
-    cvar = c('Temp', 'Temp'),
-    xvar = c('Omega[ar]', 'Omega[ar]'), 
-    yvar = c('Abundance', 'Dissolution'), 
-    legr = c('Temp', 'Temp'),
-    pos = c('left', 'right'), 
-    fct = c(0.92, 1.04), 
-    hjs = c(0, 1)
+    cvar = c('Temp', 'Ara', 'O2', 'Temp'),
+    xvar = c('Omega[ar]', 'Temp', 'Temp','Omega[ar]'), 
+    yvar = c('Dissolution', 'Dissolution', 'Dissolution', 'Abundance'), 
+    legr = c('Temp', 'Omega[ar]', 'O[2]','Temp'),
+    pos = c('right', 'left', 'left','left'), 
+    fct = c(1.04, 0.97, 0.97, 0.92), 
+    hjs = c(1, 0, 0, 0)
     ) %>% 
@@
 
-p1 <- toplo$plos[[1]] + ggtitle('(b)')
-p2 <- toplo$plos[[2]] + ggtitle('(a)')
+p1 <- toplo$plos[[1]] + ggtitle('(a)')
+p2 <- toplo$plos[[2]] + ggtitle('(b)')
+p3 <- toplo$plos[[3]] + ggtitle('(c)')
+p4 <- toplo$plos[[4]] + ggtitle('(d)')
 
@@
 pB <- ggplot_gtable(ggplot_build(p2))
+pC <- ggplot_gtable(ggplot_build(p3))
+pD <- ggplot_gtable(ggplot_build(p4))
 
-maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3])
+maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[2:3], pD$widths[2:3])
 
@@
 pB$widths[2:3] <- maxWidth
+pC$widths[2:3] <- maxWidth
+pD$widths[2:3] <- maxWidth
 
 grid.arrange(
- pB, pA, ncol = 1 
+ pA, pB, pC, pD, ncol = 2 
 )
@@
 Fig.
-5 An example of model interactions of co-occuring environmental variables with biomineralization (shell dissolution) and population (abundance) responses.
-The interaction effects between aragonite saturation state (Ωar) and temperature on shell dissolution, as well as the interaction between Ωar and oxygen is synergistic (a, b, respectively).
+7 Examples of model interactions and additive effects of co-occurring environmental variables with biomineralization (shell dissolution) and population (abundance) responses.
+Co-occurring effects between aragonite saturation state (Ωar) and temperature on shell dissolution are shown in (a, b), temperature and O2 on dissolution in (c), and temperature and aragonite saturation state on abundance in (d).
 All y-axes are transformed to conform to model output.

### 2018-07-31 figures.Rmd
removed 7b from figs

@@
 
-```{r effphy, fig.height = 5.33, fig.width = 12, eval = T}
+```{r effphy, fig.height = 8, fig.width = 6, eval = T}
 toplo <- rbind(
-    phymod[phymod$Model == 'mod8', ],
     phymod[phymod$Model == 'mod8', ],
@@
   mutate(
-    cvar = c('Temp', 'Ara', 'O2', 'Temp'),
-    xvar = c('Omega[ar]', 'Temp', 'Temp','Omega[ar]'), 
-    yvar = c('Dissolution', 'Dissolution', 'Dissolution', 'Abundance'), 
-    legr = c('Temp', 'Omega[ar]', 'O[2]','Temp'),
-    pos = c('right', 'left', 'left','left'), 
-    fct = c(1.04, 0.97, 0.97, 0.92), 
-    hjs = c(1, 0, 0, 0)
+    cvar = c('Temp', 'O2', 'Temp'),
+    xvar = c('Omega[ar]', 'Temp','Omega[ar]'), 
+    yvar = c('Dissolution', 'Dissolution', 'Abundance'), 
+    legr = c('Temp', 'O[2]','Temp'),
+    pos = c('right', 'left','left'), 
+    fct = c(1.04, 0.97, 0.92), 
+    hjs = c(1, 0, 0)
     ) %>% 
@@
 p3 <- toplo$plos[[3]] + ggtitle('(c)')
-p4 <- toplo$plos[[4]] + ggtitle('(d)')
 
@@
 pC <- ggplot_gtable(ggplot_build(p3))
-pD <- ggplot_gtable(ggplot_build(p4))
 
-maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[2:3], pD$widths[2:3])
+maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[2:3])
 
@@
 pC$widths[2:3] <- maxWidth
-pD$widths[2:3] <- maxWidth
 
 grid.arrange(
- pA, pB, pC, pD, ncol = 2 
+ pA, pB, pC, ncol = 1 
 )
@@
 7 Examples of model interactions and additive effects of co-occurring environmental variables with biomineralization (shell dissolution) and population (abundance) responses.
-Co-occurring effects between aragonite saturation state (Ωar) and temperature on shell dissolution are shown in (a, b), temperature and O2 on dissolution in (c), and temperature and aragonite saturation state on abundance in (d).
+Co-occurring effects between aragonite saturation state (Ωar) and temperature on shell dissolution are shown in (a), temperature and O2 on dissolution in (b), and temperature and aragonite saturation state on abundance in (c).
 All y-axes are transformed to conform to model output.

### 2018-08-03 figures.Rmd
flour to chla lab, figures only

@@
   select(-Lat) %>% 
+  rename(Chla = Fluor) %>% 
   decostand(method = 'range')
@@
   select(-Lat) %>%
+  rename(Chla = Fluor) %>% 
   decostand(method = 'range')
@@
 levs <- c(sort(envchr), sort(biochr), sort(phychr))
-labs <- c('Omega[ar]', 'Fluor', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
+labs <- c('Omega[ar]', 'Chla', 'Lat', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>% 
@@
 
-p1 <- toplo$plos[[1]] + ggtitle('(b)')
+p1 <- toplo$plos[[1]] + ggtitle('(b)') + xlab('Chla')
 p2 <- toplo$plos[[2]] + ggtitle('(a)')

### 2018-08-24 figures.Rmd
figs 6 and 7 legend as discrete

@@
     }),
-    plos = pmap(list(pldat, xvar, legr), function(pldat, xvar, legr){
+    plos = pmap(list(pldat, xvar, cvar, legr), function(pldat, xvar, cvar, legr){
 
@@
       pl2 <- pldat[[2]]
+
+      # leg vals as factor 
+      pl1[[cvar]] <- factor(pl1[[cvar]], 
+                            levels = sort(unique(pl1[[cvar]]), decreasing = T),
+                            labels = round(sort(unique(pl1[[cvar]]), decreasing = T), 1)
+      )
 
@@
         xlab(parse(text = xvar)) +
-        scale_colour_gradientn(parse(text = legr), colours = cols) +
+        scale_colour_manual(parse(text = legr), values = c(rev(cols)[1], cols[1])) +
         scale_y_continuous(nms[1])
@@
     }),
-    plos = pmap(list(pldat, xvar, yvar, legr, hjs), function(pldat, xvar, yvar, legr, hjs){
+    plos = pmap(list(pldat, xvar, yvar, cvar, legr, hjs), function(pldat, xvar, yvar, cvar, legr, hjs){
 
@@
         xlab(parse(text = xvar)) +
-        scale_colour_gradientn(parse(text = legr), colours = cols) +
+        scale_colour_manual(parse(text = legr), values = c(rev(cols)[1], cols[1])) +
         scale_y_continuous(yvar)

### 2018-08-28 doc/reviewer2_response.Rmd
response to reviewer 2 added

---
title: "Reviewer 2 response"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
urlcolor: blue
link-citations: true
---

__The information that the authors added about the methods is useful for better understanding the design of the study. I still have unresolved questions about the statistical analysis and therefore the results. It is still not clear how the statistical analyses were performed with respect to colinear drivers. For example, in Figure 4, you show that pCO2, pH, and O2 all covary - so I interpret that to mean that if you detect a significant correlation of XX bioindicator with pCO2, you cannot distinguish that effect from a correlation with O2. If pCO2 and O2 covary, you cannot separate their effects and test for them independently, as shown in Figure 5. Furthermore, in your forum response, you say "Because the VIF value for carbonate chemistry parameters and oxygen were above the threshold, we have not attempted to separate the effects of carbonate chemistry parameters (Ωar, pH and/or pCO2) from oxygen." However, O2 and pH and pCO2 are all shown with independent p-values in Figure 5. So, it is not clear to me how you used the VIF results. Perhaps additional text to the methods is needed to describe what you did with the results of the VIFs and how that influenced the statistical analysis that resultsed in Figure 5. Perhaps you could include the list of VIF values to support your decisions.__

We appreciate the reviewer's concern regarding clarification of the methods and we offer an explanation here to better describe our approach.  The reviewer is absolutely correct about collinearity between some of the drivers and the inability to separate the effects, as for pCO2, ph, and O2.  Our previous statistical design, although insufficiently described, was developed specifically to emphasize these limitations and to account for these issues in both analysis and interpretation of results.  Our analysis design first explored general relationships among environmental and pteropod response measures using multivariate analysis (RDA) and simple pairwise correlations (Figs. 4 and 5).  The second portion of our analysis provided a more formal framework to identify combined effects of stressors using linear models of pairwise combinations of environmental variables to explain pteropod response (Figs. 6 and 7).  This second portion was informed by the preceding exploratory analyses and included restrictions on which variables were evaluated due to collinaerity, using VIF thresholds.  As such, the results from the second set of analyses were used directly to infer relationships becuase of a robust design that accounted for collinearity.  Figures 4 and 5 were intended to demonstrate how the parameters covaried and we did not use them to directly infer causation.

To address these concerns, we have added some content to the methods to better describe the methods and partciularly to distinguish between the different objectives of the first and second set of analyses (beginning on line 213, first paragraph of statistical analyses section):

"Several of the environmental variables were correlated, which presented a challenge for inferring causal relationships with pteropod response measures.  To address this issue, the analysis was conducted in two stages where the first was exploratory and the second was a more formal approach using a subset of the environmental variables to directly infer causation.  The first set of analyses included all variables to demonstrate the complexity of the environmental dataset and directly informed the second set of analyses, whereby variables that were associated with the response measures were further screened to reduce collinearity in subsequent models.  Details for both sets of analyses follow."

Finally, we note that the VIF values for all pairwise comparisons of collinearity of the environmental variables can be viewed in the supplementary online material indicated at the end of the mansucript.  Therefore, we have not included additional tables for this information.

### 2018-09-28 doc/prcompex.Rmd
prcomp for oa environ params

---
output:
  html_document:
  keep_md: yes
self_contained: no
code_folding: hide
---

```{r, echo = F, warning = F, message = F}

knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(knitr)
library(gridExtra)
library(grid)
library(stargazer)
library(ggord)

source("../R/funcs.R")

load(file = '../data/envdat.RData')
load(file = '../data/ptedat.RData')
load(file = '../data/phymod.RData')
load(file = '../data/biomod.RData')

envchr <- c('Lat', 'pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
biochr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
phychr <- c('abu', 'dis', 'len', 'ty2', 'ty3', 'scr')

# physiology rda model
envpca <- ptedat %>% 
  select(one_of('CTD', phychr, biochr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD') %>% 
  select(one_of(envchr)) %>% 
  na.omit
prmod <- prcomp(envpca, center = T, scale = T)
```

### Principal components analysis of all environmental variables.

```{r pcaplo, fig.height = 5, fig.width = 8}
p1 <- ggord(prmod, vec_ext = 5, size = 2, alpha = 0.5)
p2 <- ggord(prmod, axes = c(2, 3), vec_ext = 5, size = 2, alpha = 0.5)
grid.arrange(p1, p2, ncol = 2)
```

Axes 1, 2, and 3 explain ~ 99% of the variation among environmental parameters: axis one explain OA parameters, axis 2 explains temp/lat, and axis 3 explains fluorescence.
```{r}
prmod
summary(prmod)
```

###  Correlation plot

The OA parameters oxygen, pH, aragonite saturation, and pCO2 were replaced with the first principal component axis for the correlation plot.

```{r corploPC1, fig.height = 6, fig.width = 7}
# physiology rda model
PC1 <- prmod$x[, 'PC1'] %>% 
  data.frame(PC1 = .) %>% 
  rownames_to_column('CTD') %>% 
  mutate(CTD = as.numeric(CTD))

dat_cor <- ptedat %>% 
  select(one_of('CTD', phychr, biochr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  left_join(PC1, by = 'CTD') %>% 
  column_to_rownames('CTD') %>% 
  select(-Ara, -pCO2, -O2, -pH)

# all correlations
crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>%
  filter(var1 != var2) %>%
  rownames_to_column() %>%
  group_by(rowname) %>%
  nest %>%
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- dat_cor[[x$var1]]
      vr2 <- dat_cor[[x$var2]]

      # pearson
      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
      pr_cr <- round(pr_ts$estimate, 2)
      pr_pv <- p_ast(pr_ts$p.value)
      pr <- paste(pr_cr, pr_pv)

      out <- data.frame(pr = pr, stringsAsFactors = F)
      return(out)

    })
  ) %>%
  unnest %>%
  select(-rowname)

levs <- c("Fluor", "Lat", "PC1", "Temp", sort(biochr), sort(phychr))
labs <- c('Chla', 'Lat', 'PC1', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
prplo <- crs %>%
  separate(pr, c('cor', 'sig'), sep = ' ') %>%
  filter(var1 %in% levs & var2 %in% levs) %>%
  mutate(
    cor = as.numeric(cor),
    var1 = factor(var1, levels = rev(levs), labels = rev(labs)),
    var2 = factor(var2, levels = rev(levs), labels = rev(labs)),
    sig = gsub('ns', '', sig)
  )

pbase <- theme(
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8),
  axis.text.y = element_text(size = 8),
  legend.position = c(0.5, 1.12),
  legend.direction = 'horizontal',
  plot.margin = unit(c(4,4,0,0), "lines"),
  strip.background = element_blank(),
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5),
  panel.background = element_rect(fill = 'black')
)

outlab <- data.frame(
  y = c(3.5, 10.5, 16.5),
  lab = c('Population/\nPhysiology', 'Cellular', 'Environment')
)

p <- ggplot(prplo) +
  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') +
  geom_text(aes(y = var1, x = var2, label = sig)) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 19, xmax = 19) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 19, xmax = 19) +
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 19, xmax = 19) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 19.5, ymax = 19.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 19.5, ymax = 19.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 19.5, ymax = 19.5) +
  pbase +
  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0))) +
  geom_hline(yintercept = 6.5, size = 1.5) +
  geom_hline(yintercept = 14.5, size = 1.5) +
  geom_vline(xintercept = 6.5, size = 1.5) +
  geom_vline(xintercept = 14.5, size = 1.5)

# Code to override clipping
gt <- ggplot_gtable(ggplot_build(p))
gt$layout$clip[gt$layout$name == "panel"] <- "off"
grid.draw(gt)
```

### 2018-10-01 doc/reviewer2_response2.Rmd
response to review2 again

---
title: "Reviewer 2 response 2"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
urlcolor: blue
link-citations: true
---

__The text addition to Methods section 2.4 is helpful, but it does not resolve my issue. I am specifically concerned about the statistical analyses behind Figure 5. I assume that these analyses were conducted using linear models for pairwise comparisons as described starting on p. 6, line 240. My argument remains that this is inappropriate, given that you found that some of the environmental variables were highly correlated in the exploratory RDA. Ultimately, the outcome may be similar and may not change your discussion and interpretation much, but I recommend redoing this part of the statistical analyses using collapsed variables that combine correlated environmental variables. For example, in the RDA, you found that carbonate chemistry parameters and dissolved oxygen are all highly correlated. Therefore, it is redundant and inappropriate to analyze their effects separately (separate rows and p-values in Figure 5). In that table, one variable that represents O2, pCO2, Omega_ar, and pH should presented (one row instead of four).__

__Furthermore, I think that latitude should be included in the RDA, as upwelling intensity generally increases with latitude in this system. Therefore, any correlations with abiotic varables and latitude should be identified and accounted for in the subsequent statistical analyses.__

We appreciate the concerns of the reviewer and we hope that our response here provides sufficient clarity on the intent of our analyses.  First and most importantly, we reiterate that the analysis in Figure 5 was presented to demonstrate exactly the issue that the reviewer notes and that the results jusified our conservative approach to screen collinear variables in the following analysis (i.e., those for Figs. 6 and 7).  Therefore, we feel there is value in presenting the correlations among all variables to emphasize the complexity of the system and to provide readers with an overview of associations that were more rigorously evaluated in the follow-up analysis.  The key distinction between figure 5 and the follow-up analyses is that we never claimed the results in figure 5 were causative.  Our conservative statistical design in the follow-up analyses provided us with a more solid foundation to infer mechanistic relationships.

We also emphasize that the text noted by the reviewer on p. 6, line 240 was not the analysis used for Figure 5.  The pairwise comparisons using linear models were the follow-up analyses that exluded collinear variables.  The stars in figure 5 are simply the significance of correlation coefficients (i.e., an association).

We feel there is still some disagreement about presenting correlation results in Figure 5 among multiple variables that co-vary, such as the parameters pH, pCO2, omega, and O2.  We recreated Figure 5 using the first principal component axis of a PCA of all environmental variables.  Nearly 100% of the variance was described by the first three axes: axis one explained the OA parameters pCO2, pH, O2, and Omega, axis 2 explained temperature and latitude, and axis 3 explained fluorescence.

![](prcompex_files/figure-html/pcaplo.png)
The first axis above was used in place of the four correlated OA parameters in the original Figure 5:

![](prcompex_files/figure-html/corploPC1.png)

Although this addresses the reviewer's concern, we feel this presentation complicates evaluation of the study system.  A reader would have to know which variables were described by PC1 and the approximate loadings.  We feel this presentation lacks the transparency provided by the original figure, nor does it change any of our conclusions.

Further, we want to emphasize the unique variation explained by latitude that was not associated with the OA parameters.  The loadings on the PC axes show a distinct separation of latitude on the second axis and the OA parameters on the first.  Therefore, we argue that latitude does not confound our interpration of pteropod response to acidification.

It is further noted that latitude is weakly associated with PC1 in the updated Figure 5, although this assocation is weak compared to the association with temperature.  Our PCA demonstrated that latitude is primarily associated with temperature and we did not include latitude in follow-up analyses because of no mechanistic interpretation on pteropod response.  We felt it unnecessary to include latitude in the RDA as a specific variable, but instead showed the points sized relative to latitude to emphasize the weak association.  We are confident that our exclusion of latitude from the analysis does not affect our conclusions.

### 2018-10-02 doc/prcompex.Rmd
no more lat in pc cor plot

@@
     keep_md: yes
-self_contained: no
+    self_contained: yes
 code_folding: hide
@@
 ```{r, echo = F, warning = F, message = F}
+library(knitr)
+opts_chunk$set(fig.align = 'center', message = F, echo = T, cache = F, dev = 'png', dev.args = list(family = 'serif'), dpi = 400, fig.pos = '!h', warning = F, background = 'white', out.width = '100%',
+               fig.process = function(x) {
+                 x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
+                 if (file.rename(x, x2)) x2 else x
+               })
 
-knitr::opts_chunk$set(echo = F, warning = F, message = F)
 library(tidyverse)
@@
   column_to_rownames('CTD') %>% 
-  select(-Ara, -pCO2, -O2, -pH)
+  select(-Ara, -pCO2, -O2, -pH, -Lat)
 
@@
 
-levs <- c("Fluor", "Lat", "PC1", "Temp", sort(biochr), sort(phychr))
-labs <- c('Chla', 'Lat', 'PC1', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
+levs <- c("Fluor", "PC1", "Temp", sort(biochr), sort(phychr))
+labs <- c('Chla', 'PC1', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
 prplo <- crs %>%
@@
 outlab <- data.frame(
-  y = c(3.5, 10.5, 16.5),
+  y = c(3.5, 10.5, 16),
   lab = c('Population/\nPhysiology', 'Cellular', 'Environment')
@@
   annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
-                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 19, xmax = 19) +
+                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 18, xmax = 18) +
   annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
-                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 19, xmax = 19) +
+                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 18, xmax = 18) +
   annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
-                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 19, xmax = 19) +
+                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 18, xmax = 18) +
   annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
-                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 19.5, ymax = 19.5) +
+                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 18.5, ymax = 18.5) +
   annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
-                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 19.5, ymax = 19.5) +
+                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 18.5, ymax = 18.5) +
   annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
-                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 19.5, ymax = 19.5) +
+                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 18.5, ymax = 18.5) +
   pbase +

### 2018-10-09 doc/reviewer2_response2.Rmd
captions for pc figs

@@
 ![](prcompex_files/figure-html/pcaplo.png)
+Figure: Results of principal components analysis of the environmental data.
+The first plot shows the estimated axis scores for the first and second principal components and the second plot shows the scores for the second and third principal components.
+The first three axes explained nearly 99% of the variation among the observations.
+Length and direction of each vector shows the relative loadings of each variable with the axes.
+
 The first axis above was used in place of the four correlated OA parameters in the original Figure 5:
@@
 ![](prcompex_files/figure-html/corploPC1.png)
+Figure: Modified correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.
+Correlated environmental variables related to ocean acidification were combined into a single principal component axis represented by PC1.
+Darker red values are strong positive correlations and darker purple values are strong negative correlations.
+Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*\*).
 

### 2019-01-21 biomark.Rmd
added oracvlpx, len, o2 v ara for abund
replaced error message in app with validate message

@@
 
+## ORACvLPX
+```{r}
+knitr::kable(filter(mods, pte_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
+```
+
 ## SOD
@@
 
-  if(is.null(trm())) stop("Model is too simple, you broke me Nina!")
+  validate(
+    need(!is.null(trm()), 'Cannot show plots for models with one predictor')
+  )
 
@@
   toplo <- md()$model
-  if(ncol(toplo) < 3) stop("Model is too simple, you broke me Nina!")
+  req(ncol(toplo) >= 3)
 

### 2021-08-10 exposure.Rmd
exposure rmd

---
output:
  html_document:
    self_contained: no
    code_folding: hide
    css: styles.css
---

```{r setup, warning = F, message = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(patchwork)
library(here)

data(expdat)
```

```{r, results = 'hide'}
toplo <- expdat %>% 
  unite('temp, pCO2', temp_trt, pco2_trt, sep = '/') %>% 
  mutate(
    `temp, pCO2` = factor(`temp, pCO2`, levels = c('lo/lo', 'lo/hi', 'hi/lo' ,'hi/hi'))
  )

p1 <- ggplot(toplo, aes(x = `temp, pCO2`, y = alive)) +
  scale_y_continuous('% alive') +
  geom_boxplot(outlier.size = -1) +
  geom_jitter(alpha = 0.3, width = 0.15) + 
  theme_bw(base_family = 'serif', base_size = 14) + 
  labs(x = 'Temp, pCO2')

toplo <- expdat %>% 
  group_by(temp_trt, pco2_trt) %>% 
  summarize(
    ave = mean(alive, na.rm = T),
    lov = t.test(alive)$conf.int[1],
    hiv = t.test(alive)$conf.int[2]
  ) %>% 
  ungroup %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  )

allv <- expdat %>% 
  mutate(
    temp_trt = factor(temp_trt, levels = c('lo', 'hi')),
    pco2_trt = factor(pco2_trt, levels = c('lo', 'hi'))
  ) %>%
  select(alive, temp_trt, pco2_trt, CTD) %>% 
  group_by(temp_trt, pco2_trt, CTD) %>%
  mutate(
    rep = 1, 
    rep = cumsum(rep)
  ) %>% 
  unite('uni', CTD, temp_trt, rep, remove = F)

p2 <- ggplot(toplo, aes(x = pco2_trt, colour = temp_trt)) + 
  # geom_line(data = allv, aes(y = alive, group = uni), linetype = 'dotted') +
  geom_point(size = 5, aes(y = ave)) + 
  geom_errorbar(aes(ymin = lov, ymax = hiv), width = 0.1, size = 1.5) +
  geom_line(aes(y = ave, group = temp_trt), size = 1.5) +
  theme_bw(base_family = 'serif', base_size = 14) + 
  labs(
    x = 'pCO2', 
    y = '% alive', 
    colour = 'Temp'
  ) +
  theme(legend.position = c(0.9, 0.2))

p <- p1 + p2 + plot_layout(ncol = 1)

png(here('raw/survive.png'), height = 6.5, width = 6.5, family = 'serif', res = 300, units = 'in')
print(p)
dev.off()
```

[download figure](raw/survive.png)

```{r, out.width = '80%'}
knitr::include_graphics(here('raw/survive.png'))
```

Run a simple test:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt * pco2_trt, data = expdat)
summary(mod)
anova(mod)
```

No interaction:
```{r}
mod <- lm(asin(alive / 100) ~ temp_trt + pco2_trt, data = expdat)
summary(mod)
anova(mod)
```
